### 2021-07-23 Old scripts/pupil_arousal_datarmd.Rmd
started working on issue #1

---
title: "R Notebook"
output: html_notebook
---

Downsample data to save space

pupiltimecourse.data is dowsnampled but doesn't have all participants < 400

but the script pupiltimecourseexp2.Rmd has what we need to do the same dowsnampling

Goal
- take pupil timecourse, dowsnample and then mrge it it behavioural data so that we can do the additional arousal pupil analyses that check for timecourse things.

check gazeR or downsampling

to do, do the same arousal correlation with valence
```{r}
library(ggplot2)
library(tidyverse)
install.packages("gazer")
library(gazer)

?gazer::downsample_gaze
?gazer::downsample_pupil

unique(X_tmp_df4_full$Condition)
# gazer::downsample_gaze%

tmp.df4_full_stim<- X_tmp_df4_full %>%
  subset(screencontent == "stim")

# create bins
# install.packages("OneR")
install.packages("OneR")
library(OneR)

?bin

tmp.df4_full_stim$timebin<- OneR::bin(tmp.df4_full_stim$timerezero3, nbin = 60)

# dowsample by averaging?
colnames(tmp.df4_full_stim)
tmp.df4_full_stim_downs <- tmp.df4_full_stim %>%
  group_by(ssid, tNo, timebin, trial,trialUnq)%>%
  summarise_at(c('timerezero3','pup_basCor', 'gaze_x_cor_pix', 'gaze_y_cor_pix'), 
               mean, na.rm = TRUE)

?saveRDS
saveRDS(tmp.df4_full_stim_downs, "tmp.df4_full_stim_downs.rds")

# create klarger bins
tmp.df4_full_stim_downs_jun2021_with_beh

range(tmp.df4_full_stim_downs$timerezero3)

tmp.df4_full_stim_downs_jun2021_with_beh$timebin2<- OneR::bin(tmp.df4_full_stim_downs_jun2021_with_beh$timerezero3, nbin = 30)

tmp.df4_full_stim_downs_jun2021_with_beh2 <- tmp.df4_full_stim_downs_jun2021_with_beh %>%
  group_by(ssid, tNo, timebin2, trial,trialUnq)%>%
  summarise_at(c('timerezero3','pup_basCor', 'gaze_x_cor_pix', 'gaze_y_cor_pix'), 
               mean, na.rm = TRUE)

nrow(tmp.df4_full_stim_downs_jun2021_with_beh2)
87529
```

pupil arousal

```{r}
library(gghalves)

# pupil tracks arousal

db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, ground_arousal_lab, Alexithymia, mediansplit_ground_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ground_arousal_lab, pup_basCor, color = ground_arousal_lab,fill = ground_arousal_lab))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")
  facet_grid(~mediansplit_ground_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, mediansplit_self_arousal, Alexithymia, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_self_arousal, pup_basCor, color = mediansplit_self_arousal,fill = mediansplit_self_arousal))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  # xlab("Arousal")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

mediansplit_self_arousal_thissample

  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, mediansplit_self_arousal_thissample, Alexithymia, mediansplit_self_valence_thissample)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_self_arousal_thissample, pup_basCor, color = mediansplit_self_arousal_thissample,fill = mediansplit_self_arousal_thissample))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  # xlab("Arousal")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence_thissample)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

  db_full4new_stim_screen_pupil_nopract%>%

     ggplot(aes( valence,arousal))+
    geom_point()+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)+
    facet_grid(~mediansplit_self_arousal)
    # geom_smooth(aes(x = arousal^2, y = valence))
   db_full4new_stim_screen_pupil_nopract%>%

    ggplot(aes( valence,arousal))+
    geom_point()+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)+
    facet_grid(~mediansplit_self_valence)

   unique(db_full4new_stim_screen_pupil_nopract$ssid)
```

```{r}
# mediansplit_self_valence is duoliating the stimuli based ion peopl who choose negative and positive
unique(db_full4new_stim_screen_pupil_nopract)
db_full4new_stim_screen_pupil_nopract <- db_full4new_stim_screen_pupil_nopract%>%
  ungroup()%>%
  # group_by(stimIAPS)%>%
  mutate(mediansplit_sample_valence2 = if_else(ValenceMeanThisSample >=
                                                  median(ValenceMeanThisSample,
                                                         na.rm = TRUE), "More positive", "More negative"))%>%
  mutate(mediansplit_sample_arousal2 = if_else(ArousalMeanThisSample >=
                                                  median(ArousalMeanThisSample,
                                                         na.rm = TRUE), "High", "Low"))

  group_by(mediansplit_sample_valence2, stimIAPS,stimDescription)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

```

brightness

```{r}
tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  group_by(stimIAPS)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ValenceMean, Mean_gray_z))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  group_by(stimIAPS)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ArousalMean, Mean_gray_z))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  group_by(stimIAPS,ground_valence_lab)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ground_valence_lab, Mean_gray_z))+
  geom_point()+
  stat_summary(geom = 'pointrange')
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

  tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  group_by(stimIAPS,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, Mean_gray_z))+
  # geom_point()+
  stat_summary(geom = 'pointrange')
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

    tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  group_by(stimIAPS,ground_valence_lab)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(Mean_gray_z, color = ground_valence_lab))+
      # geom_histogram()+
      geom_density()
  # geom_point()+
  stat_summary(aes(group = mediansplit_ground_valence),geom = 'line')
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

  #

tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  group_by(stimIAPS,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, ArousalMean))+
      # geom_histog `vram()
      # geom_density()
  # geom_point()+
  stat_summary(geom = 'pointrange')
  # geom_smooth(method = 'lm', se = F)+
  # ggpubr::stat_cor()

      ground_valence_lab
```

Participants section

```{r}
participants_demog <- db_full4new_stim_screen_pupil_nopract %>%
    # subset(pp_arousal_outler== FALSE)%>%
  subset(Group == "NT")%>%
  group_by(ssid, Gender)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

unique(participants_demog)

# gender
table(participants_demog$Gender)

colnames(participants_demog)
psych::describe(participants_demog[c("Age")])

summary(participants_demog$Age)

sd(participants_demog$Age, na.rm = TRUE)

table(is.na(participants_demog$Age))

participants_demog %>%
  group_by(Gender)%>%
  mutate(genderspecif_age = sd(TAS))%>%
  ungroup()%>%
  mutate(nongenderspecif_age = sd(TAS))%>%
  group_by(Gender)%>%
  summarise_at(c("genderspecif_age","nongenderspecif_age"), mean, na.rm = TRUE)

(10.557140+8.042743)/2

8.042743

9.299941

```

```{r}
  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(Group == "NT")%>%
  # subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, mediansplit_ground_arousal, Alexithymia, mediansplit_ground_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_arousal, pup_basCor, color = mediansplit_ground_arousal,fill = mediansplit_ground_arousal))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  # xlab("Arousal")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_ground_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)
```

```{r}

    db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, mediansplit_ground_arousal, Alexithymia, mediansplit_ground_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_arousal, Mean, color = mediansplit_ground_arousal,fill = mediansplit_ground_arousal))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  # xlab("Arousal")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_ground_valence)+
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(90,130)

```

  okay let's create averages frio this entire sample and use those to devide the stimuli

  self aropusal might be biased by problematic folks but averages for the entire sample should not

```{r}

db_full4new_stim_screen_pupil_nopract %>%
  group_by(stimIAPS)%>%
  mutate()

  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid_num< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, self_arousal_lab, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(self_arousal_lab, pup_basCor, color = self_arousal_lab,fill = self_arousal_lab))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  # xlab("Arousal")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)

sample_iaps_arousal<-  db_full4new_stim_screen_pupil_nopract %>%

    group_by(stimIAPS)%>%
    subset(ssid_num< 500)%>%
    summarise_at(c('ArousalMeanThisSample','ArousalMean'), mean, na.rm = TRUE)%>%
    ggplot(aes(ArousalMeanThisSample, ArousalMean))+
    geom_point()+
    geom_smooth(method = 'lm', se = F)+
    ggpubr::stat_cor()+
    theme_classic()

  sample_iaps_val<-   db_full4new_stim_screen_pupil_nopract %>%

     group_by(stimIAPS)%>%
    subset(ssid_num< 500)%>%
    summarise_at(c('ValenceMeanThisSample','ValenceMean'), mean, na.rm = TRUE)%>%
    ggplot(aes(ValenceMeanThisSample, ValenceMean))+
    geom_point()+
    geom_smooth(method = 'lm', se = F)+
    ggpubr::stat_cor()+
      theme_classic()

library(patchwork)
sample_iaps_arousal+
sample_iaps_val
```

```{r} 
  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid_num< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, self_arousal_lab, Alexithymia, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(self_arousal_lab, pup_basCor, color = self_arousal_lab,fill = self_arousal_lab))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid_num< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, ground_arousal_lab, Alexithymia, mediansplit_ground_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ground_arousal_lab, pup_basCor, color = ground_arousal_lab,fill = ground_arousal_lab))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")
  facet_grid(~mediansplit_ground_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

```

Resposne bias
```{r}
library(ggExtra)
arous_pup_by_val<- db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(ssid, stimIAPS, mediansplit_self_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, pup_basCor))+
      geom_point(alpha = .2)+
    # stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)
  # ggpubr::stat_cor()
  ylim(-3.3,1.3)

  arous_pup_by_val<- arous_pup_by_val+ggside::geom_xsidedensity(aes(y=stat(density),fill = mediansplit_self_valence), alpha = .4)

  arous_pup_by_val+ggside::geom_ysidedensity(aes(x=stat(density),fill = mediansplit_self_valence), alpha = .4)

ggMarginal(arous_pup_by_val, type = "density")
```

ceiliing effect?

```{r}
db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(ssid, stimIAPS, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, pup_basCor))+
      geom_point()+
    # stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()+
  ylim(-3.3,1.3)

```

ceiling effect?
```{r}
db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(mediansplit_ground_valence)%>%
  mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  group_by(mediansplit_ground_valence)%>%
  mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, pup_basCor))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .4, size = 1, colour = "Black") +
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

```

```{r}  
  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(mediansplit_ground_valence)%>%
  mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  group_by(mediansplit_ground_valence)%>%
  mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, pup_basCor))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .4, size = 1, colour = "Black") +
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_self_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_self_valence, pup_basCor))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .5, size = 1, colour = "Black") +
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

```

```{r}

# db_full4new_stim_screen_pupil_nopract$arousal_outler
  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(ssid, mediansplit_sample_arousal2, mediansplit_sample_valence2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_sample_arousal2, arousal))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .5, size = 1, colour = "Black") +
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_sample_valence2)+
  ggpubr::stat_cor()

# stim
  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(stimIAPS, mediansplit_self_valence, mediansplit_self_arousal)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_self_arousal, arousal))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .5, size = 1, colour = "Black") +
  # stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)+
  ggpubr::stat_cor()

  # ground

  # stim

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  group_by(stimIAPS, mediansplit_ground_arousal, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_arousal, ArousalMean))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary( geom = 'pointrange', alpha = .5, size = 1, colour = "Black") +
  # stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()

```

otehr confounds - eye movements

```{r}
db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, mean_fix_dur))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

   unique(db_full4new_stim_screen_pupil_nopract$ValenceMeanThisSample)

   median(db_full4new_stim_screen_pupil_nopract$ValenceMeanThisSample)
    median(db_full4new_stim_screen_pupil_nopract$valence)

     median(db_full4new_stim_screen_pupil_nopract$arousal)
    median(db_full4new_stim_screen_pupil_nopract$ArousalMeanThisSample)

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_self_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_self_valence, mean_fix_dur))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

```

```{r}
  # data loss gaze

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, fix_count))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, SacAmp))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

  db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, sdevgaze))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

 db_full4new_stim_screen_pupil_nopract %>%
  subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, gaze_loss_prop))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()
  ylim(-3.3,1.3)

  db_full4new_stim_screen_pupil_nopract%>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE), sd_pup = sd(pup_basCor, na.rm = TRUE), subj = n())%>%
  group_by(ssid, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # group_by(mediansplit_ground_valence)%>%
  # mutate(pup_basCor_se = sd_pup/(sqrt(subj)))%>%
  ggplot(aes(mediansplit_ground_valence, total_fix_dur))+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
  stat_summary(geom = 'pointrange', alpha = .1, size = 1, colour = "Black") +
  # geom_errorbar(aes(ymin = pup_basCor_mean-pup_basCor_se, ymax = pup_basCor_mean+pup_basCor_se ))+
  geom_smooth(method = 'lm', se = F)+
  p$graphstyle+
  xlab("Valence")+
  # ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()

```

# let's model it formally

# remember baseline corection is equivalendt to stimuli intercept

```{r} 
pupil_arousal_findings<- list()
library(lmerTest)

unique(db_full4new_stim_screen_pupil_nopract$stimIAPS)
unique(substr(db_full4new_stim_screen_pupil_nopract$stimIAPS, 1,8))
db_full4new_stim_screen_pupil_nopract$Label<- substr(db_full4new_stim_screen_pupil_nopract$stimIAPS, 1,8)

db_full4new_stim_screen_pupil_nopract<- left_join(db_full4new_stim_screen_pupil_nopract, imageJ_IAPS, by = "Label")
nrow(db_full4new_stim_screen_pupil_nopract)
# rows = 2858

db_full4new_stim_screen_pupil_nopract$Mean_gray_z<- scale(db_full4new_stim_screen_pupil_nopract$Mean, center = TRUE, scale = TRUE)[,1]

options(contrasts = c("contr.sum","contr.poly"))
options(scipen = 999)

db_full4new_stim_screen_pupil_nopract$ssid_num <- as.numeric(as.character(db_full4new_stim_screen_pupil_nopract$ssid))

pupil_arousal_findings$pupil_from_arousal <- lmer(pup_basCor ~ arousal_c*Mean_gray_z  +(1 +Mean_gray_z | ssid) + (0+arousal_c| ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500 & 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_arousal)
# singularity (drop arousal c slope)

# is there a coorelationbtween mean arousal rating and meagray(

db_full4new_stim_screen_pupil_nopract%>%
  group_by(stimIAPS,ground_valence_lab) %>%
  summarise_at(c('arousal_c', 'Mean_gray_z'), mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, Mean_gray_z, color = ground_valence_lab))+
  geom_point()+
  ggpubr::stat_cor()

# <- predict(pupil_arousal_findings$pupil_from_arousal)
db_full4new_stim_screen_pupil_nopract %>%
  group_by(stimIAPS) %>%
  # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # summarise_at(c('arousal_c', 'Mean_gray_z'), mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, pup))+
  geom_smooth(aes(group = ssid, alpha = ssid), method = 'lm', se = F)
# geom_smooth(aes(group = stimIAPS), method = 'lm', se = F)
  ggpubr::stat_cor()

```
cor.test(db_full4new_stim_screen_pupil_nopract$, db_full4new_stim_screen_pupil_nopract$)

```{r}
# if we put brighteness and arousal we see no interaction which means we probably don't need to worry about brighteness, it doesn't change effects of arousal self report
# db_full4new_stim_pupil$pup
pupil_arousal_findings$pupil_from_arousal <- lmer(pup_basCor ~ arousal_c+Mean_gray_z  +(1 +Mean_gray_z | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract$pup, pupil_outlier == FALSE & ssid_num< 500& 
                                      arousal_outler == FALSE))

install.packages('usdm')

usdm::vif(as.data.frame(db_full4new_stim_screen_pupil_nopract[,c('arousal', 'valence')]))

pupil_arousal_findings$pupil_from_arousal_withval <- lmer(pup_basCor ~ arousal_c*valence_c* Mean_gray_z  +(1 +Mean_gray_z | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500& 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_arousal_withval)
car::vif(pupil_arousal_findings$pupil_from_arousal_withval)

lmerTest::step(pupil_arousal_findings$pupil_from_arousal_withval)

# control for baseline arousal ratings

pupil_arousal_findings$pupil_from_arousal_withval

summary(lmer(pup_basCor ~ arousal_c*valence_c+ Mean_gray_z  +(1 | ssid) + (1|stimIAPS),
               # (1|ArousalMean)+ (1|ArousalMean), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & Group == "NT"  & 
                                      arousal_outler == FALSE)))

# arousal_c:valence_c                      0  35.52   35.52     1 2150.19  13.5760 0.0002348 ***

pupil_arousal_findings$arousal_from_pupil <- lmer(arousal ~ pup_basCor * Mean_gray_z  +(1+pup_basCor  | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid!= 610& 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$arousal_from_pupil)
# what about valence

pupil_arousal_findings$pupil_from_valence <- lmer(pup_basCor ~ valence_c+Mean_gray_z  +(1 +Mean_gray_z | ssid) + 
                                                    (0 +valence_c | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid!= 610& 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_arousal)
summary(pupil_arousal_findings$pupil_from_valence)

```

# Valence and arousal
```{r}

pupil_arousal_findings$pupil_from_ar_vale <- lmer(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z  +
                                                    (1 + Mean_gray_z | ssid),
                                                    # (0+ valence_c* arousal_c | ssid),
                                                  REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500 & 
                                      arousal_outler == FALSE))

db_full4new_stim_screen_pupil_nopract$total_fix_dur_z<- scale(db_full4new_stim_screen_pupil_nopract$total_fix_dur, center = TRUE, scale = TRUE)[,1]

pupil_arousal_findings$pupil_from_ar_val_gaze <- lmer(pup_basCor ~ (valence_c * arousal_c * total_fix_dur_z)+ Mean_gray_z  +
                                                        +
                                                    (1 + Mean_gray_z | ssid),
                                                    # (0+  total_fix_dur_z | ssid),
                                                  REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500 & 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_ar_val_gaze )
summary(pupil_arousal_findings$pupil_from_ar_vale)

car::vif(pupil_arousal_findings$pupil_from_ar_val_gaze)

anova(pupil_arousal_findings$pupil_from_ar_vale, pupil_arousal_findings$pupil_from_arousal)
# interaction valence and arousal

```

control for gaze

```{r}
# avg fix dur
db_full4new_stim_screen_pupil_nopract$mean_fix_dur_z<- scale(db_full4new_stim_screen_pupil_nopract$mean_fix_dur, center = TRUE, scale = TRUE)[,1]

pupil_arousal_findings$pupil_from_ar_val_gaze <- lmer(pup_basCor ~ (valence_c * arousal_c * mean_fix_dur_z)+ Mean_gray_z  +
                                                        +
                                                    (1 + Mean_gray_z | ssid),
                                                    # (0+  total_fix_dur_z | ssid),
                                                  REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500 & 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_ar_val_gaze )
interactions::interact_plot(pupil_arousal_findings$pupil_from_ar_val_gaze , pred = arousal_c, modx = mean_fix_dur_z)

# fix count

db_full4new_stim_screen_pupil_nopract$fix_count_z<- scale(db_full4new_stim_screen_pupil_nopract$fix_count, center = TRUE, scale = TRUE)[,1]

pupil_arousal_findings$pupil_from_ar_val_gaze <- lmer(pup_basCor ~ (valence_c * arousal_c * fix_count_z)+ Mean_gray_z  +
                                                        +
                                                    (1 + Mean_gray_z | ssid),
                                                    # (0+  total_fix_dur_z | ssid),
                                                  REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid_num< 500 & 
                                      arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_ar_val_gaze )

interactions::interact_plot(pupil_arousal_findings$pupil_from_ar_val_gaze , pred = arousal_c, modx = mean_fix_dur_z)
```

```{r}
interactions::interact_plot(pupil_arousal_findings$pupil_from_ar_vale , pred = valence_c, modx = arousal_c)

interactions::interact_plot(pupil_arousal_findings$pupil_from_ar_vale , pred =arousal_c , modx =  valence_c)

car::vif(pupil_arousal_findings$pupil_from_ar_vale) #curoff 2.5?

cor.test(db_full4new_stim_screen_pupil_nopract$valence_c, db_full4new_stim_screen_pupil_nopract$arousal_c)

rmcorr_valence_arousal <- rmcorr::rmcorr(participant = "ssid", measure1 = "valence_c", measure2 = "arousal_c", data = db_full4new_stim_screen_pupil_nopract)

plot(rmcorr_valence_arousal)

?rmcorr

db_full4new_stim_screen_pupil_nopract %>%
  group_by(stimIAPS)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ArousalMean, ValenceMean))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  theme_classic()

db_full4new_stim_screen_pupil_nopract %>%
  group_by(stimIAPS)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, valence_c))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  theme_classic()

```

```{r}
# with arousal as DV
pupil_arousal_findings$aroual_from_pup <- lmer(arousal ~ valence_c*pup_basCor *  BRIGHTNESSc  + (1| stimIAPS)+
                                                 (1 | ssid)+(0 +valence_c | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, pupil_outlier == FALSE & ssid!= 610& 
                                      arousal_outler == FALSE))
summary(pupil_arousal_findings$aroual_from_pup)

anova(pupil_arousal_findings$aroual_from_pup)

interactions::interact_plot(pupil_arousal_findings$aroual_from_pup , pred = valence_c, modx = pup_basCor)

interactions::interact_plot(pupil_arousal_findings$aroual_from_pup, pred =pup_basCor , modx =  valence_c)
# for positive valenced stimuli participants reporte higher arousal to more dilated pupils
#for negatively valence they report feeling negative even when they don't show strog pupil dilation
# same pattern
# pupil is more alighned with objective arousal in the positive trials

```

mixed model is consistent wth the previous plot, arousal effects on pupil is modulated by valence. in other words there is a consistent effect between pupil and self-repot of arousal more for positive emotions.

analyse the correlationla way
since we are aggregating we don't need to worry too much about brighteness as every person sees every stimuli in all brightness levels
CONCORDANCE

```{r}

db_full4new_stim_screen_pupil_nopract%>%
  group_by(mediansplit_self_valence_thissample, stimIAPS)%>%
  summarise_at(c("arousal", 'valence'), mean, na.rm = TRUE)%>%
  arrange(stimIAPS)

```

````{r}

test_anovas<- db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_valence2,mediansplit_sample_arousal2)%>%
    mutate(cortest = cor(arousal_c, pup_basCor, use = "complete"))%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)

 db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_valence2,mediansplit_sample_arousal2)%>%
    mutate(cortest = cor(arousal_c, pup_basCor, use = "complete"))%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_sample_valence2, cortest, color = mediansplit_sample_valence2))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
  theme_classic()+
  geom_hline(yintercept = .2, linetype = 'dashed', alpha = .2)+
   # ggpubr::stat_compare_means(paired = TRUE)+
  ggpubr::stat_compare_means()+
  facet_grid(~mediansplit_sample_arousal2)

  # let's identify the folks that show less of a difference between conditions.

flag_pos_neg<-  db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_valence2)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor)$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
    mutate(cor_pos_neg_diff = cortest[mediansplit_sample_valence2 == "More positive"] -
             cortest[mediansplit_sample_valence2 == "More negative"])%>%
    mutate(cor_pos_neg_diff_lab = if_else(cor_pos_neg_diff>0, "pos > neg", "neg > pos"))%>%
     group_by(ssid,cor_pos_neg_diff_lab)%>%
    summarise_at(c("cor_pos_neg_diff"), mean, na.rm = TRUE)

    # subset(cor_pos_neg_diff<= 0)%>%
        ggplot(aes(mediansplit_sample_valence2, cortest, color = mediansplit_sample_valence2))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
  theme_classic()+
  geom_hline(yintercept = .2, linetype = 'dashed', alpha = .2)
   # ggpubr::stat_compare_means(paired = TRUE)+
  ggpubr::stat_compare_means()

```

```{r}
testcor <-cor(db_full4new_stim_screen_pupil_nopract$arousal, db_full4new_stim_screen_pupil_nopract$valence)

testcor 
db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_valence2, mediansplit_sample_arousal2)%>%
    # mutate(cortest = cor.test(arousal_c, pup_basCor)$estimate)%>%
   mutate(cortest = cor(arousal_c, pup_basCor, method = 'kendall'))%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_sample_valence2, cortest, color = mediansplit_sample_valence2))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
  theme_classic()+
   facet_grid(~mediansplit_sample_arousal2)+
  geom_hline(yintercept = .2, linetype = 'dashed', alpha = .2)
   # ggpubr::stat_compare_means(paired = TRUE)+
  # ggpubr::stat_compare_means()

db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_arousal2, mediansplit_sample_valence2)%>%
    # mutate(cortest = cor.test(arousal_c, pup_basCor)$estimate)%>%
   mutate(cortest = cor(valence_c, pup_basCor, method = 'kendall'))%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_sample_arousal2, cortest, color = mediansplit_sample_arousal2))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
  theme_classic()+
   facet_grid(~mediansplit_sample_valence2)+
  geom_hline(yintercept = .2, linetype = 'dashed', alpha = .2)

```

# same as plot above but with valence in the correlation
```{r}
db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_sample_arousal2)%>%
    mutate(cortest = cor.test(valence_c, pup_basCor)$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_sample_arousal2, cortest, color = mediansplit_sample_arousal2))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
  theme_classic()
   # ggpubr::stat_compare_means(paired = TRUE)+
  ggpubr::stat_compare_means()

```

```{r}
# ground

 db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_ground_valence)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor, use = "complete")$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_ground_valence, cortest, color = mediansplit_ground_valence))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .3)+
        stat_summary( geom = 'pointrange')+
        stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
   # ggpubr::stat_compare_means(paired = TRUE)+
  # ggpubr::stat_compare_means()
   theme_classic()
```

```{r}
 # ground arousal with valence in the  corelation

 db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_ground_valence)%>%
    mutate(cortest = cor.test(valence_c, pup_basCor, use = "complete")$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_ground_valence, cortest, color = mediansplit_ground_valence))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .3)+
        stat_summary( geom = 'pointrange')+
        stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
   # ggpubr::stat_compare_means(paired = TRUE)+
  # ggpubr::stat_compare_means()
   theme_classic()

 db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_ground_valence, mediansplit_ground_arousal)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor, use = "complete")$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_ground_valence, cortest, color = mediansplit_ground_valence))+
           geom_bar(stat="summary", fun.y = "mean", alpha = .4)+

    geom_jitter(width = .1, alpha = .3)+
        stat_summary( geom = 'pointrange')+
        stat_summary(aes(group = ssid), geom = 'line', color = "black", alpha = .1)+
    ylab("R pupil vs self report arousal")+
   # ggpubr::stat_compare_means(paired = TRUE)+
  # ggpubr::stat_compare_means()
   theme_classic()+
   facet_grid(~mediansplit_ground_arousal)

```

```{r}
db_pupil_concordance_agg_ground <-  db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_ground_valence)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor, use = "complete")$estimate)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)

afex::aov_ez(id = "ssid",
             data = db_pupil_concordance_agg,
             between = NULL,
             within = 'mediansplit_ground_valence',
             dv = 'cortest')

unique(db_full4new_stim_screen_pupil_nopract$tNo)
db_pupil_concordance_agg_self <-  db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE)%>%
  subset( arousal_outler == FALSE & !is.na(Alexithymia))%>%
  subset(!is.na(Alexithymia))%>%
  # subset(Group == "NT")%>%
  subset(ssid_num<500)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_self_arousal, mediansplit_self_valence)%>%
    mutate(cortest = cor(arousal_c, pup_basCor, use = "complete"))%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)

afex::aov_ez(id = "ssid",
             data = test_anovas,
             between = NULL,
             within = 'mediansplit_sample_valence2',
             dv = 'cortest')

afex::aov_ez(id = "ssid",
             data = subset(test_anovas, mediansplit_sample_valence2 == "More positive"),
             between = NULL,
             within = 'mediansplit_sample_arousal2',
             dv = 'cortest')

afex::aov_ez(id = "ssid",
             data = db_pupil_concordance_agg_self,
             between = NULL,
             within = 'mediansplit_ground_valence',
             dv = 'cortest')

# valence and arousal

afex::aov_ez(id = "ssid",
             data = test_anovas,
             between = NULL,
             within = c('mediansplit_sample_arousal2','mediansplit_sample_valence2'),
             dv = 'cortest')

db_parsed_f_summ<- db_parsed_f %>%
  group_by(tNo, ssid)%>%
  summarise_at(c('Duration', 'mean_x', 'mean_y'), sum, na.rm = TRUE)
  ggplot(aes(tNo, Duration))+
  stat_summary(geom = 'pointrange')+
  geom_point(alpha = .1)

unique(db_parsed_f_summ$ssid)
unique(db_parsed_f_summ$tNo)
unique(db_full4new_stim_screen_pupil_nopract$tNo)
unique(db_full4new_stim_screen_pupil_nopract$ssid)
nrow(db_full4new_stim_screen_pupil_nopract)
nrow(db_parsed_f_summ)

table(is.na(db_parsed_f_summ$ssid))

db_parsed_f_summ$total_fix_dur<- db_parsed_f_summ$Duration

db_parsed_f_summ$tNo<- as.character(db_parsed_f_summ$tNo)
db_parsed_f_summ$ssid<- as.character(db_parsed_f_summ$ssid)
db_full4new_stim_screen_pupil_nopract$Duration<- NULL
db_full4new_stim_screen_pupil_nopract$total_fix_dur<- NULL

db_full4new_stim_screen_pupil_nopract<- left_join(db_full4new_stim_screen_pupil_nopract, db_parsed_f_summ)

```

I HAVE THE IMPRESSION THAT WHAT THIS ANALYSES GIVES US IS THE SAME AS THE MIXED MODEL AS THIS CONCIORDANCE CAN BE EXPRESSED IN TERMS FO THE SLOPE OF THE AOUSAL AND PUPIL RELATION

SO ION OTHER WORDS IS not that pupil tracksvalence is that arousal based pupil is modulated by valence

by why is that
- bias is responding? eg.. people response they feel negative even when they don't (ie. when we have no objective indicator that they do feel)
- is it something about the stimuli, positive arousal is sexual here?
- or arousal interferes with appraisal. high arousing stimuli are distracting??

timecourse models of pupil

Pupil tracks valence

```{r}

db_full4new_stim_screen_pupil_Anopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
  group_by(ssid)%>%
  mutate(pup_basCor_z = scale(pup_basCor, center = TRUE, scale = TRUE))%>%
    group_by(ssid, ground_valence_lab, Alexithymia)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(ground_valence_lab, pup_basCor_z,fill = ground_valence_lab))+
 geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
    # geom_ysidedensity(aes(x=stat(density),color = ground_valence_lab), alpha = .4)+
  p$graphstyle+
  xlab("valence")+
  ylab("Baseline corrected pupil")+
  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")

mediansplit_self_valence

db_full4new_stim_screen_pupil_nopract %>%
  # subset(pp_arousal_outler== FALSE)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
  subset(ssid_num<500)%>%
  group_by(ssid)%>%
  # mutate(pup_basCor_z = scale(pup_basCor, center = TRUE, scale = TRUE))%>%
    group_by(ssid, self_arousal_lab, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(self_arousal_lab, pup_basCor,fill = self_arousal_lab))+
 geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
    # geom_ysidedensity(aes(x=stat(density),color = ground_valence_lab), alpha = .4)+
  p$graphstyle+
  xlab("arousal")+
  ylab("Baseline corrected pupil")+
  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)

  db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler== FALSE)%>%
  subset(ssid< 500)%>%
  subset(pupil_outlier == FALSE)%>%
  subset(!is.na(Alexithymia))%>%
    group_by(ssid, self_arousal_lab, Alexithymia, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(self_arousal_lab, pup_basCor, color = self_arousal_lab,fill = self_arousal_lab))+
    geom_hline(yintercept = 0, linetype = "dashed", size = 2, alpha = .2)+
 # geom_half_violin(colour = FALSE, alpha = .3, size = 5)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .4, color = FALSE)+
      geom_jitter(width = .1, alpha=.2)+
    stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = 1, colour = "Black") +
  stat_summary(geom = 'pointrange', width =1)+
  p$graphstyle+
  xlab("Arousal")+
  ylab("Baseline corrected pupil")+

  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(~mediansplit_self_valence)
  # facet_grid(~BRIGHTNESS_mediansplit)+
  ylim(-6,2.3)

table(is.na(db_full4new_stim_screen_pupil_nopract$mediansplit_self_valence))
```

combine timecourse

let's put pupil scr and ht prediction on subjective arousal

```{r}
rm(justone_test, lucy_plots, test)

# import the timecourse

tmp.df4_full_stim_downs_jun2021 <- readRDS("~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/tmp.df4_full_stim_downs_jun2021.rds")

# merge with behavioural data
unique(db_full4new_stim_screen_pupil_nopract$ssid)
colnames(db_full4new_stim_screen_pupil_nopract)
colnames(tmp.df4_full_stim_downs_jun2021)

db_full4new_stim
colnames(db_full4new_stim_screen_pupil_nopract)
db_full4new_stim_subset<- db_full4new_stim_screen_pupil_nopract[,c(1:4,6:7,12:13,16,21:27, 31:56,80:81,84:96,98:114,116:124)]

# match types of merging variables
tmp.df4_full_stim_downs_jun2021_with_beh2$ssid<- as.character(tmp.df4_full_stim_downs_jun2021_with_beh2$ssid)

db_full4new_stim_subset$ssid<- as.character(db_full4new_stim_subset$ssid)

# tNo

tmp.df4_full_stim_downs_jun2021_with_beh2$tNo <- as.character(tmp.df4_full_stim_downs_jun2021_with_beh2$tNo)
db_full4new_stim_subset$tNo <- as.character(db_full4new_stim_subset$tNo)
unique(db_full4new_stim_subset$tNo)
unique(tmp.df4_full_stim_downs_jun2021_with_beh2$tNo)
# merge
nrow(tmp.df4_full_stim_downs_jun2021_with_beh2)
# 172168
# 87529

table(is.na(db_full4new_stim_subset$mediansplit_self_valence))

tmp.df4_full_stim_downs_jun2021_with_beh2$tNo<- as.numeric(tmp.df4_full_stim_downs_jun2021_with_beh2$tNo)
tmp.df4_full_stim_downs_jun2021_with_beh2<- tmp.df4_full_stim_downs_jun2021_with_beh2%>%
  arrange(ssid,tNo, timerezero3)

db_full4new_stim_subset$t
tmp.df4_full_stim_downs_jun2021_with_beh<- left_join(tmp.df4_full_stim_downs_jun2021_with_beh2, db_full4new_stim_subset)

unique(db_full4new_stim_subset$stimIAPS)

tmp.df4_full_stim_downs_jun2021_with_beh2$trialUnq

unique(tmp.df4_full_stim_downs_jun2021_with_beh2$ssid)
unique(db_full4new_stim_subset$ssid)

db_full4new_stim_subset$tNo<- as.numeric(db_full4new_stim_subset$tNo)

table(is.na(db_full4new_stim_subset$arousal))

tmp.df4_full_stim_downs_jun2021_with_behnew<- right_join(tmp.df4_full_stim_downs_jun2021_with_beh2, db_full4new_stim_subset,
                                                     by = c('ssid', 'tNo'))

nrow(tmp.df4_full_stim_downs_jun2021_with_beh)
# 172168 good

nrow(tmp.df4_full_stim_downs_jun2021_with_beh2)

nrow(tmp.df4_full_stim_downs_jun2021_with_behnew)
87529

table(is.na(tmp.df4_full_stim_downs_jun2021_with_behnew$mediansplit_self_valence))

subset(tmp.df4_full_stim_downs_jun2021_with_behnew, is.na(arousal))

unique(tmp.df4_full_stim_downs_jun2021_with_beh2$trialUnq)

```

```{r}
# check
tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(ground_arousal_lab))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= ground_arousal_lab, color = ground_arousal_lab), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=ground_arousal_lab, color = ground_arousal_lab), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   facet_grid(~mediansplit_self_valence)

```

pupil timecourse

```{r}

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= mediansplit_sample_valence2, color = mediansplit_sample_valence2), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=mediansplit_sample_valence2, color = mediansplit_sample_valence2), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   facet_grid(~mediansplit_sample_arousal2)

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= mediansplit_ground_arousal, color = mediansplit_ground_arousal), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=mediansplit_ground_arousal, color = mediansplit_ground_arousal), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   facet_grid(~mediansplit_ground_valence)

```

unique(tmp.df4_full_stim_downs_jun2021_with_beh$condition)

```{r}
tmp.df4_full_stim_downs_jun2021_with_beh$valencearousal_outliers<- if_else(tmp.df4_full_stim_downs_jun2021_with_beh$ArousalMean< 6.7 & 
                                                                    tmp.df4_full_stim_downs_jun2021_with_beh$mediansplit_ground_valence == "More negative", TRUE,
                                                                    if_else(tmp.df4_full_stim_downs_jun2021_with_beh$ArousalMean> 2.94 & 
                                                                    tmp.df4_full_stim_downs_jun2021_with_beh$mediansplit_ground_valence == "More positive", TRUE, FALSE))

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  # subset(valencearousal_outliers == TRUE) %>%
  group_by(stimIAPS,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, ArousalMean))+
  # geom_boxplot()+
  # geom_point()+
  # geom_text(aes(label = stimIAPS))+
  stat_summary(geom = 'pointrange')+
  geom_hline(yintercept = 7.2)+
    geom_hline(yintercept = 6.7)+
   geom_hline(yintercept = 2.94)

# thismplot raises question regarding wthtehr the better correlation bwteen subjective and objective arousal for positive stomuli reflects the fact that for negative we don't have many low arousal ones. But this is unlikelly because the oposite us also true for posituive stimuli, notmany positive stomuli are rated as highly arousing in sekf report, so we would not expect

# less restruct varoability
tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  subset(valencearousal_outliers == TRUE) %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= mediansplit_sample_arousal2, color = mediansplit_sample_arousal2), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=mediansplit_sample_arousal2, color = mediansplit_sample_arousal2), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   facet_grid(~mediansplit_sample_valence2)

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  subset(valencearousal_outliers == TRUE) %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= mediansplit_sample_arousal2, color = mediansplit_sample_arousal2), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=mediansplit_sample_arousal2, color = mediansplit_sample_arousal2), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   facet_grid(mediansplit_sample_valence2~.)+
  geom_hline(yintercept = 0)

```

unique(tmp.df4_full_stim_downs_jun2021_with_beh$timebin2)
# 30 time bins

```{r}
# pval<- cor.test(tmp.df4_full_stim_downs_jun2021_with_beh$sdevgaze_parsed, tmp.df4_full_stim_downs_jun2021_with_beh$pup_basCor)
# pval$estimate

tmp.df4_full_stim_downs_jun2021_with_beh$ssid_num<- as.numeric(as.character(tmp.df4_full_stim_downs_jun2021_with_beh$ssid))

library(purrr)

tmp.df4_full_stim_downs_jun2021_with_behnew<-tmp.df4_full_stim_downs_jun2021_with_behnew%>%
ungroup()%>%
  mutate(mediansplit_sample_valence2 = 
           if_else(ValenceMeanThisSample >= 
                     median(ValenceMeanThisSample,
                            na.rm = TRUE), 
                   "More positive", "More negative"))%>%
  mutate(mediansplit_sample_arousal2 = if_else(ArousalMeanThisSample >=
                                                  median(ArousalMeanThisSample,
                                                         na.rm = TRUE), "High", "Low"))

tmp.df4_full_stim_downs_jun2021_with_behnew$ssid_split_val_time<- paste0(tmp.df4_full_stim_downs_jun2021_with_behnew$ssid,          
paste0(tmp.df4_full_stim_downs_jun2021_with_behnew$mediansplit_sample_valence2,
       paste0(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin2)))

table(is.na(tmp.df4_full_stim_downs_jun2021_with_beh$mediansplit_self_valence))
unique(tmp.df4_full_stim_downs_jun2021_with_beh$ssid_split_val_time)
subset(tmp.df4_full_stim_downs_jun2021_with_beh, is.na(mediansplit_self_valence))

tmp.df4_full_stim_downs_jun2021_with_behnew$ssid_num<- as.numeric(tmp.df4_full_stim_downs_jun2021_with_behnew$ssid)

tmp.df4_full_stim_downs_jun2021_with_behnew <- tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  subset(ssid_num<500)%>%
  # subset(ssid == 320)%>%
  ungroup()%>%
  group_by(ssid_split_val_time)%>%
  mutate(cor_arous_pup = cor.test(pup_basCor, arousal_c, use = "complete")$estimate,
         cor_arous_pup_pval = cor.test(pup_basCor, arousal_c, use = "complete")$p.value)

tmp.df4_full_stim_downs_jun2021_with_behnew<- tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  group_by(timebin2)%>%
  mutate(mean_cor_pup = mean(cor_arous_pup, na.rm = TRUE), sd_cor_pup = sd(cor_arous_pup, na.rm = TRUE),
         mean_cor_pup_pva = mean(cor_arous_pup_pval, na.rm = TRUE), sd_cor_arous_pup_pval = sd(cor_arous_pup_pval, na.rm = TRUE))

tmp.df4_full_stim_downs_jun2021_with_behnew$pval_sign<- if_else(tmp.df4_full_stim_downs_jun2021_with_behnew$mean_cor_pup_pva < .05, tmp.df4_full_stim_downs_jun2021_with_behnew$mean_cor_pup_pva, NULL)

ttestcheck<- t.test(tmp.df4_full_stim_downs_jun2021_with_behnew$cor_arous_pup)
ttestcheck$p.value
ttestcheck$parameter

unique(tmp.df4_full_stim_downs_jun2021_with_behnew$ssid_split_val_time)

tmp.df4_full_stim_downs_jun2021_with_behnew$vale_timebin<- paste0(tmp.df4_full_stim_downs_jun2021_with_behnew$mediansplit_sample_valence2, paste0(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin2))

# subset(tmp.df4_full_stim_downs_jun2021_with_behnew, vale_timebin == "More negative(2.6,2.82]")

tmp.df4_full_stim_downs_jun2021_with_behnew$ssid<- as.character(tmp.df4_full_stim_downs_jun2021_with_behnew$ssid)

tmp.df4_full_stim_downs_jun2021_with_behnew %>%
    group_by(ssid, mediansplit_sample_valence2, timebin2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  group_by(mediansplit_sample_valence2,timebin2)%>%
  mutate(pval = t.test(cor_arous_pup)$p.value, df = t.test(cor_arous_pup)$parameter)%>%
  group_by(mediansplit_sample_valence2,timebin2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(timerezero3, cor_arous_pup, color = mediansplit_sample_valence2))+
  # geom_s
  stat_summary(fun = mean, geom = 'line', size = 4)+
  # stat_summary(aes(y = pval), fun = mean, geom = 'line', linetype = "dashed", alpha = .2)+
  geom_hline(yintercept = .05, linetype = "dashed")+

  geom_hline(yintercept = .1, linetype = "dotted")+
  theme_classic()

nrow(left_join(tmp.df4_full_stim_downs_jun2021_with_behnew, flag_pos_neg))
nrow(tmp.df4_full_stim_downs_jun2021_with_behnew)

tmp.df4_full_stim_downs_jun2021_with_behnew$cor_pos_neg_diff_lab<- NULL
tmp.df4_full_stim_downs_jun2021_with_behnew$cor_pos_neg_diff<-NULL
tmp.df4_full_stim_downs_jun2021_with_behnew<- left_join(tmp.df4_full_stim_downs_jun2021_with_behnew, flag_pos_neg)

tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  subset(!is.na(cor_pos_neg_diff_lab))%>%
  # subset(cor_pos_neg_diff_lab == "pos > neg") %>%
  group_by(ssid, mediansplit_sample_valence2, timebin2,cor_pos_neg_diff_lab) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  group_by(mediansplit_sample_valence2,timebin2) %>%
  mutate(pval = t.test(cor_arous_pup)$p.value, df = t.test(cor_arous_pup)$parameter)%>%
  group_by(mediansplit_sample_valence2,timebin2,cor_pos_neg_diff_lab)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  # ggplot(aes(timerezero3, cor_arous_pup, color = mediansplit_sample_valence2))+
  ggplot(aes(timerezero3, cor_arous_pup, color =  cor_pos_neg_diff_lab))+
  # geom_s
  stat_summary(fun = mean, geom = 'line', size = 4)+
  # stat_summary(aes(y = pval), fun = mean, geom = 'line', linetype = "dashed", alpha = .2)+
  geom_hline(yintercept = .05, linetype = "dashed")+

  geom_hline(yintercept = .1, linetype = "dotted")+
  geom_vline(xintercept = 1, linetype = "dotted")+
  # geom_rect(xmin = .5, xmax = 1, ymin = -.2, ymax = 1, alpha = .01, color = FALSE)+
  theme_classic()+
    facet_grid(~mediansplit_sample_valence2)

range(tmp.df4_full_stim_downs_jun2021_with_behnew$pval_sign, na.rm = TRUE)
tmp.df4_full_stim_downs_jun2021_with_behnew$ssid_split_val_time

# let's identify the people whio show the negative arousal effect

```

timecourse analyss using lmer

1- select a bin
2 - fit a lmer
3 - save statistics (estimate and p value)
3 - create clusters (p value < .05 in at least 3 consecutive bins)

```{r}

tmp.df4_full_stim_downs_jun2021_with_behnew$
i = 1

nbins<- unique(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin2)

tmp.df4_full_stim_downs_jun2021_with_behnew

unique(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin2)
lmer_bin_pup<- list()

# <- nrow(nbins)

nbins<- unique(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin2)

tmp.df4_full_stim_downs_jun2021_with_behnew$Label<- substr(tmp.df4_full_stim_downs_jun2021_with_behnew$stimIAPS, 1,8)

imageJ_IAPS$Label
nrow(tmp.df4_full_stim_downs_jun2021_with_behnew)
# 62921
nrow(left_join(tmp.df4_full_stim_downs_jun2021_with_behnew, imageJ_IAPS))

tmp.df4_full_stim_downs_jun2021_with_behnew<- left_join(tmp.df4_full_stim_downs_jun2021_with_behnew, imageJ_IAPS)

tmp.df4_full_stim_downs_jun2021_with_behnew$Mean_gray_z<- scale(tmp.df4_full_stim_downs_jun2021_with_behnew$Mean, scale = TRUE, center = TRUE)[,1]

test<- lmer(pup_basCor ~ arousal_c+Mean_gray_z  +(1 +Mean_gray_z | ssid), 
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    timebin2 == nbins[b]))

summary(test)

tmp.df4_full_stim_downs_jun2021_with_behnew$tNo
"(0.0411,0.261]"
subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    timebin2 == nbins[28] & ssid == 341 & tNo == 3)

test
rm(i)

i = 1 # remember to always rezero it
# b = 34
nbins<- nbins[1:28]
unique(nbins)

# nbins<- as

# positive

# create number s for time bins
tmp.df4_full_stim_downs_jun2021_with_behnew<- tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  group_by(ssid, stimIAPS)%>%
  mutate(timebin_no = 1:n())

tmp.df4_full_stim_downs_jun2021_with_behnew%>%
  # group_by(ssid, stimIAPS)
  ggplot(aes(timebin_no))+
  geom_histogram()+
  geom_vline(xintercept = 28)+
  geom_hline(yintercept = 200)

```

```{r}

nrow(nbins)
nbins = 30
timecourse_result_df <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

View(timecourse_result_df)

nbins<- nbins[1:28]

i = 1 # remember to always rezero it
# b = 34

nbins<- unique(tmp.df4_full_stim_downs_jun2021_with_behnew$timebin_no)

unique(nbins)
for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ arousal_c+Mean_gray_z  +(1 +Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_self_valence == "More positive" & 
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

# timecourse_result_df_pos<- timecourse_result_df

timecourse_result_df%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

```

# negative lmer
```{r}

timecourse_result_df_neg <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1

for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ arousal_c+Mean_gray_z  +(1 +Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_ground_valence == "More negative" & 
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_neg[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_neg[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_neg[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_neg[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_neg[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

View(timecourse_result_df_neg)

# whether I use ground valence, self valemce or this sample average valence, I get the same patterns

timecourse_result_df_neg%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .1, color = 'red', size = 1.5)

```

```{r}
timecourse_result_df$Valence <- "More negative"

timecourse_result_df_neg$Valence<- "More positive"

bind_rows(timecourse_result_df,timecourse_result_df_neg)%>%
    ggplot(aes(timebins, t, color = Valence))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

```
# use valence to pupil
```{r}
timecourse_result_df_val <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1

for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ arousal_c+ valence_c+ Mean_gray_z  +(1 + Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    # mediansplit_self_valence == "More negative" & 
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_val[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_val[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_val[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_val[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_val[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

# valence to pupil
timecourse_result_df_val%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

```

# use valence to pupil

```{r}
timecourse_result_df_val_and_ar <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1

for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ valence_c+ arousal+Mean_gray_z  +(1 + Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    # mediansplit_self_valence == "More negative" & 
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_val_and_ar[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_val_and_ar[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_val_and_ar[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_val_and_ar[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_val_and_ar[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

timecourse_result_df_val_and_ar%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

```

valence for high vs low
```{r}
timecourse_result_df_val_and_ar_neg <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1

for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ valence_c + Mean_gray_z  +(1 + Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_sample_arousal2 == "High" &
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_val_and_ar_neg[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_val_and_ar_neg[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_val_and_ar_neg[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_val_and_ar_neg[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_val_and_ar_neg[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

View(timecourse_result_df_val_and_ar_neg)

timecourse_result_df_val_and_ar_neg%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)+
  ggtitle("pupil ~ valence for high arousal")

```

#

```{r}
timecourse_result_df_val_and_ar_low <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1

for (b in 1:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_basCor ~ valence_c +Mean_gray_z  +(1 + Mean_gray_z | ssid) +
                                (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_sample_arousal2 == "Low" &
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_val_and_ar_low[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_val_and_ar_low[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_val_and_ar_low[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_val_and_ar_low[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_val_and_ar_low[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

View(timecourse_result_df_val_and_ar_neg)

timecourse_result_df_val_and_ar_low%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

```

velocity analyses
```{r}
# thought maybe it has soemthing to to with the velcity, velocity is sidderent

install.packages("pspline")
library(pspline)

predict(sm.spline(tmp.df4_full_stim_downs_jun2021_with_beh$timerezero3, 
                  tmp.df4_full_stim_downs_jun2021_with_beh$pup_basCor), 
        tmp.df4_full_stim_downs_jun2021_with_beh$timerezero3, 1)

D(tmp.df4_full_stim_downs_jun2021_with_beh$pup_basCor, 'x')

tmp.df4_full_stim_downs_jun2021_with_behnew

tmp.df4_full_stim_downs_jun2021_with_behnew <- tmp.df4_full_stim_downs_jun2021_with_behnew %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  # subset(valencearousal_outliers == TRUE) %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  group_by(ssid, stimIAPS) %>%
  mutate(pup_velocity = (abs(lag(pup_basCor)-pup_basCor)/abs(lag(timerezero3)-timerezero3)))
  # mutate(pup_velocity_peak = max(pup_velocity, na.rm = TRUE))

tmp.df4_full_stim_downs_jun2021_with_behnew$pup_velocity

# global
timecourse_result_df_vel <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1
b

tmp.df4_full_stim_downs_jun2021_with_behnew$pup_velocity

lmer_bin_pup  <- lmer(pup_velocity ~ arousal_c + Mean_gray_z  + (1 + Mean_gray_z | ssid),
                      REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    timebin_no == nbins[2]))

for (b in 2:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_velocity ~ arousal_c + Mean_gray_z  + (1 + Mean_gray_z | ssid),
                                # (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    # !is.na== "More negative" &
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_vel[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_vel[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_vel[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_vel[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_vel[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

View(timecourse_result_df_vel)

# whether I use ground valence, self valemce or this sample average valence, I get the same patterns

timecourse_result_df_vel%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

# arousal rating is related to velocity around the recovery time 1:2is, and later around 4

```

Velocity by valence
```{r}

timecourse_result_df_vel_pos <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1
# b

tmp.df4_full_stim_downs_jun2021_with_behnew$pup_velocity

for (b in 2:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_velocity ~ arousal_c + Mean_gray_z  + (1  | ssid),
                                # (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_sample_valence2== "More positive" &
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_vel_pos[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_vel_pos[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_vel_pos[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_vel_pos[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_vel_pos[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

# 11 singularity cases

View(timecourse_result_df_vel_pos)

# whether I use ground valence, self valemce or this sample average valence, I get the same patterns

timecourse_result_df_vel_pos%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

# arousal rating is related to velocity around the recovery time 1 to 1.6

```

Negative velocity

```{r}

timecourse_result_df_vel_neg <- data.frame(timebins= rep(NA, 30), Estimate= rep(NA, 30), 
                                   t=rep(NA, 30), p=rep(NA, 30), cov= as.character(rep(NA,30)))

i = 1
b

tmp.df4_full_stim_downs_jun2021_with_behnew$pup_velocity

for (b in 2:length(nbins)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer %i", nbins[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lmer_bin_pup  <- lmer(pup_velocity ~ arousal_c + Mean_gray_z  + (1 | ssid),
                                # (1|stimIAPS),
                              REML = FALSE,
                      data = subset(tmp.df4_full_stim_downs_jun2021_with_behnew,
                                    mediansplit_sample_valence2== "More negative" &
                                    timebin_no == nbins[b]))

        #store results from the simulation
            lmer_bin_pup_summary<- summary(lmer_bin_pup)
            timecourse_result_df_vel_neg[i,1]<- tmp.df4_full_stim_downs_jun2021_with_behnew$timerezero3[b] #save the exact value of time bin
            timecourse_result_df_vel_neg[i,2]<- lmer_bin_pup_summary[["coefficients"]][2,1] # s
            timecourse_result_df_vel_neg[i,3]<-lmer_bin_pup_summary[["coefficients"]][2,4] # t statistic
            timecourse_result_df_vel_neg[i,4]<-lmer_bin_pup_summary[["coefficients"]][2,5] # p value
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            timecourse_result_df_vel_neg[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
                                                   != 0,
                                               lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

# 11 singularity cases

View(timecourse_result_df_vel_neg)

# whether I use ground valence, self valemce or this sample average valence, I get the same patterns

timecourse_result_df_vel_neg%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .05, color = 'red', size = 1.5)

# arousal rating is related to velocity around the recovery time 1 to 1.6

```

A few plots for velocity
```{r}
  ggplot(aes(x = timerezero3, y = pup_velocity))+
  stat_smooth(aes(group= mediansplit_sample_valence2, color = mediansplit_sample_valence2), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=mediansplit_sample_valence2, color = mediansplit_sample_valence2), size = 3)+  
theme_classic()+
  ylab("Pupil velocity")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
   # facet_grid(~mediansplit_sample_valence2)+
     geom_vline(xintercept = 1)

tmp.df4_full_stim_downs_jun2021_with_beh$timerezero3
tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  subset(timerezero3>1 ) %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  group_by(ssid, stimIAPS)%>%
  mutate(pup_velocity = (abs(lag(pup_basCor)-pup_basCor))/(abs(lag(timerezero3)-timerezero3)))%>%
  mutate(pup_velocity_peak = max(pup_velocity, na.rm = TRUE))%>%
  group_by(ssid,mediansplit_sample_valence2,mediansplit_sample_arousal2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(x = mediansplit_sample_valence2, y = pup_velocity))+
  # geom_jitter(width = .1, alpha = .1)+
  stat_summary(geom = 'pointrange')+
  theme_classic()+
  ggpubr::stat_compare_means()+
  ylab("Pupil velocity (z)")+
  xlab("Time (s)")+
  facet_grid(~mediansplit_sample_arousal2)

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(mediansplit_sample_arousal2))%>%
  subset(timerezero3>1) %>%
  # subset(!is.na(mediansplit_sample_arousal2))%>%
  group_by(ssid, stimIAPS)%>%
  mutate(pup_velocity = (abs(lag(pup_basCor)-pup_basCor))/(abs(lag(timerezero3)-timerezero3)))%>%
  mutate(pup_velocity_peak = max(pup_velocity, na.rm = TRUE))%>%
  group_by(ssid,mediansplit_sample_valence2,mediansplit_sample_arousal2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(x = mediansplit_sample_valence2, y = pup_velocity_peak))+
  geom_jitter(width = .1, alpha = .1)+
  stat_summary(geom = 'pointrange')+
  theme_classic()+
  ggpubr::stat_compare_means()+
  ylab("peak Pupil velocity (z)")+
  xlab("Time (s)")+
  facet_grid(~mediansplit_sample_arousal2)

```
```{r}
tmp.df4_full_stim_downs_jun2021_with_behnew%>%
    group_by(mediansplit_self_valence, timebin2)%>%
summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(timerezero3, cor_arous_pup, color = mediansplit_self_valence))+
  # geom_s
  stat_summary(fun = mean, geom = 'pointrange')+
  stat_summary(aes(y = cor_arous_pup_pval), fun = mean, geom = 'line', linetype = "dashed")+
  geom_hline(yintercept = .1, linetype = "dashed")
  geom_rect(aes(xmin = timerezero3, xmax = dplyr::lead(timerezero3), ymin =.63 , ymax =.66 , fill = pval_sign), 
            alpha = 0.8)

  # stat_smooth(aes(group = ssid), se = F, alpha = .1)
  # mutate(
  #   model = map(data, ~lm(pup_basCor ~ arousal_c, data = .)),
  #   cor = map(data, possibly(
  #     ~tidy(cor.test(.x$pup_basCor, .x$arousal_c), 3), otherwise = data.frame())
  #   )
  # )

```

# self arousal

```{r}

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(ground_arousal_lab))%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= ground_arousal_lab, color = ground_arousal_lab), fun = mean,geom = "line",
               se = F, alpha = .1, size = 1.5)+
  stat_smooth(aes(group=ground_arousal_lab, color = ground_arousal_lab), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)+
  facet_grid(~mediansplit_self_valence)

```

```{r}

# plot with valence
tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(ground_arousal_lab))%>%
  group_by(ssid,Alexithymia, mediansplit_ground_valence, timebin2)%>%
  ggplot(aes(x = timerezero3, y = pup_basCor))+
  stat_smooth(aes(group= mediansplit_ground_valence, color = mediansplit_ground_valence), fun = mean,geom = "line",
               se = F, size = 1.5)+
  # stat_smooth(aes(group=ground_valence_lab, color = ground_valence_lab), size = 3)+  
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  xlim(0,6)

test<- cor.test(db_2021_pupil_summ$hr_interp,db_2021_pupil_summ$pup_basCor)

test$p.value
tmp.df4_full_stim_downs_jun2021_with_beh$ssid_ground_valence<- paste0(tmp.df4_full_stim_downs_jun2021_with_beh$ssid, paste0(tmp.df4_full_stim_downs_jun2021_with_beh$ground_arousal_lab))

tmp.df4_full_stim_downs_jun2021_with_beh %>%
  subset(!is.na(self_arousal_lab))%>%

  db_full4new_stim_screen_pupil_nopract

  group_by(ssid,Alexithymia, ground_valence_lab, timebin2)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor)$estimate)%>%
  mutate(pcortest = cor.test(arousal_c, pup_basCor)$p.value)%>%
  group_by(ssid,Alexithymia, self_arousal_lab, timebin2)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ggplot(aes(x = timerezero3, y = cortest))+

  stat_smooth(aes(group= self_arousal_lab, color = self_arousal_lab), fun = mean,geom = "line",
               se = F, size = 1.5)+
theme_classic()+
  ylab("Pupil size (z)")+
  xlab("Time (s)")+
  # p$graphstyle+
  scale_color_brewer(palette = "Dark2")+
  geom_hline(yintercept = .05, linetype = "dashed")+
  xlim(0,6)

# does arousing stimuli impairs appraisal

 tmp.df4_full_stim_downs_jun2021_with_beh %>%
   ggplot(aes(ground_arousal_lab, BRIGHTNESS))+
   geom_boxplot()+
   geom_point()+
   # stat_summary(geom = 'boxplot')+
   facet_grid(~mediansplit_self_valence)
```

```{r}
 db_full4new_stim_screen_pupil_nopract %>%
  subset(pupil_outlier == FALSE & arousal_outler == FALSE & !is.na(Alexithymia) & ssid!= 610)%>%
    # group_by(ssid,Alexithymia, is_valence_high)%>%
   group_by(ssid,Alexithymia, mediansplit_self_valence)%>%
    mutate(cortest = cor.test(arousal_c, pup_basCor)$estimate)%>%
   group_by(ssid,Alexithymia, mediansplit_self_valence, self_arousal_lab)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
        ggplot(aes(mediansplit_self_valence, cortest, color = mediansplit_self_valence))+
           geom_bar(stat="summary", fun.y = "mean")+

    geom_jitter(width = .1, alpha = .1)+
        stat_summary( geom = 'pointrange')+
          # stat_summary(aes(group = ssid), geom = 'line')+
    ylab("R pupil vs self report arousal")+
   # ggpubr::stat_compare_means(paired = TRUE)+
  # ggpubr::stat_compare_means(comparisons = my_comparisons, paired = TRUE)
      facet_grid(~self_arousal_lab)

```

SCR

```{r}

unique(db_full4new_stim_screen_pupil_nopract$ssid)
View(flag_pos_neg)
db_full4new_stim_screen_pupil_nopract$mediansplit_self_arousal <- if_else(db_full4new_stim_screen_pupil_nopract$arousal> median(db_full4new_stim_screen_pupil_nopract$arousal, na.rm = TRUE), "High", "Low")

db_full4new_stim_screen_pupil_nopract$mediansplit_self_arousal_thissample <- if_else(db_full4new_stim_screen_pupil_nopract$arousal> median(db_full4new_stim_screen_pupil_nopract$ArousalMeanThisSample, na.rm = TRUE), "High", "Low")

unique(db_full4new_stim_screen_pupil_nopract2$ssid)

db_full4new_stim_screen_pupil_nopract_nt<- subset(db_full4new_stim_screen_pupil_nopract, ssid_num<500)

nrow(db_full4new_stim_screen_pupil_nopract_nt)
# 2233
# nrow(left_join(db_full4new_stim_screen_pupil_nopract_nt, flag_pos_neg))

db_full4new_stim_screen_pupil_nopract_nt<- left_join(db_full4new_stim_screen_pupil_nopract_nt, flag_pos_neg)

unique(db_full4new_stim_screen_pupil_nopract_nt$stimDescription)
db_full4new_stim_screen_pupil_nopract_nt %>%
  subset(ssid_num< 500)%>%
    subset(pp_arousal_outler == FALSE)%>%
  subset( arousal_outler == FALSE )%>%
       subset(!is.na(Alexithymia))%>%
   group_by(ssid)%>%
  mutate(BIO_CDA.SCR_z = scale((BIO_CDA.SCR))[,1])%>%
  ungroup()%>%
    group_by(ssid,cor_pos_neg_diff_lab, mediansplit_self_valence)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_self_valence, BIO_CDA.SCR_z))+
      stat_summary(geom = 'pointrange')+

     # geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  geom_jitter(width = .2, alpha = .2)+
  stat_summary(aes(group = ssid), geom = 'line', color = "gray", alpha = .1)+
  # geom_boxplot(alpha = .1)+
  # facet_grid(~cor_pos_neg_diff_lab)+
  ggpubr::stat_compare_means()

BIO_CDA.ISCR_z

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid)%>%
  mutate(BIO_TTP.AmpSum_z = scale(BIO_TTP.AmpSum, center = TRUE, scale = TRUE)[,1])%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, BIO_TTP.AmpSum_z))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid,mediansplit_sample_valence2)%>%
  # mutate(BIO_TTP.AmpSum_z = scale(BIO_TTP.AmpSum, center = TRUE, scale = TRUE)[,1])%>%
  # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal, BIO_CDA.ISCR_z))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  facet_grid(~mediansplit_sample_valence2)

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid,mediansplit_sample_valence2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, BIO_CDA.PhasicMax_z, color =mediansplit_sample_valence2))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid,ground_arousal_lab)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_c, BIO_CDA.PhasicMax_z, color =ground_arousal_lab))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

```

# scr model

<!-- # average aroual and rating over all trials don't track well the expected arousal relation -->
```{r}

unique(db_full4new_stim_screen_pupil_nopract_nt$condition)
pupil_arousal_findings$arousal_scr <- lmer(log(BIO_CDA.SCR+.1) ~ arousal_c*TASc +(1|ssid) + (1|stimIAPS),
                        REML = FALSE,
                        data = subset(db_full4new_stim_screen_pupil_nopract_nt,
                                      pp_arousal_outler == FALSE &
                                      mediansplit_self_valence == "More positive"))

summary(pupil_arousal_findings$arousal_scr)
anova(pupil_arousal_findings$arousal_phas_max)

BIO_CDA.PhasicMax

pupil_arousal_findings$arousal_scr <- lmer(BIO_CDA.PhasicMax_z ~ valence_c +(1+arousal_c|ssid),
                        REML = FALSE,
                        data = subset(db_full4new_stim_screen_pupil_nopract_nt,
                                      pp_arousal_outler == FALSE))
                                      # mediansplit_self_valence == "More positive"))

summary(pupil_arousal_findings$arousal_scr)
anova(pupil_arousal_findings$arousal_phas_max)

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid, cor_pos_neg_diff_lab)%>%
  ggplot(aes(cor_pos_neg_diff_lab, TAS))+
  geom_point()+
  stat_summary(geom = 'pointrange')
```

summary(pupil_arousal_findings$arousal_scr)
anova(pupil_arousal_findings$arousal_phas_max)

test<- subset(db_full4new_stim_screen_pupil_nopract, pp_arousal_outler == FALSE & ssid == 310)
test$Alexithymia
table(test$mediansplit_self_valence)

# try the anova grouping
```{r}

afex::aov_ez(data = (db_full4new_stim_screen_pupil_nopract %>%
    subset(pp_arousal_outler == FALSE)%>%
      # subset(is_scr_nonresponder == FALSE )%>%
      subset(!is.na(Alexithymia))%>%
      group_by(ssid, mediansplit_ground_valence)%>%
      # group_by(ssid, Alexithymia,mediansplit_self_arousal, mediansplit_self_valence)%>%
      summarise_if(is.numeric, mean, na.rm = TRUE)),
    id = "ssid",
    between = NULL,
    within = c("mediansplit_ground_valence"),
    dv = "BIO_CDA.SCR_z")

colnames(db_full4new_stim_screen_pupil_nopract)

ggcorrplot::ggcorrplot(cor(db_full4new_stim_screen_pupil_nopract[,c(19,115,125:128,27,26)], use = "complete"))

```

BIO_CDA.SCR_z

# indovidual plots
```{r}

facet_hist_scr<- db_full4new_stim_screen_pupil_nopract %>%
  subset(ssid > 303)%>%
  subset(arousal_outler == FALSE )%>%
       subset(!is.na(Alexithymia))%>%
  ggplot(aes(log1p(BIO_CDA.SCR+1)))+
  geom_histogram(bins = )+
  facet_wrap(~ssid)

```

# what's up with non responders

# 312
# 318 only one rspinse?
# 320
# 321
# 327
# 329
# 332
# 335 (only one response)
# 342
# 343
# 600
# 603
# 606 (only one response)
# 608
# 900
# 56- 15

# that would leave us with 41 people

```{r}
unique(db_full4new_stim_screen_pupil_nopract$ssid)

range(db_full4new_stim_screen_pupil_nopract$BIO_CDA.SCR, na.rm = TRUE)

mean(db_full4new_stim_screen_pupil_nopract$BIO_CDA.SCR, na.rm = TRUE)
sd(db_full4new_stim_screen_pupil_nopract$BIO_CDA.SCR, na.rm = TRUE)

db_full4new_stim_screen_pupil_nopract%>%
  ggplot(aes(BIO_CDA.SCR))+
  geom_histogram(bins = 500)
  xlim(0,.01)

db_full4new_stim_screen_pupil_nopract%>%
  subset(ssid == 312)%>%
  ggplot(aes(BIO_CDA.AmpSum))+
  geom_histogram(bins = 500)+
  xlim(0,.05)

```

```{r}
db_full4new_stim_screen_pupil_nopract$is_scr_nonresponder<- if_else(db_full4new_stim_screen_pupil_nopract$BIO_CDA.AmpSum< .1, TRUE, FALSE)

 db_full4new_stim_screen_pupil_nopract %>%
  subset(ssid > 303)%>%
  subset(arousal_outler == FALSE )%>%
       subset(!is.na(Alexithymia))%>%
  ggplot(aes(log1p(BIO_CDA.SCR+1)))+
  geom_histogram(bins = )+
   facet_grid(~is_scr_nonresponder)
  facet_wrap(~ssid)

  db_full4new_stim_screen_pupil_nopract%>%
    group_by(ssid,is_scr_nonresponder)%>%
    mutate(scr_nonresp_prop = n())%>%
      group_by(ssid)%>%
     mutate(scr_nonresp_prop = scr_nonresp_prop/n())%>%
    ggplot(aes(as.factor(ssid), scr_nonresp_prop))+
    geom_bar(stat="summary", fun.y = "mean", alpha = .2, width = .5)

# test<-  subset(db_full4new_stim_screen_pupil_nopract, ssid == 321)

  table(db_full4new_stim_screen_pupil_nopract_nt$is_scr_nonresponder)
  db_full4new_stim_screen_pupil_nopract_nt %>%
    subset(is_scr_nonresponder != FALSE)%>%
    subset(pp_arousal_outler == FALSE)%>%
    subset(!ssid %in% c("304","321", "323", "318","339","331","325"))%>%
  # subset( arousal_outler == FALSE )%>%
       subset(!is.na(Alexithymia))%>%
   group_by(ssid)%>%
  mutate(BIO_CDA.SCR_z = scale(log1p(BIO_CDA.PhasicMax +.1))[,1])%>%
  # ungroup()%>%
    group_by(ssid, Alexithymia,mediansplit_sample_arousal2, mediansplit_sample_valence2)%>%
    summarise_if(is.numeric, mean, na.rm = TRUE)%>%
    # group_by(ssid)%>%
    # mutate(SCR_trend = BIO_CDA.SCR_z[mediansplit_ground_arousal == "Higher arousal"] -
             # BIO_CDA.SCR_z[mediansplit_ground_arousal == "Lower arousal"])
    # mutate(trendscr = if_else(SCR_trend> 0, "high to high", "high to low"))

    # group_by(ssid, Alexithymia,mediansplit_ground_arousal,trendscr)%>%
    # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_sample_arousal2, BIO_CDA.SCR_z))+
      stat_summary(geom = 'pointrange')+
     stat_summary(aes(group = ssid), geom = 'line')+

     # geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  geom_jitter(width = .2, alpha = .2)
    facet_grid(~mediansplit_sample_valence2)
    # geom_boxplot(alpha = .1)
    # geom_text(aes(label = ssid))
    facet_grid(~trendscr)
    # ylim(-1,1)

304, 318, 339, 331,325

scr_models<- list()

db_full4new_stim_screen_pupil_nopract_nt %>%
    subset(is_scr_nonresponder == FALSE)%>%
    subset(pp_arousal_outler == FALSE)%>%
    subset(!ssid %in% c("304","321", "323", "318","339","331","325"))
  # subset( arousal_outler == FALSE )%>%

db_full4new_stim_screen_pupil_nopract_nt_scr_excluding_atypical<- db_full4new_stim_screen_pupil_nopract_nt%>%
                        subset(!ssid %in% c("304","321", "323", "318","339","331","325"))
  # subset(is_scr_nonresponder == FALSE)%>%
  group_by(ssid)%>%
  mutate(BIO_CDA.PhasicMax_z = scale(BIO_CDA.PhasicMax, center = TRUE, scale = TRUE)[,1])%>%
 mutate(valence_z = scale(valence, center = TRUE, scale = TRUE)[,1])%>%
    mutate(arousal_z = scale(arousal, center = TRUE, scale = TRUE)[,1])

scr_models$scr_excluding_atypical<- lmer(log(BIO_CDA.PhasicMax+.1) ~ valence_z + arousal_z+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data = db_full4new_stim_screen_pupil_nopract_nt_scr_excluding_atypical)

scr_models$scr_excluding_atypical<- lmer(log(BIO_CDA.PhasicMax+.1) ~ valence_c + arousal_c+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  db_full4new_stim_screen_pupil_nopract_nt)
                        # subset(!ssid %in% c("304","321", "323", "318","339","331","325"))%>%
  subset(is_scr_nonresponder != FALSE))

summary(scr_models$scr_excluding_atypical)

(db_full4new_stim_screen_pupil_nopract_nt%>%
                        subset(!ssid %in% c("304","321", "323", "318","339","331","325"))))

summary(scr_models$scr_excluding_atypical)
```
<!-- one of the worries I had was that maybe there was a bias such that valence was confounded with arousal, but if this was true, then positive stimuli woul;d be rated as more arousing -->

    db_full4new_stim_screen_pupil_nopract %>%
    subset(is_scr_nonresponder == FALSE)%>%
    subset(pp_arousal_outler == FALSE)%>%
  subset( arousal_outler == FALSE )%>%
       subset(!is.na(Alexithymia))%>%
   group_by(ssid)%>%
  mutate(BIO_CDA.SCR_z = scale(log1p(BIO_CDA.PhasicMax +.1))[,1])%>%
      mutate(arousal_z = scale(arousal +.1))%>%
  ungroup()%>%
    # group_by(ssid, Alexithymia,mediansplit_self_arousal)%>%
    # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal_z, BIO_CDA.SCR_z))+
      geom_point()+
      # geom_smooth(aes(group = ssid),method = 'lm', se = F)+
      geom_smooth(method = 'lm', se = F)+
      facet_grid(~ground_arousal_lab)

     # geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  geom_jitter(width = .2, alpha = .2)+
    ylim(-1,1)

  # panel for many folks
  db_full4new_stim_screen_pupil_nopract %>%
  # subset(ssid == 314)%>%
    group_by(ssid)%>%
    mutate(BIO_CDA.SCR_z = scale(log1p(BIO_CDA.PhasicMax +.1))[,1])%>%
    mutate(arousal_z = scale(arousal +.1))%>%
  ggplot(aes(arousal_z,BIO_CDA.SCR_z, color =mediansplit_ground_valence , shape =mediansplit_ground_arousal))+
    geom_point()+
    geom_smooth(method = 'lm', se = F)+
    facet_wrap(~ssid)+
    ggpubr::stat_cor()

    db_full4new_stim_screen_pupil_nopract %>%
  # subset(ssid == 314)%>%
    group_by(ssid)%>%
    mutate(BIO_CDA.SCR_z = scale(log1p(BIO_CDA.PhasicMax +.1))[,1])%>%
    mutate(arousal_z = scale(arousal +.1))%>%
  ggplot(aes(pup_basCor,BIO_CDA.SCR_z))+
    geom_point()+
    geom_smooth(method = 'lm', se = F)+
    facet_wrap(~ssid)+
    ggpubr::stat_cor()

        db_full %>%
  # subset(ssid == 314)%>%
    group_by(ssid)%>%
    mutate(BIO_Mean_HR = scale(BIO_Mean_HR, scale = FALSE)[,1])%>%
    mutate(arousal_z = scale(arousal +.1))%>%
  ggplot(aes(BIO_Mean_HR,log1p(BIO_CDA.PhasicMax+.1)))+
    geom_point()+
    geom_smooth(aes(group = ssid),method = 'lm', se = F)
    facet_wrap(~ssid)+
    ggpubr::stat_cor()

Let's try some 3 d visualisations
continue here

````{r}

db_full4new_stim_screen_pupil_nopract_nt %>%
  group_by(ssid, mediansplit_ground_arousal, mediansplit_ground_valence,scr_utl)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  subset(scr_utl == "notoutlier")%>%
  ggplot(aes(mediansplit_ground_arousal , BIO_CDA.PhasicMax_z))+
    geom_jitter(alpha = .1, width = .1)+
  stat_summary(geom = 'pointrange')+
  stat_summary(aes(group = ssid), geom = "line", alpha = .1)+
  facet_grid(~mediansplit_ground_valence)

db_full4new_stim_screen_pupil_nopract_nt%>%
  ggplot(aes(BIO_CDA.PhasicMax_z))+
  geom_histogram()

db_full4new_stim_screen_pupil_nopract_nt<- db_full4new_stim_screen_pupil_nopract_nt%>%
  mutate(scr_utl = if_else(BIO_CDA.PhasicMax_z > 5, "outlier", "notoutlier"))

```

  # + arousal_c

```{r}
scr_models$scr_excluding_atypical<- lmer(log(BIO_CDA.AmpSum_collapsed+.1) ~  arousal_c*valence_c * TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract_nt, scr_utl == "notoutlier"))

summary(scr_models$scr_excluding_atypical)
plot(scr_models$scr_excluding_atypical)

interactions::interact_plot(scr_models$scr_excluding_atypical, pred = arousal_c, modx = valence_c, mod2 = TASc)

db_full4new_stim_screen_pupil_nopract_nt%>%
  ggplot(aes())

scr_models$scr1<- lmer(log(BIO_CDA.PhasicMax_collapsed+.1) ~  arousal_c*valence_c * TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract_nt, scr_utl == "notoutlier"))

summary(scr_models$scr1)
scr_models$scr1.2<- lmer(log1p(BIO_CDA.PhasicMax+.1) ~  TASc*arousal_c * valence_c+
                                           (1 | ssid)+
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract, Group == "NT", !is.na(Alexithymia)))

unique(db_full4new_stim_screen_pupil_nopract$ssid)
summary(scr_models$scr1.2)

db_full4new_stim_screen_pupil_nopract %>%
  group_by(ssid,Group)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  subset(Group == "NT")%>%
  ggplot(aes(TAS))+
  geom_histogram()

interactions::interact_plot(scr_models$scr1, pred = arousal_c, modx = valence_c, mod2 = TASc)

scr_models$scr1.1<- lmer(CDA.PhasicMax ~  arousal_c*valence_c * TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract_nt, scr_utl == "notoutlier"))

summary(scr_models$scr1.1)

interactions::interact_plot(scr_models$scr1.1, pred = arousal_c, modx = valence_c, mod2 = TASc)

scr_models$scr2<- lmer(BIO_CDA.ISCR_collapsed ~  arousal_c*valence_c *TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract_nt, scr_utl == "notoutlier"))

summary(scr_models$scr2)

interactions::interact_plot(scr_models$scr2, pred = arousal_c, modx = valence_c)

scr_models$scr2<- lmer(BIO_CDA.SCR_collapsed ~  arousal_c* valence_c *TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract, Group = "NT"))

summary(scr_models$scr2)

interactions::interact_plot(scr_models$scr2, pred = arousal_c, modx = valence_c)

# the effect of pupil is the opposite on scr

scr_models$scr3<- lmer(BIO_CDA.ISCR_collapsed ~  arousal_c * valence_c * TASc+
                                           (1| ssid) +
                                           (1 | stimIAPS),
                        REML = FALSE,
                        data =  subset(db_full4new_stim_screen_pupil_nopract_nt, scr_utl == "notoutlier"))

summary(scr_models$scr3)

interactions::interact_plot(scr_models$scr2, pred = arousal_c, modx = valence_c, mod2 = TASc)

# the effect of pupil is the opposite on scr

```

check scr on within individuals

```{r}

lm_objtest<- lm(BIO_CDA.PhasicMax_z ~ arousal * valence, data = db_full4new_stim_screen_pupil_nopract_nt)
lm_objtest_summ<-summary(lm_objtest)
lm_objtest_summ$coefficients
db_full4new_stim_screen_pupil_nopract_nt

lm_objtest[["coefficients"]][[2]]
lm_objtest[["coefficients"]][["arousal"]]
lm_objtest[["coefficients"]][["valence"]]

regression_df <- data.frame(ssid= rep(NA, 50), Estimate_arousal= rep(NA, 50), 
                         Estimate_valence= rep(NA, 50),
                         Estimate_int= rep(NA, 50),
                         p_arousal= rep(NA, 50), 
                         p_valence= rep(NA, 50),
                         p_int= rep(NA, 50))

lm_ind_summary$coefficients[[14]]

ssid<- unique(db_full4new_stim_screen_pupil_nopract_nt$ssid)

db_full4new_stim_screen_pupil_nopract_nt<- db_full4new_stim_screen_pupil_nopract_nt%>%
  group_by(ssid)%>%
  mutate(arousal_z_ssid = scale(arousal - mean(arousal, na.rm = TRUE))[,1],
         valence_z_ssid = scale(valence - mean(valence, na.rm = TRUE))[,1])

i = 1

rm(b)

# debug
ssid_tes<- unique(db_full4new_stim_screen_pupil_nopract_nt$ssid_num)

table(is.na(db_full4new_stim_screen_pupil_nopract_nt$BIO_CDA.PhasicMax_z))
lm_ind  <- lm(BIO_CDA.PhasicMax_z ~ arousal_z_ssid *valence_z_ssid,
                      data = subset(db_full4new_stim_screen_pupil_nopract_nt, !is.na(BIO_CDA.PhasicMax_z) &
                                    ssid_num == ssid_tes[10]))

subset(db_full4new_stim_screen_pupil_nopract_nt, !is.na(BIO_CDA.PhasicMax_z) &
                                    ssid_num == ssid_tes[10])

```

```{r}
i = 1
rm(b)

for (b in 5:length(ssid_tes)) {
  # for (s in 1:length(nsim)) {

    message(sprintf("$$$$$RUNING lmer for ssid %i", ssid_tes[b]))
    # db_of7_new_JEFFE$VideoType_contrast_sample<- sample((rep(c(-.5, .5), each =  1920)), replace = FALSE) 
        #run the model on the current time bin and simulation sumber
        lm_ind  <- lm(BIO_CDA.PhasicMax_z ~ arousal_z_ssid *valence_z_ssid,
                      data = subset(db_full4new_stim_screen_pupil_nopract_nt, !is.na(BIO_CDA.PhasicMax_z) & 
                                    mediansplit_ground_valence == "More negative"&
                                     ssid_num == ssid_tes[b]))

        #store results from the simulation
            lm_ind_summary<- summary(lm_ind)
            regression_df[i,1]<- ssid_tes[b] #save ssid
            regression_df[i,2]<- lm_ind_summary$coefficients[[2]] # arousal beta
            regression_df[i,3]<-lm_ind_summary$coefficients[[3]] # valence beta
            regression_df[i,4]<-lm_ind_summary$coefficients[[4]] # inyteraction beta
            regression_df[i,4]<-lm_ind_summary$coefficients[[4]] 
            regression_df[i,5]<- lm_ind_summary$coefficients[[14]]
            regression_df[i,6]<- lm_ind_summary$coefficients[[15]]
            regression_df[i,7]<- lm_ind_summary$coefficients[[16]]

            lm_ind_summary$coefficients[[14]]
            # simulated_clusters_JEFFE[i,6]<- nsim[s] #store simulation ount
            # timecourse_result_df_neg[i,5] <-ifelse(length(lmer_bin_pup_summary$optinfo$conv$lme4$message)
            #                                        != 0,
            #                                    lmer_bin_pup_summary$optinfo$conv$lme4$message, 'pass')

     i = i+1
}

View(regression_df)

 lm_ind_summary$coefficients[[4]]
# whether I use ground valence, self valemce or this sample average valence, I get the same patterns

timecourse_result_df_neg%>%
  ggplot(aes(timebins, t))+
  geom_line(size = 2)+
  geom_line(aes(y = p), linetype = "dashed", size = 2)+
  geom_hline(yintercept = .1, color = 'red', size = 1.5)

t.test(regression_df$Estimate_arousal)

ssid# just positive
regression_df <- data.frame(ssid= rep(NA, 50), Estimate_arousal= rep(NA, 50), 
                         Estimate_valence= rep(NA, 50),
                         Estimate_int= rep(NA, 50),
                         p_arousal= rep(NA, 50), 
                         p_valence= rep(NA, 50),
                         p_int= rep(NA, 50))

```

```{r}

scatter3D(x, y, z, clab = c("Sepal", "Width (cm)"))

db_full4new_stim_screen_pupil_nopract_nt

install.packages("plot3D")
install.packages("misc3d")
library(misc)

plot3D::scatter3D(db_full4new_stim_screen_pupil_nopract_nt$pup_basCor, 
          db_full4new_stim_screen_pupil_nopract_nt$arousal_c, 
          db_full4new_stim_screen_pupil_nopract_nt$valence_c)

 install.packages("plot3D", dependencies = TRUE)

  install.packages("misc3d", repo = 'https://mac.R-project.org')

install.packages(c("rgl", "car"))

library(car)

install.packages("car", dependencies = TRUE)

car::scatter3d(db_full4new_stim_screen_pupil_nopract_nt$arousal_c, 
          db_full4new_stim_screen_pupil_nopract_nt$pup_basCor, 
          db_full4new_stim_screen_pupil_nopract_nt$valence_c,
          fit = "smooth"
          # point = FALSE,

          # surface = FALSE,
          # point.col = ""
          )

```

gg = ggplot(db_full4new_stim_screen_pupil_nopract_nt, aes(BIO_CDA.PhasicMax_z, arousal_c)) +
  stat_density_2d(aes(fill = stat(db_full4new_stim_screen_pupil_nopract_nt$valence_c)), 
                  geom = "polygon",
                  n = 100,bins = 10,contour = TRUE) +
  # facet_wrap(clarity~.) +
  scale_fill_viridis_c(option = "A")

rayshader::plot_gg(gg,multicore=TRUE,width=5,height=5,scale=250)

scr<- db_full4new_stim_screen_pupil_nopract_nt$BIO_CDA.PhasicMax_z
arousal<- db_full4new_stim_screen_pupil_nopract_nt$arousal_c

valence = db_full4new_stim_screen_pupil_nopract_nt$valence_c

gg = ggplot(db_full4new_stim_screen_pupil_nopract_nt, aes(BIO_CDA.PhasicMax_z, arousal_c)) +
  stat_density_2d(aes(fill = stat(valence_c)), 
                  geom = "polygon",
                  # n = 100,bins = 10,
                  contour = TRUE) +
  # facet_wrap(clarity~.) +
  scale_fill_viridis_c(option = "A")
rayshader::plot_gg(gg,multicore=TRUE,width=5,height=5,scale=250)
gg
#No lines
pp_nolines = ggplot(db_full4new_stim_screen_pupil_nopract_nt, aes(x=BIO_CDA.PhasicMax_z, y=arousal_c)) +
  geom_point(aes(color = valence_c))
  scale_fill_viridis_c(a)

rayshader::plot_gg(pp_nolines, width = 4, height = 4, scale = 300, multicore = TRUE)
pp_nolines

?scatter3d
scatter3d(x = arousal, 
          y = valence, z = scr, 
          fit = "smooth",

          # surface = FALSE,
          point.col = "blue", 
          # point
          groups = as.factor(db_full4new_stim_screen_pupil_nopract_nt$mediansplit_sample_valence2), 
          ellipsoid = FALSE)

scatter3d(x = db_full4new_stim_screen_pupil_nopract_nt$pup_basCor, 
          y = db_full4new_stim_screen_pupil_nopract_nt$BIO_CDA.PhasicMax_z, z = db_full4new_stim_screen_pupil_nopract_nt, 
          fit = "smooth",

          # surface = FALSE,
          point.col = "blue", 
          # point
          groups = as.factor(db_full4new_stim_screen_pupil_nopract_nt$mediansplit_ground_valence), 
          ellipsoid = FALSE)

library(viridisLite)
install.packages("viridis")
library(viridis)
deathgg = ggplot(db_full4new_stim_screen_pupil_nopract_nt) +
  geom_tile(aes(x=arousal_c,y=pup_basCor,fill=valence_c))+
  # scale_x_continuous("Year",expand=c(0,0),breaks=seq(1900,2010,10)) +
  # scale_y_continuous("Age",expand=c(0,0),breaks=seq(0,100,10),limits=c(0,100)) +
  scale_fill_viridis()
  # ggtitle("Death Probability vs Age and Year for the USA") +
  # labs(caption = "Data Source: US Dept. of Social Security")

rayshader:: plot_gg(deathgg, multicore=TRUE)
height=5,width=6,scale=500

```

heart rate

```{r}

unique(db_full6new$Bio_Mean_HR)
unique(db_full6new$Bio_Mean_HR_fix)

View(db_full6new)

db_full4new_stim_screen_pupil_nopract
  db_full4new_stim_screen_pupil_nopract %>%
  # subset(ssid == 314)%>%
    group_by(ssid)%>%
    mutate(BIO_CDA.SCR_z = scale(log1p(BIO_CDA.PhasicMax +.1))[,1])%>%
    mutate(arousal_z = scale(arousal +.1))%>%
  ggplot(aes(arousal_z,BIO_CDA.SCR_z, color =mediansplit_ground_valence , shape =mediansplit_ground_arousal))+
    geom_point()+
    geom_smooth(method = 'lm', se = F)+
    facet_wrap(~ssid)+
    ggpubr::stat_cor()

```

```

```{r}

# keep stim and and fixation

unique(db_full6new$screencontent)
unique(db_full6new$screencontent_no)
db_full6new_hr_fix_stim<- subset(db_full6new, screencontent_no <3)
View(db_full6new_hr_fix_stim)
# fixation"    "stim"

# keep controls
unique(db_full6new_hr_fix_stim$ssid)
db_full6new_hr_fix_stim$ssid_num <- as.numeric(as.character(db_full6new_hr_fix_stim$ssid))

db_full6new_hr_fix_stim_nt<- subset(db_full6new_hr_fix_stim, Group == "NT")
unique(db_full6new_hr_fix_stim_nt$ssid_num)

db_full6new_hr_fix_stim_nt<- subset(db_full6new_hr_fix_stim, ssid_num> 303)
unique(db_full6new_hr_fix_stim_nt$ssid)

# HR differences
db_full6new_hr_fix_stim_nt$Bio_Mean_HR_fix<- if_else(db_full6new_hr_fix_stim_nt$screencontent_no ==1, db_full6new_hr_fix_stim_nt$Bio_Mean_HR, NULL)

View(db_full6new_hr_fix_stim_nt)

library(dplyr)
library(tidyverse)
db_full6new_hr_fix_stim_nt1<- db_full6new_hr_fix_stim_nt %>%
  group_by(ssid, tNo) %>%
  arrange(ssid, tNo,screencontent_no)%>%
  fill(Bio_Mean_HR_fix, .direction = "down")%>%
  subset(screencontent_no == 2)

View(db_full6new_hr_fix_stim_nt1)

db_full6new_hr_fix_stim_nt1<- db_full6new_hr_fix_stim_nt1 %>%
  mutate(Bio_Mean_HR_dif = Bio_Mean_HR - Bio_Mean_HR_fix)
# 2 1
stim - fix
db_full6new_hr_fix_stim_nt1%>%
  ggplot(aes(Bio_Mean_HR_dif))+
  geom_histogram()

?scatter3d

db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(pup_basCor))%>%
   subset(!is.na(BIO_CDA.PhasicMax_z))%>%
  subset(!is.na(Bio_Mean_HR_dif))

db_full6new_hr_fix_stim_nt1<- db_full6new_hr_fix_stim_nt1%>%
  group_by(ssid)%>%
  mutate(BIO_CDA.PhasicMax_z= scale(BIO_CDA.PhasicMax, center = TRUE, scale = TRUE)[,1])

scatter3d(x = db_full6new_hr_fix_stim_nt1$pup_basCor, 
          y = db_full6new_hr_fix_stim_nt1$BIO_CDA.PhasicMax_z, 
          z = db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif, 
          fit = "smooth",

          surface = FALSE,
          point.col = "blue",
          # point
          groups = as.factor(db_full6new_hr_fix_stim_nt1$mediansplit_ground_arousal),
          ellipsoid = TRUE)

cor(db_full6new_hr_fix_stim_nt1[,c("pup_basCor", "BIO_CDA.PhasicMax_z","Bio_Mean_HR_dif")], use = "complete")

db_full6new_hr_fix_stim_nt1%>%
  group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(pup_basCor,BIO_CDA.PhasicMax_z, color= mediansplit_ground_valence))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ylim(0,.4)+
  ggpubr::stat_cor()

db_full6new_hr_fix_stim_nt1%>%
  # group_by(ssid)%>%
  # mutate(Bio_Mean_HR_dif_z = scale(Bio_Mean_HR_dif))
  group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(pup_basCor,Bio_Mean_HR_dif_z, color= mediansplit_ground_valence))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  # ylim(-7,2)+
  ggpubr::stat_cor()

db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  # group_by(ssid)%>%
  # mutate(Bio_Mean_HR_dif_z = scale(Bio_Mean_HR_dif))
  group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(BIO_CDA.PhasicMax_z,Bio_Mean_HR_dif, color= mediansplit_ground_valence))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  # ylim(-7,2)+
  ggpubr::stat_cor()+
  facet_grid(~mediansplit_ground_valence)

```

```{r}

db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif_outl<- if_else(abs(db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif)>
                                                             (median(abs(db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif),na.rm = TRUE) +                                                                                                        (4*(sd(abs(db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif), na.rm = TRUE)))), "outlier", "not outlier")

db_full6new_hr_fix_stim_nt1%>%
  ggplot(aes(Bio_Mean_HR_dif))+
  geom_histogram()+
  facet_grid(~Bio_Mean_HR_dif_outl)

```

```{r}
db_full6new_hr_fix_stim_nt1 %>%
    subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  subset(ssid_num<500)%>%
  group_by(ssid)%>%
  mutate(Bio_Mean_HR_dif_ssid = mean(Bio_Mean_HR_dif, na.rm = TRUE))%>%
  group_by(ssid)%>%
  # mutate(Bio_Mean_HR_dif_stim = mean(Bio_Mean_HR_dif, na.rm = TRUE))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(TASc, Bio_Mean_HR_dif))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

db_full6new_hr_fix_stim_nt1 %>%
    subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  subset(ssid_num<500)%>%
  # group_by(ssid)%>%
  mutate(Bio_Mean_HR_dif = abs(Bio_Mean_HR_dif))%>%
  group_by(ssid, stimIAPS)%>%
  mutate(Bio_Mean_HR_dif_stim = abs(Bio_Mean_HR_dif))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(TASc, Bio_Mean_HR_dif))+
  # geom_point()

  geom_smooth(aes(group = stimIAPS, y =Bio_Mean_HR_dif_stim), method=lm, se=FALSE, colour = "gray40", alpha = .01,linetype="dashed", size = .2 )+
  geom_smooth( method=lm, se=FALSE, linetype="dashed",
             color="darkred", size = 2)+
  stat_summary(geom = 'pointrange', pch = 1)+

  # geom_point()+
  # stat_summary(aes(group =ssid, y = Bio_Mean_HR_dif),method = "line",alpha = .1)+
  # ggpubr::stat_cor()+
  p$graphstyle+
  xlab("Alexithymia (Z)")+
  ylab("HR difference")+
p$graphstyle

hr_bio <- lmer(Bio_Mean_HR_dif ~ TASc *arousal_c + (1|ssid) + (1| stimIAPS),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt1, Bio_Mean_HR_dif_outl == "not outlier"))

interactions::interact_plot(hr_bio, pred = arousal_c, modx = TASc)

subset(db_full6new_hr_fix_stim_nt1, Bio_Mean_HR_dif_outl == "not outlier")

```

```{r}
db_full6new_hr_fix_stim_nt1 %>%
  subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  group_by(ssid)%>%
  mutate(cor_hrdif_ar = cor(Bio_Mean_HR_dif, arousal_c))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(TAS, cor_hrdif_ar))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  p$graphstyle

db_full6new_hr_fix_stim_nt1 %>%
  subset(TAS<70)%>%
  subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  group_by(ssid,mediansplit_ground_valence)%>%
  mutate(cor_hrdif_ar = cor(Bio_Mean_HR_dif, arousal_c))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, cor_hrdif_ar))+
  geom_jitter(alpha = .1, width = .1)+
  stat_summary(geom = 'pointrange')
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  p$graphstyle

```

db_full6new_hr_fix_stim_nt1 %>%
  # subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  group_by(ssid)%>%
  mutate(cor_hrdif_ar = cor(Bio_Mean_HR_dif, arousal_c))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(TAS, cor_hrdif_ar))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

```{r}
db_full6new_hr_fix_stim_nt1<- db_full6new_hr_fix_stim_nt1%>%
  group_by(stimIAPS)%>%
  mutate(ValenceMeanThisSample = mean(valence, na.rm = TRUE), 
         ArousalMeanThisSample = mean(arousal, na.rm = TRUE))

db_full6new_hr_fix_stim_nt1<- db_full6new_hr_fix_stim_nt1%>%
  ungroup()%>%
  mutate(mediansplit_ground_valence = if_else(ValenceMean >=
                                                  median(ValenceMean,
                                                         na.rm = TRUE), "More positive", "More negative"))%>%
  mutate(mediansplit_ground_arousal = if_else(ArousalMean >=
                                                  median(ArousalMean,
                                                         na.rm = TRUE), "High", "Low"))%>%
    mutate(mediansplit_sample_valence2 = if_else(ValenceMeanThisSample >=
                                                  median(ValenceMeanThisSample,
                                                         na.rm = TRUE), "More positive", "More negative"))%>%
  mutate(mediansplit_sample_arousal2 = if_else(ArousalMeanThisSample >=
                                                  median(ArousalMeanThisSample,
                                                         na.rm = TRUE), "High", "Low"))

db_full6new_hr_fix_stim_nt1%>%
  subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  group_by(ssid,mediansplit_ground_arousal, mediansplit_ground_valence,)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_arousal, Bio_Mean_HR_dif))+
  geom_jitter(alpha = .2, width = .2)+
   geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
  # stat_smooth(aes(group = ssid), se = F, method = 'lm', alpha = .2)+
    facet_grid(~mediansplit_ground_valence)

```

```{r}  
db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(Bio_Mean_HR_dif_outl== "not outlier")%>%
  group_by(ssid, mediansplit_ground_arousal,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_arousal, Bio_Mean_HR_dif, fill = mediansplit_ground_valence))+
  geom_jitter(alpha = .2, width = .2)+
   geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
  stat_summary(aes(group = ssid), geom = 'line', alpha = .2)+

  facet_grid(~mediansplit_ground_valence)+
     ggpubr::stat_compare_means()
  # p$graphstyle+
  scale_fill_brewer(palette = "Dark2")+

```
  # stat_smooth(aes(group = ssid), se = F, method = 'lm', alpha = .2)

db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(Bio_Mean_HR_dif_outl!= "outlier")%>%
  group_by(ssid, mediansplit_ground_arousal,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_arousal, Bio_Mean_HR_dif))+
  geom_jitter(alpha = .2, width = .2)+
   geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
  ggpubr::stat_compare_means()+
  facet_grid(~mediansplit_ground_valence)

# self

db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(Bio_Mean_HR_dif_outl!= "outlier")%>%
  group_by(ssid, mediansplit_sample_arousal2,mediansplit_sample_valence2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_sample_arousal2, Bio_Mean_HR_dif))+
  geom_jitter(alpha = .2, width = .2)+
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1, size = .4, colour = "Black") +
   # geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
  ggpubr::stat_compare_means()+
  facet_grid(~mediansplit_sample_valence2)

lmer(
  db_full6new_hr_fix_stim_nt1)

db_full6new_hr_fix_stim_nt1%>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal))+
  geom_histogram()

```{r}
# test - analyse just positive
hr_dif_models$arousal_val <- lmer(Bio_Mean_HR_dif ~  valence_c *arousal_c+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt1, mediansplit_ground_valence == "More positive" & Bio_Mean_HR_dif_outl == "not outlier"  & Group == "NT"))

summary(hr_dif_models$arousal_val)

unique(db_full6new_hr_fix_stim_nt1$ssid)
hr_dif_models$valfrom_hr <- lmer(valence_c ~   Bio_Mean_HR_dif* arousal_c+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt1, mediansplit_ground_valence == "More positive" &
                   Bio_Mean_HR_dif_outl == "not outlier" & Group == "NT"))
arousal_c

summary(hr_dif_models$valfrom_hr)

interactions::interact_plot(hr_dif_models$valfrom_hr, pred = Bio_Mean_HR_dif, modx = arousal_c)

plot(hr_dif_models$arousal_val)

car::vif(hr_dif_models$arousal_val)
 # valence_c           arousal_c valence_c:arousal_c 
           # 1.767605            1.708642            1.070635

?step
?lmerTest::step()
lmerTest::step(hr_dif_models$arousal_val)

# favours valence

# check the response bias thing
db_full6new_hr_fix_stim_nt1 %>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  ggplot(aes(log10(arousal+10.5) ,Bio_Mean_HR_dif))+
    # geom_point(alpha = .1)+
    facet_grid(~mediansplit_ground_valence)+
  geom_smooth(method = 'lm', se = F)+
  ggside::geom_ysidedensity(aes(x=stat(density, fill = mediansplit_self_valence), alpha = .4))+
ggside::geom_xsidedensity(aes(y=stat(density, fill = mediansplit_self_valence), alpha = .4))

scatter3d(x = db_full6new_hr_fix_stim_nt1$valence_c, 
          y =db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif, z =db_full6new_hr_fix_stim_nt1$arousal_c,
          # fit = "smooth",

          # surface = FALSE,
          point.col = "blue", 
          # point
          # ellipsoid = db_full6new_hr_fix_stim_nt1$mediansplit_sample_valence2)
          ellipsoid = TRUE)

db_full6new_hr_fix_stim_nt1%>%
  group_by(stimIAPS, mediansplit_ground_valence, mediansplit_ground_arousal)%>%
  ggplot(aes(mediansplit_ground_valence, ValenceMean))+
  geom_point()

db_full6new_hr_fix_stim_nt1%>%
  subset(Condition == "IAPS")%>%
  group_by(stimIAPS, mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, arousal))+
  stat_summary(geom = 'pointrange')+
  ggpubr::stat_compare_means()+
  geom_text(aes(label = stimIAPS))

```

check corelation between arousal and valence on and individual basis

```{r}

db_full6new_hr_fix_stim_nt1$arousal_c
subset(db_full6new_hr_fix_stim_nt1, mediansplit_ground_valence == "More positive" & Bio_Mean_HR_dif_outl == "not outlier") %>%
  group_by(ssid)%>%
  mutate(cor_vale_ar = cor(arousal, valence, use  = "complete"))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(cor_vale_ar))+
    geom_histogram()+
  geom_vline(xintercept = -.7)

subset(db_full6new_hr_fix_stim_nt1)%>%
  # group_by(mediansplit_ground_valence, stimIAPS)%>%
  # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes( valence, arousal))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  # geom_smooth(aes(group = ssid), se  = F, method = 'lm')+
  # facet_grid(~mediansplit_ground_valence)+
  ggpubr::stat_cor()

subset(db_full6new_hr_fix_stim_nt1)%>%
  subset(ssid == 331)%>%
  # group_by(mediansplit_ground_valence, stimIAPS)%>%
  # summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes( as.numeric(tNo), arousal))+
  geom_point()+
  stat_summary(geom = "line")+
  geom_point(aes(y = valence), color = "green")+
  stat_summary(aes(y = valence),geom = "line", linetype = "dashed", color = "green")
  # geom_smooth(method = 'lm', se = F)+
  # geom_smooth(aes(group = ssid), se  = F, method = 'lm')+
  # facet_grid(~mediansplit_ground_valence)+
  # ggpubr::stat_cor()

```

# try and regress valence our oarousal_val

db_full6new_hr_fix_stim_nt2

db_full6new_hr_fix_stim_nt3<- subset(db_full6new_hr_fix_stim_nt2, !is.na(valence))

hr_dif_models$arosal_from_val <- lmer(arousal_c~  valence_c +
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt3))

hr_dif_models$val_from_arousal <- lmer(valence_c~arousal_c+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt3))

hr_dif_models$arosal_from_val
plot(hr_dif_models$arosal_from_val)

?residuals
?resid
?residuals.merMod
db_full6new_hr_fix_stim_nt3$arous_val_resid <- residuals(hr_dif_models$arosal_from_val)

db_full6new_hr_fix_stim_nt3$val_from_arous_resid<- residuals(hr_dif_models$val_from_arousal)

# residuals.merMod(hr_dif_models$arosal_from_val)

ar_frm_val<- lm(arousal_c ~ valence_c, data = db_full6new_hr_fix_stim_nt3)

val_frm_val<- lm( valence_c~arousal_c, data = db_full6new_hr_fix_stim_nt3)

db_full6new_hr_fix_stim_nt3%>%ggplot(aes(val_from_arous_resid, arous_val_resid))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

```

resid(ar_frm_val)
resid(val_frm_val)
 #           1            2            3            4            5            6            7            8 
 # 0.149962843  0.505815507
 # 
 # -0.637512958  0.044580602

Is the residual of A~ B the same as B~A 
```{r}

hr_dif_models$val_from_arousal <- lmer(valence_c~arousal_c+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt3))

hr_dif_models$biohr_dif_arous_res <- lmer(Bio_Mean_HR_dif~  val_from_arous_resid *arous_val_resid+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = subset(db_full6new_hr_fix_stim_nt3, Bio_Mean_HR_dif_outl == "not outlier"))

summary(hr_dif_models$biohr_dif_arous_res)

# analyses of residuals also suggets that effects on bio_hr are related to common arousal and valence relation

```

interactions::interact_plot(hr_dif_models$arousal_val, pred= arousal_c, modx = valence_c)

# alexithjymia
db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif_z<- scale(db_full6new_hr_fix_stim_nt1$Bio_Mean_HR_dif, center = TRUE, scale = TRUE)[,1]
hr_dif_models$arousal_val_cat <- lmer(Bio_Mean_HR_dif_z ~TASc*arousal_c+
                                        + (1 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = db_full6new_hr_fix_stim_nt1)

interactions::interact_plot(hr_dif_models$arousal_val_cat, pred= arousal_c, modx = TASc)

summary(hr_dif_models$arousal_val_cat)

plot(hr_dif_models$arousal_val)
plot(hr_dif_models$arousal_val_cat)

```
summary(hr_dif_models$arousal_val_cat)

hr_dif_models$arousal_val_cat <- lmer(Bio_Mean_HR_dif~  mediansplit_sample_arousal2* mediansplit_sample_valence2+
                                        + (1+mediansplit_sample_valence2 | ssid) + (1|stimIAPS),
                                        # (1|mediansplit_sample_arousal2),
     REML = FALSE,
     data = db_full6new_hr_fix_stim_nt1)

summary(hr_dif_models$arousal_val_cat)

interactions::cat_plot(hr_dif_models$arousal_val_cat, pred = mediansplit_sample_arousal2, modx = mediansplit_sample_valence2)

```
# correlation

```{r}
db_full6new_hr_fix_stim_nt1%>%
  subset(!is.na(mediansplit_ground_valence))%>%
  subset(Bio_Mean_HR_dif_outl!= "outlier")%>%
  group_by(ssid,mediansplit_ground_valence, mediansplit_ground_arousal)%>%
  mutate(cor_hr_arous = cor(Bio_Mean_HR_dif, arousal, use = "complete"))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(mediansplit_ground_valence, cor_hr_arous))+
  # geom_jitter(alpha = .2, width = .2)+
  geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
  # geom_smooth(aes(group = ssid), color = 'gray', method = 'lm', se = F)+
  ggpubr::stat_compare_means()+
  facet_grid(~mediansplit_ground_arousal)

```
  # xlim(-5,3)

```{r}
db_full6new_hr_fix_stim_nt1 %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(Age, TASc))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

hr_dif_models<- list()

hr_dif_models$arousal_val <- lmer(Bio_Mean_HR_dif~  (arousal_c*valence_c )* TASc +(1+arousal_c | ssid) + (1|stimIAPS),
     REML = FALSE,
     data = db_full6new_hr_fix_stim_nt1)

summary(hr_dif_models$arousal_val)

interactions::interact_plot(hr_dif_models$arousal_val, pred = arousal_c, modx = TASc)

hr_dif_models$arousal_val <- lmer(Bio_Mean_HR~  arousal_c* valence_c +(1+arousal_c| ssid),
                                  # + (1|stimIAPS),
     REML = FALSE,
     data = db_full6new_hr_fix_stim_nt1)

summary(hr_dif_models$arousal_val)

# plot(hr_dif_models$arousal_val)

interactions::interact_plot(hr_dif_models$arousal_val, pred = valence_c, modx = arousal_c)

```

```

import HRV resting
```{r}

Rest_HRV <- read_excel("Rest HRV.xlsx")
View(Rest_HRV)

Rest_HRV%>%
  ggplot(aes(RMSSD))+
  geom_histogram()

Rest_HRV%>%
  ggplot(aes(RMSSD))+
  geom_histogram()

Rest_HRV%>%
  ggplot(aes(MEANHR))+
  geom_histogram()

BIO_HRVanalysys$ssid<- as.character(BIO_HRVanalysys$ssid)

left_join(BIO_HRVanalysys, Rest_HRV, by = "ssid")%>%
  ggplot(aes(RMSSD, RMMSSD))+
  geom_point()+
  ggpubr::stat_cor()

# library(readxl)

Rest_practice_HRV <- read_excel("Rest + practice HRV.xlsx")
View(Rest_practice_HRV)
Rest_practice_HRV$ssid<- as.character(Rest_practice_HRV$ssid)

left_join(Rest_practice_HRV, BIO_HRVanalysys, by = "ssid")%>%
  ggplot(aes(RMSSD, RMMSSD))+
  geom_point()+
  ggpubr::stat_cor()

left_join(Rest_practice_HRV, Rest_HRV, by = "ssid")%>%
  ggplot(aes(RMSSD.x, RMSSD.y))+
  geom_point()+
  ggpubr::stat_cor()

Rest_practice_HRV%>%
  ggplot(aes(MEANHR))+
  geom_histogram()

# import hrv task
library(readxl)
hrvResultsTask <- read_excel("Summary_hrvResultsTask.xls")
View(hrvResultsTask)

```

combine with data

```{r}
Rest_HRV

Rest_HRV$subject<- NULL
nrow(db_full6new_hr_fix_stim_nt1)
# 2160

db_full6new_hr_fix_stim_nt2<- left_join(db_full6new_hr_fix_stim_nt1, Rest_HRV)

# hrv rest vs hrtaskCallbackManager(
left_join(Rest_HRV, hrvResultsTask, by = "ssid") %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(RMSSD.x, RMSSD.y))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  xlim(0,100)+
  ylim(0,100)

Rest_HRV$hrvcondition<- "rest"
hrvResultsTask$hrvcondition<- "task"

left_join(Rest_HRV, hrvResultsTask, by = "ssid") %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(RMSSD.x, RMSSD.y))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()+
  xlim(0,100)+
  ylim(0,100)

colnames(hrvResultsTask)
bind_rows(Rest_HRV, hrvResultsTask)%>%
  subset(RMSSD<100)%>%
  ggplot(aes(hrvcondition, hfpower))+
  geom_jitter(width = .1, alpha = .1)+
  stat_summary(geom = 'pointrange')+
  stat_summary(aes(group = ssid), geom = 'line', alpha = .1)
```

```{r}
left_join(db_full6new_hr_fix_stim_nt1, hrvResultsTask) %>%
  group_by(ssid)%>%
  mutate(cor = cor(arousal, pup_basCor, use = "complete"))%>%
    group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(NN50,cor ))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  xlim(0,500)+
  ggpubr::stat_cor()

  left_join(db_full6new_hr_fix_stim_nt1, hrvResultsTask) %>%
  group_by(ssid)%>%
  mutate(cor = cor(arousal, pup_basCor, use = "complete"))%>%
    group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(TASc, NN50))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
      # ylim(0,200)+
  ggpubr::stat_cor()

  hrvResultsTask%>%
    ggplot(aes(RMSSD))+
    geom_histogram()

  left_join(db_full6new_hr_fix_stim_nt1, hrvResultsTask) %>%
    subset(!is.na(mediansplit_ground_valence))%>%
  # group_by(ssid)%>%
  # mutate(cor = cor(arousal, pup_basCor, use = "complete"))%>%
    group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(RMSSD,pup_basCor,color = mediansplit_ground_valence ))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  # xlim(0,100)+
  ggpubr::stat_cor()

    left_join(db_full6new_hr_fix_stim_nt1, hrvResultsTask) %>%
    subset(!is.na(mediansplit_ground_valence))%>%
  # group_by(ssid)%>%
  # mutate(cor = cor(arousal, pup_basCor, use = "complete"))%>%
    group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(RMSSD,BIO_CDA.PhasicMax_z,color = mediansplit_ground_valence ))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  xlim(0,100)+
  ggpubr::stat_cor()

    left_join(db_full6new_hr_fix_stim_nt1, hrvResultsTask) %>%
    subset(!is.na(mediansplit_ground_valence))%>%
       subset(Bio_Mean_HR_dif_outl == "not outlier")%>%
  # group_by(ssid)%>%
  # mutate(cor = cor(arousal, pup_basCor, use = "complete"))%>%
    group_by(ssid,mediansplit_ground_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(`LF/HF`,BIO_CDA.PhasicMax_z,color = mediansplit_ground_valence ))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  # xlim(0,5000)+
  ggpubr::stat_cor()
```

db_full6new_hr_fix_stim_nt2 %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(Bio_Mean_HR_dif, lfpower))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

db_full6new_hr_fix_stim_nt2 %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(MEANHR, RMSSD))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

db_full6new_hr_fix_stim_nt2 %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(MEANHR, Bio_Mean_HR))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

```

```{r}
# work on this
db_full6new_304_up$
colnames(db_full4new_stim_screen_pupil_nopract)
colnames(db_full6new_304_up)
# db_full6new
unique(db_full6new_304_up$screencontent)

View(db_full6new)

# store heart rate baseline to compute difference
db_full6new$Bio_Mean_HR_fix = if_else(db_full6new$screencontent == "fixation", db_full6new$Bio_Mean_HR, NULL)

# compute heart rate difference to fixation screen

db_full6new_304_up$Bio_Mean_HR
View(db_full6new_304_up)

db_full6new_304_up<- db_full6new_304_up%>%
  group_by(ssid)%>%
  mutate(BIO_CDA.PhasicMax_z = scale(BIO_CDA.PhasicMax)[,1])

pupulresidtest<- lmer(pup_basCor ~ BRIGHTNESSc +(1+BRIGHTNESSc|ssid), REML = FALSE,
     data = testresiddf)

summary(pupulresidtest)

plot(pupulresidtest)

resid(pupulresidtest)
tstresid <- resid(pupulresidtest)

testresiddf$puopil_resid <- tstresid

testresiddf<- subset(db_full6new_304_up, !is.na(pup_basCor))

summary(lmer(arousal ~ pupil_resid + (1|ssid), REML = FALSE,
     data = testresiddf))

summary(lmer(arousal ~ puopil_resid + Bio_Mean_HR+BIO_CDA.PhasicMax + (1|ssid),
              # (0+puopil_resid + Bio_Mean_HR+BIO_CDA.PhasicMax|ssid), 
             REML = FALSE,
     data = subset(testresiddf, as.numeric(as.character(ssid))<500)))

subset(testresiddf, as.numeric(as.character(ssid))<500)%>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(arousal, BIO_CDA.PhasicMax))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)+
  ggpubr::stat_cor()

subset(testresiddf, as.numeric(as.character(ssid))<500)%>%
  mutate(median_ground_arousal = if_else(ArousalMean> median(ArousalMean, na.rm = TRUE), "High", "Low"))%>%
  mutate(median_ground_valence = if_else(ValenceMean> median(ValenceMean, na.rm = TRUE), "More positive", "More negative"))%>%
  group_by(ssid)%>%
  ungroup()%>%
  mutate(median_self_arousal = if_else(arousal> median(arousal), "High", "Low"))%>%
  mutate(median_self_valence = if_else(valence> median(valence), "Positive", "Negative"))%>%
group_by(ssid,median_self_arousal,median_self_valence)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(median_self_arousal, puopil_resid,))+
  geom_jitter(width = .2, alpha = .2)+
   geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
  stat_summary(geom = 'pointrange')+
    facet_grid(~median_self_valence)
  ggpubr::stat_cor()

  subset(testresiddf, as.numeric(as.character(ssid))<500)%>%
  mutate(median_ground_arousal = if_else(ArousalMean> median(ArousalMean, na.rm = TRUE), "High", "Low"))%>%
  mutate(median_ground_valence = if_else(ValenceMean> median(ValenceMean, na.rm = TRUE), "More positive", "More negative"))%>%
  group_by(ssid)%>%
  mutate(median_self_arousal = if_else(arousal> median(arousal), "High", "Low"))%>%
  mutate(median_self_valence = if_else(valence> median(valence), "Positive", "Negative"))%>%
group_by(ssid,median_self_arousal)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
    subset(!is.na(median_self_arousal))%>%
  ggplot(aes(median_self_arousal, BIO_CDA.PhasicMax_z))+
  geom_jitter(width = .2, alpha = .2)+
   geom_bar(stat="summary", fun.y = "mean", alpha = .2)+
    # geom_smooth(aes(group = ssid),method = 'lm', se = F)+
    # geom_text(aes(label = ssid))
  stat_summary(geom = 'pointrange')
    facet_grid(~median_self_valence)
  ggpubr::stat_cor()

  summary(lmer(arousal ~ log(BIO_CDA.PhasicMax+.1) + (1|ssid) + (1+log(BIO_CDA.PhasicMax+.1)|stimIAPS),
              # (0+puopil_resid + Bio_Mean_HR+BIO_CDA.PhasicMax|ssid), 
             REML = FALSE,
     data = subset(testresiddf, as.numeric(as.character(ssid))<500 &BIO_CDA.AmpSum> .1)))

```