### 2021-08-20 Experiment 2.Rmd
Preprocessing
This is the updated code with the correct repixelation. This was tested in the validation paper, so use it as a model

---
title: "InteroceptionStudy_GPValidation"
Author: "Helio Cuve & "
date: "8 January 2021"
output: html_document
---

# Loading rdhf5 
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.10")

BiocManager::install("rhdf5")

library(rhdf5)

```

# Preferences
list necessary package install them if not already and load them
```{r}
# devtools::install_github("dmirman/gazer")

list.of.packages <- c("ggplot2", "data.table", 'readr', "rhdf5", 'tidyverse', "gazer", 'lmerTest')
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only = TRUE)
lapply(new.packages, require, character.only = TRUE)

library(purrr)
library(data.table)
library(dplyr)
library(tidyr)

# setting a workspace
# change approproatelly

setwd("D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020")
hdf5FilesPath<- "D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020"

# 2021 data
setwd("~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking/2021")
hdf5FilesPath<- "~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking/2021"
```

# gp extractor function
takes gp files and returns a clean files with messages parsed and deletes unnecessary stuff

note that starting from 322 we have gsr-v and hr-v, and pulse

```{r, gp extractor}

# hdf5FilesPath<- "D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking"

hdf5FilesPath<- "~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking/2021"

gp.extractor = function(hdf5FilesPath) {

  setwd(hdf5FilesPath)
  #getwd()
  hdfFilesList = list.files(hdf5FilesPath, pattern = "\\.hdf5$")

  if (length(hdfFilesList) == 0){

    return(message('No HDF5 files found in the directory provided.'))
  }
   f = 30
  for (f in 1:length(hdfFilesList)) {

    message(sprintf("Extracting file %i / %i - %s", f, length(hdfFilesList), hdfFilesList[f]))

    #note that some r-studio bug sometimes causes this to fail when runnin from notebook, 
    # in that case run it directly on command line
    tmp.events =h5read(hdfFilesList[f], '/data_collection/events/experiment/MessageEvent')

    tmp.eyetr  = h5read(hdfFilesList[f], '/data_collection/events/eyetracker/BinocularEyeSampleEvent')

    # Get subject id from Events
    ssid = sapply(strsplit(grep('SubjJID', tmp.events$text, value = TRUE), split = ": "), "[", 2)
    date = sapply(strsplit(grep('DATE', tmp.events$text, value = TRUE), split = ": "), "[", 2)
    session = sapply(strsplit(grep('Session:', tmp.events$text, value = TRUE), split = ":"), "[", 2)
    cali<- subset(tmp.events, grepl('Cali', tmp.events$text) == TRUE)

    # Create DF for trials data
    tmp.df = data.frame(matrix(ncol = ncol(tmp.eyetr)))
    names(tmp.df) = names(tmp.eyetr)
    colnames(tmp.df)

    # Prepare Events, keep only start/end messages
    phrases = c('tStart', 'tEnd')
    tmp.events_backup<- tmp.events
    # tmp.events<-  tmp.events_backup
    tmp.events = subset(tmp.events, grepl(paste(phrases, collapse = "|"), tmp.events$text))

    # Create start/end references
    tmp.events$tStart   = NA
    tmp.events$tEnd     = NA

    # Trial number extraction 
    tmp.events$tNo      = NA
    tmp.df$tNo          = NA
    tmp.df$Condition    = NA

    # Space for subject id
    tmp.df$ssid         = NA
    tmp.df$text<-NA

    for (l in 1:nrow(tmp.events)) {

      message(sprintf("Grabing start messages %s", tmp.events$text[l]))

      if ((grepl('tStart', tmp.events$text[l])) == TRUE) {
        tmp.events$tStart[l] = tmp.events$time[l]

        if ((grepl('tEnd', tmp.events$text[l+1])) == TRUE){
          tmp.events$tEnd[l] = tmp.events$time[l+1]
        }

        else {
          message('trial start/end structure not valid')
        }
      }

    }
    message(sprintf("done grabbing messages"))
    # readline(prompt="Press [enter] to continue")

    tmp.events$tNo<- as.numeric(as.character(gsub(".*_","",tmp.events$text)))

     message(sprintf("making screencontent"))

     tmp.eventsbackup<- tmp.events
    tmp.events$screencontent<- if_else(grepl('fixation', tmp.events$text), 'fixation',
                                       if_else(grepl('Valence', tmp.events$text), 'ValenceResp',
                                               if_else(grepl('rest', tmp.events$text), 'rest1', 
                                                       if_else(grepl('IntResp', tmp.events$text), 'IntResp','stim'))))

    # unique(tmp.events$screencontent)

    # atorw condition

    tmp.events$Condition<- if_else(grepl('IAPS', tmp.events$text), 'IAPS',
                                   if_else(grepl('Example', tmp.events$text), 'Practice',
                                           if_else(grepl('rest', tmp.events$text), 'Rest', NULL)))
    tmp.events$stim<- if_else(tmp.events$Condition == 'IAPS', sub(".*_ *(.*?) *jpg.*", "\\1", tmp.events$text),
                              if_else(tmp.events$Condition == 'Practice', sub(".*_ *(.*?) *jpg.*", "\\1", tmp.events$text), 'REST'))

    # substitute periods
    tmp.events$stim<- gsub('\\.', '.jpg', tmp.events$stim)

    tmp.events$tNo<- if_else(tmp.events$Condition == 'Practice', (tmp.events$tNo)-6,
                             if_else(tmp.events$Condition == 'Rest',-7, tmp.events$tNo))

    # create trial number < 0 are practice trials
    tmp.events$tNo<- if_else(tmp.events$tNo> -1, tmp.events$tNo+1, tmp.events$tNo) # avoid zeros

    tmp.events1 = subset(tmp.events, grepl('tStart',tmp.events$text))
    # tmp.events1$tNo1<- NULL
    # tmp.events1$tNo2<- NULL

    tmp.events1<- subset(tmp.events1, is.na(tmp.events1$tNo) == FALSE)

    # colnames(tmp.df)
    # colnames(tmp.df)
    # colnames(tmp.events1)
    # colnames(tmp.eyetr)

    tmp.df$screencontent<- NA
    tmp.df$stim<- NA
    tmp.events1$tEnd<- if_else(tmp.events1$tNo == -7, tmp.events1$tStart+120, tmp.events1$tEnd)

    message(sprintf("subsetting eyetracking data"))

    for(e in 1:nrow(tmp.events1)) {

      # Subset the EyeTracking data based on start/end
      tmp.raw.trial = subset(tmp.eyetr, tmp.eyetr$time >= tmp.events1$tStart[e] & tmp.eyetr$time <= tmp.events1$tEnd[e])

      tmp.raw.trial$tNo = tmp.events1$tNo[e]

      tmp.raw.trial$Condition = tmp.events1$Condition[e]
      tmp.raw.trial$screencontent = tmp.events1$screencontent[e]
      tmp.raw.trial$stim = tmp.events1$stim[e]
      # tmp.raw.trial$stim = tmp.events1$[e]
      tmp.raw.trial$ssid = ssid
      tmp.raw.trial$text = tmp.events1$text[e]

      message(sprintf("binding tmp.df"))

      tmp.df= rbind(tmp.raw.trial, tmp.df)
    }
     message(sprintf("done binding"))

    # Arrange the dataframe byu time
    tmp.df<- tmp.df%>%
      arrange(time)

    # Save file for the task with subject name
    name = sprintf("ss%s_GPdata",ssid)
    message(sprintf("saving data"))
    write_csv(tmp.df, paste(name,".csv"))

    # write calibration file
    name2 = sprintf("ss%s_GPdata_cali", ssid)
    write_csv(cali, paste(name2,".csv"))

  }

  message("Processing Completed.")
  tmp.df
}

# run function
tmp_full<- gp.extractor(hdf5FilesPath)

```

Problem files
```{r}
# 355  - did not finish IAPS
# adjust below manually

hdf5FilesPath<- "~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking/2021"

gp.extractor = function(hdf5FilesPath) {

  setwd(hdf5FilesPath)
  #getwd()
  hdfFilesList = list.files(hdf5FilesPath, pattern = "\\.hdf5$")

  if (length(hdfFilesList) == 0){
    hdfFilesList

    return(message('No HDF5 files found in the directory provided.'))
  }
   f = 10
  for (f in 1:length(hdfFilesList)) {

    message(sprintf("Extracting file %i / %i - %s", f, length(hdfFilesList), hdfFilesList[f]))

    #note that some r-studio bug sometimes causes this to fail when runnin from notebook, 
    # in that case run it directly on command line
    tmp.events =h5read(hdfFilesList[f], '/data_collection/events/experiment/MessageEvent')

    tmp.eyetr  = h5read(hdfFilesList[f], '/data_collection/events/eyetracker/BinocularEyeSampleEvent')

    # Get subject id from Events
    ssid = sapply(strsplit(grep('SubjJID', tmp.events$text, value = TRUE), split = ": "), "[", 2)
    date = sapply(strsplit(grep('DATE', tmp.events$text, value = TRUE), split = ": "), "[", 2)
    session = sapply(strsplit(grep('Session:', tmp.events$text, value = TRUE), split = ":"), "[", 2)
    cali<- subset(tmp.events, grepl('Cali', tmp.events$text) == TRUE)

    # Create DF for trials data
    tmp.df = data.frame(matrix(ncol = ncol(tmp.eyetr)))
    names(tmp.df) = names(tmp.eyetr)
    colnames(tmp.df)

    # Prepare Events, keep only start/end messages
    phrases = c('tStart', 'tEnd')
    tmp.events_backup<- tmp.events
    # tmp.events<-  tmp.events_backup
    tmp.events = subset(tmp.events, grepl(paste(phrases, collapse = "|"), tmp.events$text))

    # Create start/end references
    tmp.events$tStart   = NA
    tmp.events$tEnd     = NA

    # Trial number extraction 
    tmp.events$tNo      = NA
    tmp.df$tNo          = NA
    tmp.df$Condition    = NA

    # Space for subject id
    tmp.df$ssid         = NA
    tmp.df$text<-NA

    for (l in 1:nrow(tmp.events)) {

      message(sprintf("Grabing start messages %s", tmp.events$text[l]))

      if ((grepl('tStart', tmp.events$text[l])) == TRUE) {
        tmp.events$tStart[l] = tmp.events$time[l]

        if ((grepl('tEnd', tmp.events$text[l+1])) == TRUE){
          tmp.events$tEnd[l] = tmp.events$time[l+1]
        }

        else {
          message('trial start/end structure not valid')
        }
      }

    }
    message(sprintf("done grabbing messages"))
    # readline(prompt="Press [enter] to continue")

    tmp.events$tNo<- as.numeric(as.character(gsub(".*_","",tmp.events$text)))

     message(sprintf("making screencontent"))

     tmp.eventsbackup<- tmp.events
    tmp.events$screencontent<- if_else(grepl('fixation', tmp.events$text), 'fixation',
                                       if_else(grepl('Valence', tmp.events$text), 'ValenceResp',
                                               if_else(grepl('rest', tmp.events$text), 'rest1', 
                                                       if_else(grepl('IntResp', tmp.events$text), 'IntResp','stim'))))

    # unique(tmp.events$screencontent)

    tmp.events$Condition<- if_else(grepl('IAPS', tmp.events$text), 'IAPS',
                                   if_else(grepl('Example', tmp.events$text), 'Practice',
                                           if_else(grepl('rest', tmp.events$text), 'Rest', NULL)))
    tmp.events$stim<- if_else(tmp.events$Condition == 'IAPS', sub(".*_ *(.*?) *jpg.*", "\\1", tmp.events$text),
                              if_else(tmp.events$Condition == 'Practice', sub(".*_ *(.*?) *jpg.*", "\\1", tmp.events$text), 'REST'))

    # substitute periods
    tmp.events$stim<- gsub('\\.', '.jpg', tmp.events$stim)

    tmp.events$tNo<- if_else(tmp.events$Condition == 'Practice', (tmp.events$tNo)-6,
                             if_else(tmp.events$Condition == 'Rest',-7, tmp.events$tNo))

    # create trial number < 0 are practice trials
    tmp.events$tNo<- if_else(tmp.events$tNo> -1, tmp.events$tNo+1, tmp.events$tNo) # avoid zeros

    tmp.events1 = subset(tmp.events, grepl('tStart',tmp.events$text))
    # tmp.events1$tNo1<- NULL
    # tmp.events1$tNo2<- NULL

    tmp.events1<- subset(tmp.events1, is.na(tmp.events1$tNo) == FALSE)

    # colnames(tmp.df)
    # colnames(tmp.df)
    # colnames(tmp.events1)
    # colnames(tmp.eyetr)

    tmp.df$screencontent<- NA
    tmp.df$stim<- NA
    tmp.events1$tEnd<- if_else(tmp.events1$tNo == -7, tmp.events1$tStart+120, tmp.events1$tEnd)

    message(sprintf("subsetting eyetracking data"))

    for(e in 1:nrow(tmp.events1)) {

      # Subset the EyeTracking data based on start/end
      tmp.raw.trial = subset(tmp.eyetr, tmp.eyetr$time >= tmp.events1$tStart[e] & tmp.eyetr$time <= tmp.events1$tEnd[e])

      tmp.raw.trial$tNo = tmp.events1$tNo[e]

      tmp.raw.trial$Condition = tmp.events1$Condition[e]
      tmp.raw.trial$screencontent = tmp.events1$screencontent[e]
      tmp.raw.trial$stim = tmp.events1$stim[e]
      # tmp.raw.trial$stim = tmp.events1$[e]
      tmp.raw.trial$ssid = ssid
      tmp.raw.trial$text = tmp.events1$text[e]

      message(sprintf("binding tmp.df"))

      tmp.df= rbind(tmp.raw.trial, tmp.df)
    }
     message(sprintf("done binding"))

    # Arrange the dataframe byu time
    tmp.df<- tmp.df%>%
      arrange(time)

    # Save file for the task with subject name
    name = sprintf("ss%s_GPdata",ssid)
    message(sprintf("saving data"))
    write_csv(tmp.df, paste(name,".csv"))

    # write calibration file
    name2 = sprintf("ss%s_GPdata_cali", ssid)
    write_csv(cali, paste(name2,".csv"))

  }

  message("Processing Completed.")
  tmp.df
}

unique(tmp.df$tNo)

```

Now cleaning gaze data and prepare data for ledalab and artifact

```{r, cleaning}
# function to read all the csvs into one dataframe

setwd("~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Eyetracking/Step1_GpExtractor_Output/346-355")

# rm(tmp_full)
read_plus <- function(flnm) {
  data.table::fread(flnm, select = c("session_id","event_id", "time", "left_gaze_x", "left_gaze_y","right_gaze_x", "right_gaze_y","left_pupil_measure1",
                                     "right_pupil_measure1","gsr", "gsrv", 'hr', "hrv", "hrp", "status", "tNo",  "Condition", "screencontent", "stim", "ssid", "text")) %>% 
    mutate(filename = flnm)
}

# read all clean data 
tmp.df2 <-
  list.files(pattern = "*.csv", 
             full.names = T) %>% 
  map_df(~read_plus(.))

unique(tmp.df2$ssid)

table(is.na(tmp.df2$ssid))
# tmp.df2$ssid<- substr(tmp.df2$filename, 5,7)
head(tmp.df2)
# check we have all participants
# unique(tmp.df2$ssid)
tmp.df2<- subset(tmp.df2, !is.na(ssid))

```

#####################################################################################
# gp preprocessing funtion, checks for losses, averages gaze, pupil and pixelate
#####################################################################################

```{r, gppreprocessing}
gp_preprocessing = function(tmp.df2) {

  message(sprintf("step1 lft gaze validity"))

  # transform invalid gaze into NA based on validity codes

  tmp.df2$left_gaze_x_cor<- if_else(tmp.df2$status != '20', tmp.df2$left_gaze_x, NULL)
  tmp.df2$left_gaze_x_cor<- if_else(tmp.df2$status != '22', tmp.df2$left_gaze_x_cor, NULL)
  tmp.df2$left_gaze_y_cor<- if_else(tmp.df2$status != '20', tmp.df2$left_gaze_y, NULL)
  tmp.df2$left_gaze_y_cor<- if_else(tmp.df2$status != '22', tmp.df2$left_gaze_y_cor, NULL)

  message(sprintf("step2 right gaze validity"))

  tmp.df2$right_gaze_x_cor<- if_else(tmp.df2$status != '2', tmp.df2$right_gaze_x, NULL)
  #
  tmp.df2$right_gaze_x_cor<- if_else(tmp.df2$status != '22', tmp.df2$right_gaze_x_cor, NULL)
  tmp.df2$right_gaze_y_cor<- if_else(tmp.df2$status != '22', tmp.df2$right_gaze_y, NULL)
  tmp.df2$right_gaze_y_cor<- if_else(tmp.df2$status != '22', tmp.df2$right_gaze_y_cor, NULL)

  # pupil validity

  message(sprintf("step3 pupil validity"))
  tmp.df2$left_pupil_measure1<- if_else(tmp.df2$status != '20', tmp.df2$left_pupil_measure1, NULL)
  tmp.df2$left_pupil_measure1<- if_else(tmp.df2$status != '22', tmp.df2$left_pupil_measure1, NULL)

  tmp.df2$right_pupil_measure1<- if_else(tmp.df2$status != '2', tmp.df2$right_pupil_measure1, NULL)
  tmp.df2$right_pupil_measure1<- if_else(tmp.df2$status != '22', tmp.df2$right_pupil_measure1, NULL)
  table(is.na(tmp.df2$right_pupil_measure1))

  # check gazepoint out of bounds nd repixelate (1920/2 = 960 by 1200/2 = 600. add half to normalised coordinates from psychopy)

  message(sprintf("step 4 cutting off screen in points and repixelate"))

  tmp.df2$left_gaze_x_cor<- if_else(tmp.df2$left_gaze_x_cor >= -960 & tmp.df2$left_gaze_x_cor <= 960, tmp.df2$left_gaze_x_cor, NULL)
  tmp.df2$right_gaze_x_cor<- if_else(tmp.df2$right_gaze_x_cor >= -960 & tmp.df2$right_gaze_x_cor <= 960, tmp.df2$right_gaze_x_cor, NULL)

  tmp.df2$left_gaze_y_cor<- if_else(tmp.df2$left_gaze_y_cor >= -600 & tmp.df2$left_gaze_y_cor <= 600, tmp.df2$left_gaze_y_cor, NULL)
  tmp.df2$right_gaze_y_cor<- if_else(tmp.df2$right_gaze_y_cor >= -600 & tmp.df2$right_gaze_y_cor <= 600, tmp.df2$right_gaze_y_cor, NULL)

  tmp.df2$left_gaze_x_cor_pix<- tmp.df2$left_gaze_x_cor+960
  tmp.df2$right_gaze_x_cor_pix<- tmp.df2$right_gaze_x_cor+960

  tmp.df2$left_gaze_y_cor_pix<- tmp.df2$left_gaze_y_cor+600
  tmp.df2$right_gaze_y_cor_pix<- tmp.df2$right_gaze_y_cor+600

  # is left or right an NA

  tmp.df2$left_na<- is.na(tmp.df2$left_gaze_x_cor) 
  tmp.df2$right_na<- is.na(tmp.df2$right_gaze_x_cor)
  tmp.df2$both_na<- if_else(tmp.df2$left_na == TRUE & tmp.df2$right_na == TRUE, TRUE, FALSE)

  message(sprintf("step5 gaze x  means"))

  tmp.df2$gaze_x_cor<-if_else(tmp.df2$left_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$right_gaze_x_cor,
                              if_else(tmp.df2$right_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$left_gaze_x_cor,
                                      ((tmp.df2$left_gaze_x_cor + tmp.df2$right_gaze_x_cor)/2)))
  message(sprintf("step6 gaze y means"))

  tmp.df2$gaze_y_cor<-if_else(tmp.df2$left_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$right_gaze_y_cor,
                              if_else(tmp.df2$right_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$left_gaze_y_cor,
                                      ((tmp.df2$left_gaze_y_cor + tmp.df2$right_gaze_y_cor)/2)))

  message(sprintf("step6 repixel gaze means"))

  tmp.df2$gaze_x_cor_pix<- tmp.df2$gaze_x_cor + 960
  min(tmp.df2$gaze_x_cor_pix, na.rm = TRUE)

  tmp.df2$gaze_y_cor_pix<- tmp.df2$gaze_y_cor+600
  min(tmp.df2$gaze_y_cor_pix, na.rm = TRUE)

  # clean_pupil

  message(sprintf("cleaning pupil"))

  max(tmp.df2$left_pupil_measure1)

  tmp.df2$pupil<- if_else(tmp.df2$left_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$right_pupil_measure1,
                          if_else(tmp.df2$right_na == TRUE & tmp.df2$both_na == FALSE, tmp.df2$left_pupil_measure1,
                                  ((tmp.df2$left_pupil_measure1 + tmp.df2$right_pupil_measure1)/2)))

  # tmp.df2$trackloss<- if_else(tmp.df2$status > 0, TRUE, FALSE)
  tmp.df2$trialUnq<- paste(tmp.df2$tNo, paste(tmp.df2$screencontent, paste(tmp.df2$stim)))
  tmp.df2$distance = 630

  # min(tmp.df2$pupil, na.rm = TRUE)
  # max(tmp.df2$pupil, na.rm = TRUE)
  # colnames(tmp.df2)
  tmp.df2<- tmp.df2%>%
    arrange(time)
}

tmp.df2<- gp_preprocessing(tmp.df2)
 colnames(tmp.df2)

min(tmp.df2$pupil, na.rm = TRUE)
max(tmp.df2$gaze_y_cor_pix, na.rm = TRUE)

```

Compute proportions outside the sccreen
repixe

# check gazepoint out of bounds nd repixelate (1920/2 = 960 by 1200/2 = 600. add half to normalised coordinates from psychopy)

  x -960 by 960
  y -600 by 600

```{r}
colnames(X_tmp_df4_full$trackloss)

range(tmp.df2$gaze_x_cor, na.rm = TRUE)

tmp.df2
range(tmp.df2$left_gaze_y, na.rm = TRUE)

range(tmp.df2$left_gaze_y, na.rm = TRUE)
range(tmp.df2$left_gaze_x, na.rm = TRUE)
tmp.df2$x_outside<- ifelse(tmp.df2$left_gaze_x < -960, TRUE,
                           ifelse(tmp.df2$left_gaze_x > 960, TRUE, FALSE))

tmp.df2$y_outside<- ifelse(tmp.df2$right_gaze_y < -600, TRUE,
                           ifelse(tmp.df2$right_gaze_y > 600, TRUE, FALSE))

table(tmp.df2$y_outside)
137/(185406+137)
0.0007383733*100

range(tmp.df, na.rm = TRUE)

```
# rezeroing times by trial, by participants

```{r}
tmp.df3<- tmp.df2

# create start at every screen content
tmp.df3<- tmp.df3%>% 
  arrange(ssid, time, tNo)%>%
  group_by(ssid)%>%
  mutate(firststart = duplicated(trialUnq))

head(tmp.df3)

# make zero-times
tmp.df3$zero<- if_else(tmp.df3$firststart == FALSE, tmp.df3$time, NULL)

# fill in the zeros
tmp.df3<- tmp.df3%>%
  group_by(ssid, Condition, tNo)%>%
  # group_by(text)%>%
  fill(zero)

unique(tmp.df3$zero)

# rezero
tmp.df3<-tmp.df3%>%
  arrange(ssid, time, tNo)%>%
  group_by(ssid, tNo)%>%
  mutate(timerezero = time - zero)

min(tmp.df3$timerezero, na.rm = TRUE)
max(tmp.df3$timerezero, na.rm = TRUE)
natest<-subset(tmp.df3, is.na(tmp.df3$ssid) == TRUE)

# let's check to see if it looks fine
ggplot(subset(tmp.df3, tmp.df3$tNo == -1), aes(timerezero, gsr, colour = as.factor(ssid)))+
  geom_line()

# create a new time rezero where before stim start, time is negative

tmp.df3<- tmp.df3%>%
  # arrange(time)%>%
  group_by(ssid, Condition, screencontent)%>%
  mutate(firststartSTIM = duplicated(tNo))

#this will create a time in the stim start which we can use to do the a second rezeroing
tmp.df3$zero_stim<- if_else(tmp.df3$firststartSTIM == FALSE & tmp.df3$screencontent == 'stim', tmp.df3$time, NULL)

    # fill down
tmp.df3<- tmp.df3%>%
      arrange(time)%>%
      group_by(tNo, ssid)%>%
      fill(zero_stim, .direction = c('updown'))

   tmp.df3$timerezero2<- tmp.df3$time-tmp.df3$zero_stim

   # check timerezero2
     range(tmp.df3$timerezero2, na.rm = TRUE)

# time 3 including rest
  # simply includes rest as is

tmp.df3$timerezero3<- if_else(tmp.df3$Condition == 'Rest', tmp.df3$timerezero, tmp.df3$timerezero2)
min(tmp.df3$timerezero3, na.rm = TRUE)

# quick check that everything is fine

ggplot(subset(tmp.df3, tmp.df3$tNo == 27 ) , aes(x = timerezero3, y= gsr, colour = as.factor(ssid)))+
    geom_smooth()+
    # geom_()+
    geom_vline(aes(xintercept = 0), colour = 'red', linetype = 'dashed', size = 2)+
    geom_vline(aes(xintercept = 4), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 8), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 6), colour = 'blue', linetype = 'dashed', size = 2)

ggplot(subset(tmp.df3, tmp.df3$tNo == -4) , aes(x = timerezero2, y= gsr, colour = as.factor(ssid)))+
  # geom_point()
      geom_line() +
    # geom_smooth(aes())+
    geom_vline(aes(xintercept = 0), colour = 'red', linetype = 'dashed', size = 2)+
    geom_vline(aes(xintercept = 4), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 8), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 6), colour = 'blue', linetype = 'dashed', size = 2)

ggplot(subset(tmp.df3, tmp.df3$tNo == -7) , aes(x = timerezero3, y= gsr, colour = as.factor(ssid)))+
  # geom_point()
      geom_line() +
    # geom_smooth(aes())+
    geom_vline(aes(xintercept = 0), colour = 'red', linetype = 'dashed', size = 2)+
    geom_vline(aes(xintercept = 4), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 8), colour = 'gray30', linetype = 'dotted', size = 2)+
    geom_vline(aes(xintercept = 6), colour = 'blue', linetype = 'dashed', size = 2)

```

triggers

```{r}
unique(tmp.df3$tNo)

natest<- subset(tmp.df3, is.na(tmp.df3$tNo) == TRUE)
tmp.df3<- subset(tmp.df3, is.na(tmp.df3$tNo) == FALSE)

tmp.df3$intercept_time<- if_else(tmp.df3$firststart == FALSE, tmp.df3$time, 
                          if_else(tmp.df3$firststartSTIM == FALSE, tmp.df3$time,
                                  NULL))

# takes a bit with a group - so don't run
# ggplot(tmp.df3, aes(time, gsr))+
#   geom_line()+
#   geom_vline(aes(xintercept = intercept_time))

unique(tmp.df3$tNo)

tmp.df3$trigger<- as.double(tmp.df3$tNo)

tmp.df3$trigger<- if_else(tmp.df3$firststart == FALSE, tmp.df3$trigger, 
                          if_else(tmp.df3$firststartSTIM == FALSE, tmp.df3$trigger, 0))
unique(tmp.df3$trigger)

tmp.df3$trigger <- if_else(tmp.df3$screencontent == 'fixation' &tmp.df3$firststart == FALSE, 100,
                          if_else(tmp.df3$screencontent == 'ValenceResp' & tmp.df3$firststart == FALSE, 101,
                             if_else(tmp.df3$screencontent == 'IntResp' & tmp.df3$firststart == FALSE, 102, tmp.df3$trigger)))

unique(tmp.df3$trigger)

tmp.df3$trigger<- if_else(tmp.df3$firststart == TRUE, 0, tmp.df3$trigger)

unique(tmp.df3$trigger)

```

Calibration analysis

```{r}

# fucntion to read all calibration files into one big dataframe
read_cali <- function(flnm) {
  data.table::fread(flnm) %>%
    mutate(filename = flnm)
}

# now read them
db_cali <-
  list.files(pattern = "*.csv",
             full.names = T) %>%
  map_df(~read_cali(.))

# view the dataset
View(db_cali)
unique(db_cali$filename)

# select needed columns
db_cali$ssid <- substring(db_cali$filename, 5,7)
db_cali$valid_points <- sub(".*_", "", db_cali$text)
db_cali$valid_points <- if_else(grepl('valid', db_cali$text), db_cali$valid_points, NULL)
unique(db_cali$valid_points)

db_cali$avg_error <- sub(".*_", "", db_cali$text)
db_cali$avg_error <- if_else(grepl('AvgError', db_cali$text), db_cali$avg_error, NULL)
db_cali$cali_no <- substring(db_cali$text, 23, 24)
db_cali$cali_no <- if_else(grepl('valid', db_cali$text), db_cali$cali_no, NULL)
db_cali$cali_no<- as.numeric(sub("_", "",db_cali$cali_no))

install.packages("useful")

library(useful)
db_cali<- db_cali%>%
  fill(cali_no)%>%
  fill(valid_points)%>%
  fill(avg_error, .direction = 'up')

#find the duplicate rows (some trigger messages were sent twice)
db_cali<- db_cali %>%
  arrange(ssid)%>%
  group_by(ssid)%>%
  mutate(isduplicate = duplicated(avg_error))

db_cali1 <- subset(db_cali, isduplicate == FALSE)

db_cali1 <- select(db_cali1, c(ssid, cali_no, valid_points, avg_error))

# create a order for the calibration
db_cali1$x = row.names(db_cali1)
db_cali1$cali_order <- ave(db_cali1$x,                 # Create numbering variable
                       db_cali1$ssid,
                       FUN = seq_along)
```

# FUNCTION TO CONVERT PIXELS TO DEGREES OF VISUAL ANGLE

```{r}

pixtoAng<- function(hw,d,r,stimsize){
  #hw - height/width of the monitor(cm) - depending on the axis
  #d - distance of the participant from the monitor (cm)
  #r - vertical/horizontal resolution of the screen (px)
  #stimsize - size of the stimulus on the screen (px)
  deg_per_px = rad2deg(atan2(0.5*hw, d))/(0.5*r)
  visAng <- stimsize * deg_per_px
  return (visAng)
}
#pixtoAng - hw - height/width of the screen in cm - use height for vertical acc/prec, and width for horizontal; d - distance of the participants to the screen (we used an average distance from the logbook); r = height/width of the screen in pixels; stimsize - basically, what you wish to convert from pixels to degrees of visual angle

```

```{r}
library(ggplot2)

db_cali1$avg_error1<- as.numeric(as.character(db_cali1$avg_error))

db_cali1 %>%
  subset(ssid!= 610)%>%
ggplot(aes(x = cali_order, avg_error1))+
  geom_line(aes(group = ssid))

db_cali1 %>%
  subset(ssid!= 610)%>%
ggplot(aes(x = cali_order, cali_no, color = ssid))+
  geom_line(aes(group = ssid))

db_cali1<- db_cali1%>%
  group_by(ssid)%>%
  mutate(cali_no_resc =  cali_no - lag(cali_no))%>%
  mutate(cali_no_resc = if_else(!is.na(cali_no_resc), cali_no_resc, cali_no))

db_cali1 %>%
  subset(ssid!= 610)%>%
ggplot(aes(x = cali_order, cali_no_resc, colour = ssid))+
  geom_line(aes(group = ssid))

#how many calibrations does it take until accepting
library(lmerTest)

cali_models<- list()
db_cali1$ssid<- as.character(db_cali1$ssid)

db_cali1$cali_order <- as.numeric(as.character(db_cali1$cali_order))
db_cali1$cali_order_z<- scale(db_cali1$cali_order, center = TRUE, scale = TRUE)

cali_models$cali_toaccept<- glmer(cali_no_resc ~ (1| ssid),
                                  family = "poisson", data = db_cali1)

summary(cali_models$cali_toaccept)

summary(glm(cali_no_resc ~ cali_order_z, data = db_cali1))

summary(cali_models$cali_toaccept)

db_cali2<- subset(db_cali1, ssid < 500)
mean(subset(db_cali2, cali_order == 1)$cali_no_resc, na.rm = TRUE)
sd(subset(db_cali2, cali_order == 1)$cali_no_resc, na.rm = TRUE)

# AVG cali
mean(db_cali1$cali_no_resc, na.rm = TRUE)

sd(db_cali1$cali_no_resc)

(mean(db_cali1$avg_error1, na.rm = TRUE)) #43.40358
(median(db_cali1$avg_error1, na.rm = TRUE)) #40.13

install.packages("rCAT")
library(rCAT)

# height and widrh of the sreen in pixes and cm
# H = 32.5 cm (108 Pixels) W = (52)  1920Pix by 1200, distance 63 - 65

pixtoAng(52, 65, 1920, 43.40358) # 0.9849095
pixtoAng(52, 65, 1920, 40.13) # 0.9849095

db_cali1$avg_error1_visang<- pixtoAng(52, 65, 1920, db_cali1$avg_error1)

mean(db_cali1$avg_error1_visang)
sd(db_cali1$avg_error1_visang)
range(db_cali1$avg_error1_visang)

db_cali2_id<- db_cali2%>%
  group_by(ssid)%>%
  summarise_at(c("avg_error"), mean, na.rm = TRUE)

db_cali2$Participant<- db_cali2$ssid

db_cali2_id$avg_error<- NULL
db_cali2$Participant<- NULL

db_cali2<- left_join(db_cali2, db_cali2_id)

db_cali2$Participant<- as.character(db_cali2$Participant)
db_cali2%>%
  ggplot(aes(cali_order, avg_error1))+
  geom_line(aes(group = Participant, colour = Participant))+
  stat_summary(aes(geom = "line"))+
  geom_smooth()

```

Trackloss

data = (tmp.df4_sum)

```{r}
# get average loss
mean(tmp.df4_sum$gaze_loss_prop)*100
sd(tmp.df4_sum$gaze_loss_prop)*100
range(accuracy_dataset$gaze_loss_prop)*100

range(accuracy_dataset)*100
range(db_cali1$cali_no_resc)

tmp.df4_sum %>%
  subset(ssid < 500)%>%
  group_by(screencontent)%>%
  summarise_at(c('gaze_loss_prop'), mean, na.rm = TRUE)

# store only neurotypicals

tmp.df4_sum_below500 <- tmp.df4_sum %>%
  subset(ssid < 500)

#plot loss overtime
tmp.df4_sum_below500%>%
  ggplot(aes(tNo, gaze_loss_prop))+
  stat_summary(geom = 'pointrange')+
  geom_vline(xintercept = 26)

#merge calibration information
db_cali2_summ<- db_cali2 %>%
  group_by(ssid)%>%
  mutate(avg_error_agg = mean(avg_error1, na.rm = TRUE), cali_no_resc_agg = mean(cali_no_resc, na.rm = TRUE))

# Save an object to a file
saveRDS(db_cali2_summ, file = "db_cali2_summ.rds")
# Restore the object
accuracy_dataset_below6s<- readRDS(file = "accuracy_dataset_below6s.rds")

View(db_cali2_summ)

tmp.df4_sum_below500$cali_order<- if_else(tmp.df4_sum_below500$tNo < 0,1, if_else(tmp.df4_sum_below500$tNo > 0 & tmp.df4_sum_below500$tNo< 26, 2, 3))

unique(tmp.df4_sum_below500$cali_order)

#merge

tmp.df4_sum_below500$cali_order<- as.character(tmp.df4_sum_below500$cali_order)
tmp.df4_sum_below500 <- left_join(tmp.df4_sum_below500, db_cali2_summ)

max(tmp.df4_sum_below500$ssid)
View(tmp.df4_sum_below500)

# just create one trackloss per trial for all screens
tmp.df4_sum_below500 %>%
  group_by(ssid)%>%
  summarise_if(is.numeric, mean, na.r = TRUE)%>%
  ggplot(aes(avg_error_agg, gaze_loss_prop))+
  geom_point()+
  geom_smooth(method = 'lm')
  facet_grid(~cali_order)

 unique(tmp.df4_sum_below500$cali_order)

# mixed models

calibration_lmerdata <-  tmp.df4_sum_below500 %>%
  group_by(ssid, tNo )%>%
  summarise_at(c('avg_error_agg','gaze_loss_prop', 'cali_no_resc_agg'), mean, na.rm = TRUE)

nrow(calibration_lmerdata)

loss_lmermodels<- list()

calibration_lmerdata$ot1<- poly(calibration_lmerdata$tNo, degree = 1, raw = FALSE)
calibration_lmerdata$ot2<- poly(calibration_lmerdata$tNo, degree = 2, raw = FALSE)[,2]

View(calibration_lmerdata)
calibration_lmerdata%>%
ggplot(aes(tNo, ot1))+
  geom_line()+
  geom_line(aes(y = ot2))

cor.test(calibration_lmerdata$ot1,calibration_lmerdata$ot2 )

calibration_lmerdata1<- subset(calibration_lmerdata, ssid> 303 & tNo > 0 & tNo < 26)

calibration_lmerdata1$ot1<- poly(calibration_lmerdata1$tNo, degree = 1, raw = FALSE)
calibration_lmerdata1$ot2<- poly(calibration_lmerdata1$tNo, degree = 2, raw = FALSE)[,2]

#
loss_lmermodels$poly1<- glmer(gaze_loss_prop ~ ot1 + ot2 + (1 + ot1 + ot2 | ssid),
                             family = "binomial",
                             data = calibration_lmerdata1)

summary(loss_lmermodels$poly1)

loss_lmermodels$poly2<- glmer(gaze_loss_prop ~ tNo+ (1  | ssid),
                             family = "binomial",
                             data = calibration_lmerdata1)

summary(loss_lmermodels$poly2)

calibration_lmerdata$gaze_loss_prop

(calibration_lmerdata1)%>%
  ggplot(aes(tNo, gaze_loss_prop))+
  #geom_smooth(aes(color = ssid), se = F)
  #stat_summary(geom = 'pointrange')
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(loss_lmermodels$poly1), colour = ssid))

# second part

calibration_lmerdata2<- subset(calibration_lmerdata, ssid> 303 & tNo > 25)

calibration_lmerdata2$ot1<- poly(calibration_lmerdata2$tNo, degree = 1, raw = FALSE)
calibration_lmerdata2$ot2<- poly(calibration_lmerdata2$tNo, degree = 2, raw = FALSE)[,2]

# lmer models on loss

calibration_lmerdata2$avg_error_agg_z<- scale(calibration_lmerdata2$avg_error_agg, center = TRUE, scale = TRUE)

loss_lmermodels$poly2.1<- glmer(gaze_loss_prop ~ (ot1 + ot2)+ avg_error_agg_z+ (1 + ot1 + ot2 | ssid),
                             family = "binomial",
                             data = calibration_lmerdata2)

summary(loss_lmermodels$poly2.1)

loss_lmermodels$poly2.2<- glmer(gaze_loss_prop ~ tNo+ (1  | ssid),
                             family = "binomial",
                             data = calibration_lmerdata2)

summary(loss_lmermodels$poly2.2)

calibration_lmerdata$gaze_loss_prop

(calibration_lmerdata2)%>%
  ggplot(aes(tNo, gaze_loss_prop))+
  geom_smooth(aes(color = ssid), se = F)
  #stat_summary(geom = 'pointrange')
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(loss_lmermodels$poly2.1), colour = ssid))

#combine both
calibration_lmerdata3$part <- subset(calibration_lmerdata, ssid> 303 & tNo > 0)

calibration_lmerdata3$ot1<- poly(calibration_lmerdata3$tNo2, degree = 1, raw = FALSE)
calibration_lmerdata3$ot2<- poly(calibration_lmerdata3$tNo2, degree = 2, raw = FALSE)[,2]

#fir full model excluding pratice

calibration_lmerdata3
calibration_lmerdata3$part<- if_else(calibration_lmerdata3$tNo< 25, -.05, .05)

calibration_lmerdata3$ot1_z<- scale(calibration_lmerdata3$ot1, center = TRUE)

calibration_lmerdata3$ot2_z<- scale(calibration_lmerdata3$ot2, center = TRUE)

loss_lmermodels$poly3.1 <- glmer(gaze_loss_prop ~ (ot1_z + ot2_z) * part +(1 + ot1_z + ot2_z | ssid),
                             family = 'quasi_binomial',
                             data = calibration_lmerdata3)

loss_lmermodels$poly3.1 <- lmer(log1p(gaze_loss_prop+.5) ~ (ot1_z + ot2_z) * part +   (0 + ot1_z + ot2_z  | ssid)+(1 | ssid),
                             REML = FALSE,
                             data = calibration_lmerdata3)

loss_lmermodels$poly3.1 <- lmer(log1p(gaze_loss_prop+.5) ~ (ot1 + ot2) * part +   (0 + ot1 + ot2  | ssid)+(1 | ssid),
                             REML = FALSE,
                             data = calibration_lmerdata3)

summary(loss_lmermodels$poly3.1)

anova(loss_lmermodels$poly3.1)
library(sjstats)
library(sjPlot)

tab_model(loss_lmermodels$poly3.1)

relgrad <- with(loss_lmermodels$poly3.1@optinfo$derivs,solve(Hessian,gradient))
 max(abs(relgrad))
 #[1] 1.152891e-05

emmeans::emmeans(loss_lmermodels$poly3.1, pairwise~ part | ot2_z )

loss_lmermodels$poly3.1_rand1 <- lmer(log1p(gaze_loss_prop+.5) ~(1 | ssid),
                             REML = FALSE,
                             data = calibration_lmerdata3)

summary(loss_lmermodels$poly3.1_rand1)

loss_lmermodels$poly3.1_rand2 <- lmer(log1p(gaze_loss_prop+.5) ~( | ssid),
                             REML = FALSE,
                             data = calibration_lmerdata3)

s

# check ID

library(psych)
library(sjstats)
library(psycho)
psycho::analyze(loss_lmermodels$poly3.1)
psych
library(MuMIn)
r.squaredGLMM(loss_lmermodels$poly3.1)
r.squaredGLMM(loss_lmermodels$poly3.2)
#what if we incorporate calibration values

calibration_lmerdata3$avg_error_agg_z <- scale(calibration_lmerdata3$avg_error_agg, center = TRUE, scale = TRUE)

3$avg_error_agg_z <- scale(calibration_lmerdata3$avg_error_agg, center = TRUE, scale = TRUE)

unique(calibration_lmerdata3$ssid)
loss_lmermodels$poly3.2 <- lmer(log1p(gaze_loss_prop+.5) ~ (ot1_z + ot2_z) * part + avg_error_agg_z +   (0 + ot1_z + ot2_z  | ssid)+(1 | ssid),
                             REML = FALSE,
                             data = calibration_lmerdata3)

summary(loss_lmermodels$poly3.2)

anova(loss_lmermodels$poly3.2, loss_lmermodels$poly3.1)

calibration_lmerdata3$tNo2<- if_else(as.double(calibration_lmerdata3$tNo)< 26, as.double(calibration_lmerdata3$tNo), calibration_lmerdata3$tNo-25)

(calibration_lmerdata3)%>%
  ggplot(aes(tNo2, gaze_loss_prop))+
  #geom_smooth(aes(color = ssid), se = F)
  #stat_summary(geom = 'pointrange')
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(loss_lmermodels$poly3.1), colour = ssid))+
    facet_grid(~part)

(calibration_lmerdata3)%>%
  ggplot(aes(tNo2, gaze_loss_prop))+
  geom_line(aes(group = ssid,color = ssid))+
 # geom_smooth(aes(color = ssid), se = F)+
    facet_grid(~part)

# Save an object to a file
saveRDS(calibration_lmerdata3, file = "calibration_lmerdata3.rds")
# Restore the object
readRDS(file = "my_data.rds")

# is calibraton related to loss acc etc
calibration_lmerdata3$avg_error_agg_z<- as.factor(as.character(calibration_lmerdata3$part))
 %>%
  group_by(ssid, part)%>%
  summarise_at(c(avg_error1_visang))

calibration_lmerdata3$avg_error_agg_z+5

tmp.df4_sum_below500_1.2<- tmp.df4_sum_below500 %>%
   subset(cali_order!= 1)%>%
  group_by(ssid, cali_order )%>%
  summarise_at(c('avg_error1','gaze_loss_prop', 'cali_no_resc_agg'), mean, na.rm = TRUE)

 nrow(tmp.df4_sum_below500_1.2)

 #avg error
cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 2)$avg_error1, subset(tmp.df4_sum_below500_1.2, cali_order == 2)$gaze_loss_prop)

cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 3)$avg_error1, subset(tmp.df4_sum_below500_1.2, cali_order == 3)$gaze_loss_prop)

cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 2)$cali_no_resc_agg, subset(tmp.df4_sum_below500_1.2, cali_order == 2)$gaze_loss_prop)

cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 3)$cali_no_resc_agg, subset(tmp.df4_sum_below500_1.2, cali_order == 3)$gaze_loss_prop)

tmp.df4_sum_below500_1.2%>%
ggplot(aes(cali_no_resc_agg, gaze_loss_prop))+
  geom_point()+
geom_smooth(method = 'lm', se = F)+
  facet_grid(~cali_order)

   ggplot(aes(avg_error1, gaze_loss_prop))+
   geom_point()+
   facet_grid(~ cali_order)+
   geom_smooth(method = 'lm', se = F)

```

Accuracy

let's workout a few concepts
, speed velocity and aceleration

Velocity =

Velocity is a physical vector quantity; both magnitude and direction are needed to define it. The scalar absolute value (magnitude) of velocity is called speed, being a coherent derived unit whose quantity is measured in the SI (metric system) as metres per second (m/s) or as the SI base unit of (m???s???1). velocity = delta space/ delta time

Speed =  speed  = distance/ time

acceleration is the rate of change of the velocity of an object with respect to time. Accelerations are vector quantities (in that they have magnitude and direction).[1][2] acceleration = delta velocity/ delta time

For example, EyeLink (SR Research Ltd., Ontario, Canada) uses a velocity threshold of 35 deg/s and an acceleration threshold of 8000 deg/s 2 as default values, although these thresholds can be altered manually

Let's compute accuracy

ACCURACY FINAL CALCULATION
STEP 1: Calculate ABSOLUTE horizontal and vertical distance for EACH SAMPLE
STEP 2: Detect sample outliers using the median + 3 * MAD (rely only on the positive side!)
STEP 3: Remove outliers (horizontal outliers should be removed ONLY when calculating horizontal accuracy and the same for vertical outliers)
STEP 4: Mean horizontal and vertical distance (accuracy) for each participant, condition, trials
STEP 5: Calculate overall accuracy using Pitagorean theorem
STEP 6: Remove outliers in the same way as the previous (see step 2)
STEP 7: Mean accuracy per participant, condition and trial

```{r}

accuracy_dataset$y_coord<- 600
accuracy_dataset$x_coord

unique(accuracy_dataset$screencontent)
unique(accuracy_dataset$distance)
#Calculate accuracy again
library(rCAT)
# remember to ignore the last few seconds later

accuracy_dataset$distance
accuracy_dataset$y_coord
accuracy_dataset$x_coord

setwd("D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Validation")

accuracy_dataset <- readRDS("D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Validation/accuracy_datasetjan16.rds")

# rm(accuracy_datasetjan16)

accuracy_dataset_below6s<- subset(accuracy_dataset, timerezero< 6)

max(accuracy_dataset_below6s$timerezero)
# STEP 1: Calculate ABSOLUTE horizontal and vertical distance for EACH SAMPLE
library(tidyverse)

accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, time) %>%
  group_by(ssid, tNo) %>%
  mutate(v_accuracy_dif = abs(y_coord - gaze_y_cor_pix)) %>%
  # mutate(v_accuracy = mean(v_accuracy_dif, na.rm = TRUE)) %>%
  mutate(h_accuracy_dif = abs(x_coord - gaze_x_cor_pix )) %>%
  # STEP 2: Detect sample outliers using the median + 3 * MAD (rely only on the positive side!)
  mutate(v_accuracy_dif_median = median(v_accuracy_dif, na.rm = TRUE),
                                        h_accuracy_dif_median = median(h_accuracy_dif, na.rm = TRUE),
                                        v_accuracy_dif_mad = mad(v_accuracy_dif, na.rm = TRUE),
                                        h_accuracy_dif_mad = mad(h_accuracy_dif, na.rm = TRUE))%>%
  mutate(v_accuracy_dif_outlier = if_else(v_accuracy_dif > (v_accuracy_dif_median + (3*v_accuracy_dif_mad)), TRUE, FALSE))%>%
  mutate(h_accuracy_dif_outlier = if_else(h_accuracy_dif > (h_accuracy_dif_median + (3*h_accuracy_dif_mad)), TRUE, FALSE))%>%
  mutate(both_acc_dif_outlier = if_else(v_accuracy_dif_outlier == TRUE & h_accuracy_dif_outlier == TRUE, TRUE, FALSE))%>%
  # create new difference columns where we don't have outliers
  mutate(v_accuracy_dif_nooutl = if_else(v_accuracy_dif_outlier == FALSE, v_accuracy_dif, NULL))%>%
  mutate(h_accuracy_dif_nooutl = if_else(h_accuracy_dif_outlier == FALSE, h_accuracy_dif, NULL))

```

```{r}

# STEP 4: Mean horizontal and vertical distance (accuracy) for each participant, condition, trials
library(rCAT)

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  group_by(ssid, tNo) %>%
  mutate(v_accuracy = mean(v_accuracy_dif_nooutl, na.rm = TRUE)) %>%
  mutate(h_accuracy = mean(h_accuracy_dif_nooutl, na.rm = TRUE)) %>%
  mutate(avg_accuracy = (v_accuracy + h_accuracy)/2) %>%
  mutate(v_accuracy_visang = pixtoAng(hw = 32, d = 65, r = 1200, stimsize = v_accuracy)) %>%
  mutate(h_accuracy_visang = pixtoAng(hw = 52, d = 65, r = 1920, stimsize = h_accuracy))

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  group_by(ssid, tNo) %>%
  mutate(avg_accuracy_visang = (v_accuracy_visang + h_accuracy_visang)/2)

accuracy_dataset_below6s %>%
  group_by(ssid,tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(v_accuracy_visang))+
  geom_histogram(bins = 300)+
  geom_vline(xintercept = 1.4, color = 'red')

```

```{r}

# STEP 5: Calculate overall accuracy using Pitagorean theorem
accuracy_dataset_below6s<- accuracy_dataset_below6s%>%
    arrange(ssid, tNo, time) %>%
  group_by(ssid, tNo) %>%
    mutate(accuracy_dif_pt = sqrt((x_coord - gaze_x_cor_pix)^2+ (x_coord - gaze_x_cor_pix)^2)) %>%
   mutate(accuracy_dif_pt_median = median(accuracy_dif_pt, na.rm = TRUE),accuracy_dif_pt_mad = mad(accuracy_dif_pt, na.rm = TRUE))%>%
  # check outliers
   mutate(accuracy_dif_pt_outlier = if_else(accuracy_dif_pt > (accuracy_dif_pt_median + (3*accuracy_dif_pt_mad)), TRUE, FALSE))%>%
   mutate(accuracy_dif_pt_nooutl = if_else(accuracy_dif_pt_outlier == FALSE, accuracy_dif_pt, NULL))

# mean accuracy
accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  group_by(ssid, tNo) %>%
  mutate(accuracy_pt = mean(accuracy_dif_pt_nooutl, na.rm = TRUE)) %>%
  mutate(avg_accuracy_visang_pt = pixtoAng(hw = 52, d = 65, r = 1920, stimsize = accuracy_pt))

# report
mean(accuracy_dataset_below6s$avg_accuracy_visang_pt, na.rm = TRUE) #0.770
sd(accuracy_dataset_below6s$avg_accuracy_visang_pt, na.rm = TRUE) #0.70
median(accuracy_dataset_below6s$avg_accuracy_visang_pt, na.rm = TRUE) # 0.6382099

# mean(accuracy_dataset_below6s$avg_accuracy_visang, na.rm = TRUE)#.93
range(accuracy_dataset_below6s$avg_accuracy_visang_pt, na.rm = TRUE) # 0.2335924 13.5357490
)
# vertical
mean(accuracy_dataset_below6s$v_accuracy_visang, na.rm = TRUE) # 1.327067
median(accuracy_dataset_below6s$v_accuracy_visang, na.rm = TRUE) #1.17
sd(accuracy_dataset_below6s$v_accuracy_visang, na.rm = TRUE) # 0.725

range(accuracy_dataset_below6s$v_accuracy_visang, na.rm = TRUE) # 0.276907 9.500563

# horizontal

mean(accuracy_dataset_below6s$h_accuracy_visang, na.rm = TRUE) #0.5448298
median(accuracy_dataset_below6s$h_accuracy_visang, na.rm = TRUE) #.451
sd(accuracy_dataset_below6s$h_accuracy_visang, na.rm = TRUE) # 0.4966967

range(accuracy_dataset_below6s$accuracy_visang, na.rm = TRUE) #0.3019886 6.3431213

acc_plots<- list()

acc_plots$accuracy<- accuracy_dataset_below6s %>%
  group_by(ssid,tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(avg_accuracy_visang_pt))+
  geom_histogram(bins = 300)+
  geom_vline(xintercept = 1, color = 'red', linetype = "dashed", size =1.5)+p$graphstyle+
  xlab("Accuracy (deg)")+
  ylab("Density")

acc_plots$accuracy<- acc_plots$accuracy+
  scale_x_continuous(limits = c(0,10), breaks = c(0, 5, 10))

acc_plots$accuracy
saveRDS(accuracy_dataset_below6s, file = "accuracy_dataset_below6sJan16.rds")

```

fixation detection

fixations

# to detect fixations and saccades we should
compute velocity and degre of displacement by participant and trial
compute peak velcoicties as well
compute the median and sd by trial and parocipant

saccades will have occured if at least 3 consecutive samples break 6* sd of the median belocity

```{r}

# unique(fixationscreen_only1$screencontent)

# we need to compute velocity so that we can exclude samples that are likelly saccades
# exclude samples where people clearly where not looking at the center
# compute accuracy and precision

View(accuracy_dataset)

#Original calculation with the squaring

accuracy_dataset_below6s

# velocity
unique(accuracy_dataset_below6s$screencontent)
accuracy_dataset_below6s$distance

# max(accuracy_dataset_below6s$timerezero )

accuracy_dataset_below6s <-

  accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo)%>%
  mutate(intdist_pix =  sqrt((gaze_x_cor_pix - lag(gaze_x_cor_pix))^2 +(gaze_y_cor_pix - lag(gaze_y_cor_pix))^2)) %>%
  mutate(inttimediff =  (timerezero)-  lag(timerezero))%>% #If time is calculated this way, it will create a mismatch between inter-sample times and inter-sample distance
  mutate(intdist_visang =  pixtoAng(hw =52, d= 65, r=1920, stimsize= intdist_pix)) %>%
  mutate(int_velocity = intdist_visang/inttimediff)%>%
  mutate(median_velocity = median(int_velocity, na.rm = TRUE,
                                  mad_velocity = mad(int_velocity, na.rm = TRUE)))%>%
  mutate(threshold_velocity = median_velocity+ (6*mad_velocity))%>%
  mutate(peak_velocity = max(int_velocity, na.rm = TRUE))%>%
  mutate(avg_velocity = mean(int_velocity, na.rm = TRUE))

accuracy_dataset_below6s %>%
  group_by(ssid, trialUnq)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(threshold_velocity))+
  geom_histogram(bins = 300)

accuracy_dataset_below6s

accuracy_dataset_below6s %>%
  group_by(ssid, trialUnq)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(peak_velocity, threshold_velocity))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)

```

```{r}

# accuracy_dataset<- accuracy_dataset %>%
#   group_by(ssid, trialUnq)%>%
#   mutate(median_velocity = median(int_velocity, na.rm = TRUE))%>% # I think his needs to be positive(absolute other wise the median will be screwed)
#   mutate(mad_velocity = mad(int_velocity, na.rm = TRUE))%>% 
#   mutate(threshold_velocity = median_velocity + (6* mad_velocity))

# mutate(threshold_velocity = median_velocity + (6* mad_velocity)) %>%

# create a boolean vector for fixationdetcation

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo)%>%
  mutate(detection = if_else(abs(int_velocity)> threshold_velocity, "s", "f"))

accuracy_dataset_below6s$detection

# test<-
# accuracy_dataset_below6s %>%
#   subset(ssid == 309 &tNo == 2)%>%
#   mutate(timesaccade = if_else(detection == "s", timerezero, NULL))

accuracy_dataset_below6s %>%
  subset(ssid == 309 &tNo == 2)%>%
  mutate(timesaccade = if_else(detection == "s", timerezero, NULL))%>%
  ggplot(aes(x = timerezero, y = abs(int_velocity)))+
  geom_line()+
  geom_line(aes(y =gaze_y_cor_pix ), color = "red")+
  geom_line(aes(y =gaze_x_cor_pix ), color = "green")+
  geom_line(aes(y = threshold_velocity) , linetype = "dashed", size = 1, color = "gray")+
    ylim(0, 1000)+
  geom_vline(aes(xintercept = timesaccade), alpha  = .3)

  geom_hline(yintercept = -35, linetype = "dashed", size = 1, color = "gray")
unique(accuracy_dataset_below6s$threshold_velocity)

```

```{r}
# now find a pattern "s - f - s" and if e two saccades slit by only a few fixatiojs (up to 3) then remove the the fixations as they are likelly saccades"
# pattern 1 "s-f-s"
table(test2$detection)
accuracy_dataset_below6s$new<-NULL
accuracy_dataset_below6s$new = ifelse(accuracy_dataset_below6s$detection == "f" & 
                            lag(accuracy_dataset_below6s$detection, n = 1) == "s" & 
                                lead(accuracy_dataset_below6s$detection, n = 1) == "s", TRUE, FALSE)

r = c("s", "f", "s")
library(zoo)
test<- which(rollapply(accuracy_dataset_below6s$detection,length(r),identical,r))+0:(length(r)-1)
View(test)
test
# 
# test
test2<- accuracy_dataset_below6s[12:20,]

View(test2)

# pattern 2 "s-f-f-s"
accuracy_dataset_below6s$new1<-NULL

accuracy_dataset_below6s <- accuracy_dataset_below6s %>% 
 mutate(new1 = ifelse(detection == "f" & 
                       lead(detection, n = 1) == "f" &
                            lag(detection, n = 1) == "s" & 
                                lead(detection, n = 2) == "s", TRUE, FALSE))

r1 = c("s", "f", "f", "s")
test3<- which(rollapply(accuracy_dataset_below6s$detection,length(r1),identical,r1))+0:(length(r1)-1)

test3

test4<- accuracy_dataset_below6s[304514 :304530,]
View(test4)

# accuracy_dataset_below6s$new1[accuracy_dataset_below6s$new1 == TRUE

# if_else(accuracy_dataset_below6s$new1 == TRUE, row.names(accuracy_dataset_below6s))

View(test2)
# 
test3<- subset(accuracy_dataset_below6s[290310:290328,])

# # pattern 3 "s-f-f-"f" s"

accuracy_dataset_below6s$new2<-NULL

accuracy_dataset_below6s<- accuracy_dataset_below6s %>% 
 mutate(new2 = ifelse(detection == "f" & 
                        lead(detection, n = 1) == "f" &
                        lead(detection, n = 2) == "f" &
                        lead(detection, n = 3) == "s" &
                            lag(detection, n = 1) == "s", TRUE, FALSE))

# r1 = c("s", "f", "f", "s")

# r2 = c("s", "f", "f", "f", "s")
# 
# test1.1 <- which(rollapply(accuracy_dataset_below6s$detection,length(r1),identical,r1))+0:(length(r1)-1)

accuracy_dataset_below6s[which(accuracy_dataset_below6s$new1 == TRUE)+1, 'new1'] <- TRUE

accuracy_dataset_below6s[which(accuracy_dataset_below6s$new2 == TRUE)+1, 'new2'] <- TRUE
accuracy_dataset_below6s[which(accuracy_dataset_below6s$new2 == TRUE)+1, 'new2'] <- TRUE

# accuracy_dataset_below6s$neW2<- NULL

#mark fixations that need to be changed to saccades
accuracy_dataset_below6s$fixation_tosacade<- (rowSums(accuracy_dataset_below6s[,startsWith(names(accuracy_dataset_below6s),"new")]=="TRUE") >= 1)

table(accuracy_dataset_below6s$fixation_tosacade)

# unique(subset(accuracy_dataset_below6s, fixation_tosacade == TRUE )$new)

# replace spurious fixations when a sacade is taling pplace

accuracy_dataset_below6s<- accuracy_dataset_below6s%>%
  mutate(detection1 = if_else(fixation_tosacade == TRUE, "s", detection))

# accuracy_dataset_below6s$sac_time <- if_else(accuracy_dataset_below6s$detection1 == "s", accuracy_dataset_below6s$timerezero, NULL)
# accuracy_dataset_below6s$sac_time <- NULL

# delete spurious saccades

# create start of a saccade

accuracy_dataset_below6s$start_sacc<- NULL
accuracy_dataset_below6s$end_sacc<- NULL

# accuracy_dataset_below6s$start_sacc <- if_else(is.na(accuracy_dataset_below6s$sac_time) == FALSE & is.na(lag(accuracy_dataset_below6s$sac_time, n = 1) == TRUE),accuracy_dataset_below6s$sac_time, NULL)

# accuracy_dataset_below6s$end_sacc <- if_else(is.na(accuracy_dataset_below6s$sac_time) == FALSE & is.na(lead(accuracy_dataset_below6s$sac_time, n = 1) == TRUE),accuracy_dataset_below6s$sac_time, NULL)
# 
# max(accuracy_dataset_below6s$sac_time, na.rm = TRUE)

# #get rid os start of old spurios saccades
# accuracy_dataset_below6s$start_sacc <- if_else(accuracy_dataset_below6s$sacc_dur > 0, accuracy_dataset_below6s$start_sacc, NULL)
# accuracy_dataset_below6s$end_sacc <- if_else(accuracy_dataset_below6s$sacc_dur > 0, accuracy_dataset_below6s$end_sacc, NULL)

# accuracy_dataset_below6s <- accuracy_dataset_below6s%>%
#   arrange(ssid, tNo, timerezero)%>%
#   group_by(ssid, tNo)%>%
#   fill(end_sacc, .direction = "up")%>%
#   fill(start_sacc, .direction = "updown")%>%
#   mutate(sacc_dur = end_sacc - start_sacc)%>%
#   mutate(detection_withw_spursacc = if_else(sacc_dur == 0, "f", detection1))
#     
accuracy_dataset_below6s $detection_withw_spursacc<- NULL

accuracy_dataset_below6s %>%
subset(ssid == 309 & tNo == 2)%>%
  ggplot(aes(timerezero, abs(int_velocity)))+
  geom_line()+
  geom_point(aes(y = gaze_y_cor_pix))+
  geom_point(aes(y = gaze_x_cor_pix))+
  geom_line(aes(y = threshold_velocity), size = 1.5)
  # geom_vline(aes(xintercept = start_sacc), color = 'green')
  geom_vline(aes(xintercept = end_sacc), color = 'red')

  # annotate("rect", xmin = start_sacc, xmax = end_sacc)

# delete single sacades
rle(accuracy_dataset_below6s$detection1)

accuracy_dataset_below6s$saccade_less3 <- rle(accuracy_dataset_below6s$detection1)

accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
  group_by(ssid, tNo, detection1, event_order) %>%
  mutate(keepF = if_else(detection1 == 'f' & n() >= 12, TRUE, FALSE)) %>%
  mutate(keepS = if_else(detection1 == 's' & n() >= 3, TRUE, FALSE)) %>%

# create a conter based on saca start

saveRDS(accuracy_dataset_below6s, file = "accuracy_dataset_below6sjan16.rds")

testsacc<- accuracy_dataset_below6s%>%
subset(ssid == 309 & tNo == 2)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

  # geom_vline(aes(xintercept = start_sacc), color = "red", size = 2, alpha = .03) +
  # geom_vline(aes(xintercept = end_sacc), color = "green", linetype = "dotted", alpha = .03) +
  ylim(0,1000)

mean(accuracy_dataset_below6s$end_sacc - accuracy_dataset_below6s$start_sacc, na.rm = TRUE)
min(accuracy_dataset_below6s$end_sacc - accuracy_dataset_below6s$start_sacc, na.rm = TRUE)
max(accuracy_dataset_below6s$end_sacc - accuracy_dataset_below6s$start_sacc, na.rm = TRUE)

accuracy_dataset_below6s %>%
  subset(ssid == 309 & tNo == 2)%>%
  ggplot(aes(timerezero, abs(int_velocity)))+
  geom_line()+
  geom_line(aes(y = intdist_visang_vert))+
  geom_point(aes(y = gaze_y_cor_pix))+
  # geom_point(aes(y = gaze_x_cor_pix))+
  # geom_line(aes(y = threshold_velocity), size = 1.5)+
  # geom_vline(aes(xintercept = sac_time), alpha = .03)+

  geom_vline(aes(xintercept = start_sacc), color = "green", size = 2, alpha = .3)+
  geom_vline(aes(xintercept = end_sacc), color = "red", size = 2, linetype = "dotted", alpha = .3) 
  ylim(0,100)+
  xlim(.6,2)

max(accuracy_dataset_below6s$timerezero3)

min(accuracy_dataset_below6s$timerezero)
```

Timecourse models for accuracy

```{r}

unique(accuracy_dataset_below6s$screencontent)
unique(accuracy_dataset_below6s$tNo)
unique(accuracy_dataset_below6s$ssid)

fixationscreen_only2<- accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

fixationscreen_only2$ssid<- as.numeric(as.character(fixationscreen_only2$ssid))

# fixationscreen_only2$ssid< 500
# fixationscreen_only2_lmer <- fixationscreen_only2  %>%
#   subset(tNo>0 & as.numeric(ssid, na.rm = TRUE) < 500)

fixationscreen_only2_lmer<- fixationscreen_only2
fixationscreen_only2_lmer$tNo2<- if_else(as.double(fixationscreen_only2_lmer$tNo)< 25, as.double(fixationscreen_only2_lmer$tNo), fixationscreen_only2_lmer$tNo - 24)

unique(fixationscreen_only2_lmer$tNo2)

#polynomial
fixationscreen_only2_lmer$ot1<- poly(fixationscreen_only2_lmer$tNo2, degree = 1, raw =FALSE)[,1]
fixationscreen_only2_lmer$ot2<- poly(fixationscreen_only2_lmer$tNo2, degree = 2, raw =FALSE)[,2]
fixationscreen_only2_lmer$ot1_z<- scale(fixationscreen_only2_lmer$ot1, center = TRUE, scale = TRUE)[,1]
fixationscreen_only2_lmer$ot2_z<- scale(fixationscreen_only2_lmer$ot2, center = TRUE, scale = TRUE)[,1]

fixationscreen_only2_lmer%>%
  ggplot(aes(tNo2, ot1_z))+
  geom_line()+
  geom_line(aes(y = ot2_z))

fixationscreen_only2_lmer$block<- if_else(fixationscreen_only2_lmer$tNo< 25, -.5, +.5)

fixationscreen_only2_lmer$subject<- as.factor(as.character(fixationscreen_only2_lmer$ssid))                                          
# models

accurcay_models$acc_1<- lmer(avg_accuracy_visang_pt ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | subject), REML = FALSE,
                             data = fixationscreen_only2_lmer)

summary(accurcay_models$acc_1)

anova(accurcay_models$acc_1)

```
Fixed effects:
               Estimate  Std. Error          df t value Pr(>|t|)   
ot1_z          0.044002    0.013721 2064.033249   3.207  0.00136 **
ot2_z         -0.001381    0.013899 2064.058090  -0.099  0.92087   
block          0.068205    0.027205 2064.999334   2.507  0.01225 * 
ot1_z:block    0.034924    0.027445 2064.297130   1.273  0.20333   
ot2_z:block    0.036084    0.027793 2064.388596   1.298  0.19432   
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Correlation of Fixed Effects:
            ot1_z  ot2_z  block  ot1_z:
ot2_z        0.054                     
block       -0.078 -0.090              
ot1_z:block -0.123 -0.162  0.029       
ot2_z:block -0.162 -0.194  0.028  0.054

```{r}

unique(fixationscreen_only2_lmer$ssid)

plot(accurcay_models$acc_1)
qqnorm(resid(accurcay_models$acc_1)) 
qqline(resid(accurcay_models$acc_1))

fixationscreen_only2_lmer$subject<- as.factor(as.character(fixationscreen_only2_lmer$ssid))

accurcay_models$acc_1_notstand <- lmer(accuracy_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | subject), REML = FALSE,
                             data = fixationscreen_only2_lmer)

summary(accurcay_models$acc_1_notstand)
# uniq

(fixationscreen_only2_lmer) %>%
  ggplot(aes(tNo2, accuracy_visang))+
  # geom_smooth(aes(color = ssid), se = F)
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(accurcay_models$acc_1, fixationscreen_only2_lmer), colour = subject))+
  facet_grid(~block)

```

```{r}

# horizontal
accurcay_models$acc_1_h <- lmer(h_accuracy_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ 
                                (1 | ssid), REML = FALSE,
                             data = fixationscreen_only2_lmer)

plot(accurcay_models$acc_1_h)
qqnorm(resid(accurcay_models$acc_1_h))
qqline(resid(accurcay_models$acc_1_h))

summary(accurcay_models$acc_1_h)

```
Linear mixed model fit by maximum likelihood . t-tests use Satterthwaite's method ['lmerModLmerTest']
Formula: h_accuracy_visang ~ (ot1_z + ot2_z) * block + (0 + ot1_z + ot2_z) +      (1 | ssid)
   Data: fixationscreen_only2_lmer

     AIC      BIC   logLik deviance df.resid 
  2686.4   2726.0  -1336.2   2672.4     2100

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-2.2019 -0.2687 -0.0533  0.1650 19.9560

Random effects:
 Groups   Name        Variance Std.Dev.
 ssid     (Intercept) 0.3487   0.5905  
 Residual             0.1899   0.4357  
Number of obs: 2107, groups:  ssid, 43

Fixed effects:
                Estimate   Std. Error           df t value Pr(>|t|)   
ot1_z          0.0311142    0.0097019 2064.0332461   3.207  0.00136 **
ot2_z         -0.0009764    0.0098278 2064.0580865  -0.099  0.92087   
block          0.0482285    0.0192370 2064.9993309   2.507  0.01225 * 
ot1_z:block    0.0246953    0.0194067 2064.2971262   1.273  0.20333   
ot2_z:block    0.0255152    0.0196523 2064.3885924   1.298  0.19432   
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Correlation of Fixed Effects:
            ot1_z  ot2_z  block  ot1_z:
ot2_z        0.054                     
block       -0.078 -0.090              
ot1_z:block -0.123 -0.162  0.029       
ot2_z:block -0.162 -0.194  0.028  0.054

# vertical
```{r}
accurcay_models$acc_1_v <- lmer(v_accuracy_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ 
                                (1 | ssid), REML = FALSE,
                             data = fixationscreen_only2_lmer)

plot(accurcay_models$acc_1_v)
qqnorm(resid(accurcay_models$acc_1_v))
qqline(resid(accurcay_models$acc_1_v))

emmeans::emmeans(accurcay_models$acc_1_v, pairwise ~ block|ot1_z )

summary(accurcay_models$acc_1_v)

(fixationscreen_only2_lmer) %>%
  ggplot(aes(tNo2, v_accuracy_visang))+
  # geom_smooth(aes(color = ssid), se = F)
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(accurcay_models$acc_1_v, fixationscreen_only2_lmer), colour = subject))+
  facet_grid(~block)

```

Linear mixed model fit by maximum likelihood . t-tests use Satterthwaite's method ['lmerModLmerTest']
Formula: v_accuracy_visang ~ (ot1_z + ot2_z) * block + (0 + ot1_z + ot2_z) +      (1 | ssid)
   Data: fixationscreen_only2_lmer

     AIC      BIC   logLik deviance df.resid 
  2795.3   2834.9  -1390.6   2781.3     2100

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-4.1827 -0.4861 -0.0329  0.4371 16.6153

Random effects:
 Groups   Name        Variance Std.Dev.
 ssid     (Intercept) 2.0756   1.4407  
 Residual             0.1929   0.4392  
Number of obs: 2107, groups:  ssid, 43

Fixed effects:
               Estimate  Std. Error          df t value Pr(>|t|)   
ot1_z          0.001718    0.009779 2064.007298   0.176  0.86056   
ot2_z          0.002415    0.009906 2064.011559   0.244  0.80743   
block          0.062299    0.019392 2064.177075   3.213  0.00134 **
ot1_z:block    0.054673    0.019561 2064.052790   2.795  0.00524 **
ot2_z:block    0.013280    0.019809 2064.068659   0.670  0.50267   
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Correlation of Fixed Effects:
            ot1_z  ot2_z  block  ot1_z:
ot2_z        0.054                     
block       -0.078 -0.090              
ot1_z:block -0.123 -0.162  0.029       
ot2_z:block -0.162 -0.194  0.028  0.054

$emmeans
ot1_z = -0.00209:
 block  emmean      SE   df lower.CL upper.CL
  -0.5 -0.0311 0.00971 2069  -0.0501  -0.0121
   0.5  0.0311 0.00971 2069   0.0121   0.0502

Degrees-of-freedom method: kenward-roger 
Confidence level used: 0.95

$contrasts
ot1_z = -0.00209:
 contrast   estimate     SE   df t.ratio p.value
 -0.5 - 0.5  -0.0622 0.0194 2069 -3.205  0.0014

Degrees-of-freedom method: kenward-roger

```{r}

# 
# 
# mean(fixationscreen_only2_lmer$accuracy_visang, na.rm = TRUE)
# 
# median(fixationscreen_only2_lmer$accuracy_visang, na.rm = TRUE)
# 
# sd(fixationscreen_only2_lmer$accuracy_visang, na.rm = TRUE)
# min(fixationscreen_only2_lmer$accuracy_visang, na.rm = TRUE)
# max(fixationscreen_only2_lmer$accuracy_visang, na.rm = TRUE)
# 
# 
# 
# mean(fixationscreen_only2_lmer$v_accuracy_visang, na.rm = TRUE)
# 
# median(fixationscreen_only2_lmer$v_accuracy_visang, na.rm = TRUE)
# 
# sd(fixationscreen_only2_lmer$v_accuracy_visang, na.rm = TRUE)
# min(fixationscreen_only2_lmer$v_accuracy_visang, na.rm = TRUE)
# max(fixationscreen_only2_lmer$v_accuracy_visang, na.rm = TRUE)
# 
# 
# mean(fixationscreen_only2_lmer$h_accuracy_visang, na.rm = TRUE)
# 
# median(fixationscreen_only2_lmer$h_accuracy_visang, na.rm = TRUE)
# 
# sd(fixationscreen_only2_lmer$h_accuracy_visang, na.rm = TRUE)
# min(fixationscreen_only2_lmer$h_accuracy_visang, na.rm = TRUE)
# max(fixationscreen_only2_lmer$h_accuracy_visang, na.rm = TRUE)

# Save an object to a file
saveRDS(fixationscreen_only2_lmer, file = "fixationscreen_only2_lmer.rds")
# Restore the object
readRDS(file = "my_data.rds")

fixationscreen_only2_lmer

 tmp.df4_sum_below500 %>%
   subset(cali_order!= 1)%>%
  group_by(ssid, cali_order )%>%
  summarise_at(c('avg_error1','gaze_loss_prop', 'cali_no_resc_agg'), mean, na.rm = TRUE)%>%
   ggplot(aes(avg_error1, gaze_loss_prop))+
   geom_point()+
   facet_grid(~ cali_order)+
   geom_smooth(method = 'lm', se = F)

 tmp.df4_sum_full_stim_gp_leda_beh_self_biopac %>%
   group_by(ssid, screencontent)%>%
   summarise_if(is.numeric, mean, na.rm = TRUE)%>%
   ggplot(aes(Global.Mean, BIO_Global.Mean))+
   geom_point()+
   geom_smooth(method = 'lm', se = F)+ facet_grid(~screencontent)

```

Precision
fixationscreen_only3
model_time_prec

start by finishing fixation

```{r eval=FALSE, include=FALSE}

table(accuracy_dataset_below6s$detection_withw_spursacc)
saveRDS(accuracy_dataset_below6s, file = "accuracy_dataset_below6sjan16.rds")

# not yet
accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  mutate(gaze_x_cor_pix_forpreci = if_else(detection1 == "f", gaze_x_cor_pix, NULL))%>%
  mutate(gaze_y_cor_pix_forpreci = if_else(detection1 == "f", gaze_y_cor_pix, NULL))
  # mutate(test = if_else(detection1 == "s", timerezero, NULL))
  # mutate(test2 = if_else(detection_withw_spursacc == "s", timerezero, NULL))

# we need to create a counter for the events so that each fix sacc or loss has a duration

# accuracy_dataset_below6s$test <- duplicated(accuracy_dataset_below6s$detection1)
# accuracy_dataset_below6s$test2 <- accuracy_dataset_below6s$detection1

library(data.table)

accuracy_dataset_below6s<-
accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo)%>%
  mutate(event_order = rleid(detection1))%>%
  ungroup()%>%
  group_by(ssid, tNo, detection1, event_order)%>%
  mutate(duplicated_event = !duplicated(event_order))%>%
  ungroup()%>%
  group_by(ssid, tNo)%>%

  mutate(duplicated_event2 = duplicated(event_order))
View(accuracy_dataset_below6s)

# mutate(start = if_else(duplicated_event == TRUE, timerezero, NULL))%>%
#   # fill(start, .direction = "down") %>%
#   mutate(end = if_else(duplicated_event2 == TRUE & lead(duplicated_event2, n=1) == FALSE, timerezero, 
#                if_else(duplicated_event2 == FALSE & lead(duplicated_event2, n=1) == FALSE, start + 0.005, NULL)))%>%
#   ungroup() %>%
#   group_by(ssid, tNo, detection1, event_order) %>%
#   # mutate(end = if_else(duplicated_event == TRUE, timerezero, NULL))%>%
#   fill(start, .direction = "down")%>%
#   fill(end, .direction = "up")%>%
#   mutate(duration = end - start)

# arrangind the ds in order
accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
   arrange(ssid, tNo, timerezero)

accuracy_dataset_below6s$start<- if_else(accuracy_dataset_below6s$duplicated_event == TRUE, accuracy_dataset_below6s$timerezero, NULL)
unique(accuracy_dataset_below6s$start)
table(is.na(accuracy_dataset_below6s$start))

# create end
accuracy_dataset_below6s$end <- 
  if_else(accuracy_dataset_below6s$duplicated_event2 == TRUE & rownames(accuracy_dataset_below6s) == nrow(accuracy_dataset_below6s), accuracy_dataset_below6s$timerezero, 
  if_else(accuracy_dataset_below6s$duplicated_event2 == FALSE & rownames(accuracy_dataset_below6s) == nrow(accuracy_dataset_below6s), accuracy_dataset_below6s$start + 0.005,
  if_else(accuracy_dataset_below6s$duplicated_event2 == TRUE & lead(accuracy_dataset_below6s$duplicated_event2, n = 1) == FALSE, accuracy_dataset_below6s$timerezero,
  if_else(accuracy_dataset_below6s$duplicated_event2 == FALSE & lead(accuracy_dataset_below6s$duplicated_event2, n=1) == FALSE, accuracy_dataset_below6s$start + 0.005, NULL))))

table(is.na(accuracy_dataset_below6s$end))

  # if_else(accuracy_dataset_below6s$duplicated_event2 == TRUE & rownames(accuracy_dataset_below6s) == nrow(accuracy_dataset_below6s), accuracy_dataset_below6s$timerezero, 
  # if_else(accuracy_dataset_below6s$duplicated_event2 == FALSE & rownames(accuracy_dataset_below6s) == nrow(accuracy_dataset_below6s), accuracy_dataset_below6s$start + 0.005, NULL))))

# TEST <- rownames(accuracy_dataset_below6s) == nrow(accuracy_dataset_below6s)
# table(TEST)
# 
# part344 <- subset(accuracy_dataset_below6s, ssid == 344 & tNo == 50 & event_order == 48)
# 
# part344$test <- if_else(rownames(part344) == nrow(part344), TRUE, FALSE)
# 
# part344$test2 <- if_else(part344$duplicated_event2 == TRUE & rownames(part344) == nrow(part344), part344$timerezero, 
#                   if_else(part344$duplicated_event2 == FALSE & rownames(part344) == nrow(part344), part344$start + 0.005, NULL))

table(part344$test)
View(part344)

accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
  arrange(ssid,tNo,timerezero) %>%
  group_by(ssid, tNo, detection1, event_order) %>%
  fill(start, .direction = "down")%>%
  fill(end, .direction = "up")%>%
  mutate(duration = end - start)

unique(accuracy_dataset_below6s$duplicated_event2)

table(is.na(accuracy_dataset_below6s$duration))
table(is.na(accuracy_dataset_below6s$start))
table(is.na(accuracy_dataset_below6s$end))

View(accuracy_dataset_below6s)

max(accuracy_dataset_below6s$duration)
min(accuracy_dataset_below6s$duration)

# natest<- subset(accuracy_dataset_below6s, is.na(duration) == TRUE)
# accuracy_dataset_below6s %>%
#   arrange(ssid, tNo, timerezero)%>%
#   # group_by(ssid, tNo) %>%
#   group_by(ssid, tNo, detection1)%>%
#   fill(end, .direction = "up")

  # mutate(end = if_else(lag(duplicated_event, n = 1) == TRUE, timerezero, NULL))

# accuracy_dataset_below6s%>%
#   ggplot(aes(duration))+
#   geom_histogram(bins = 300)+
#   xlim(0,.01)+
#   facet_grid(~detection1)+
#   geom_vline(xintercept = .005, size = 1, linetype = "dashed", alpha = .5)

# now lets cmpute the average duration of events
saveRDS(accuracy_dataset_below6s, "accuracy_dataset_below6sjan16withevents.rds")

accuracy_dataset_below6s %>%
  group_by(ssid, tNo, detection1)%>%
  summarise_at(c('duration'), mean, na.rm = TRUE)%>%
  ggplot(aes(duration))+
  geom_density(aes(fill = detection1), alpha = .5)
  facet_grid(~detection1)

  table(accuracy_dataset_below6s$detection1 == "f")

  table(is.na(accuracy_dataset_below6s$gaze_x_cor_pix), is.na(accuracy_dataset_below6s$gaze_y_cor_pix))
  unique(accuracy_dataset_below6s$detection1)

  # table(is.na(accuracy_dataset_below6s$detection1))

# eliminate implausbe fixations (< 80ms = 0.08) and saccades( < 12ms = .012)

  saveRDS(  accuracy_dataset_below6s, "  accuracy_dataset_below6sjan16withevents.rds")

accuracy_dataset_below6s<-  readRDS(file = "  accuracy_dataset_below6sjan16withevents.rds")

   accuracy_dataset_below6s$detection2<- if_else(is.na(accuracy_dataset_below6s$detection1) == TRUE,
                                                 "loss",  accuracy_dataset_below6s$detection1)

   table( accuracy_dataset_below6s$detection2)

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  group_by(ssid, tNo, detection2, event_order)%>%
    mutate(fix_validity = if_else(detection2 == "f" & duration > .08, TRUE, FALSE))%>%
    mutate(sacc_validity = if_else(detection2 == "s", duration > .012, TRUE,
                                        FALSE))%>%
    mutate(new_dur = if_else(fix_validity == TRUE & detection2 == "f", duration,
                             if_else(sacc_validity == TRUE & detection2 == "s", duration,
                                    if_else(detection2 == "loss", duration,  NULL))))

# nee detectin ignoring small fix or saccades
accuracy_dataset_below6s$detection3<- if_else(accuracy_dataset_below6s$fix_validity == TRUE & accuracy_dataset_below6s$detection2 == "f", accuracy_dataset_below6s$detection2,
                                              if_else(accuracy_dataset_below6s$sacc_validity == TRUE & 
                                                        accuracy_dataset_below6s$detection2 == "s", accuracy_dataset_below6s$detection2, "loss"))

# this shousld result in less fixation and less saccades

table(accuracy_dataset_below6s$detection1)
table(accuracy_dataset_below6s$detection2)

table(accuracy_dataset_below6s$detection3)

# new event 3 that matches the new detection since the old fixations that were deleted will cause the order not to match

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo,)%>%
  mutate(event_order_new = rleid(detection3))

unique(accuracy_dataset_below6s$event_order_new)

```

  # ungroup()%>%
  # group_by(ssid, tNo, detection1, event_order)%>%
  # mutate(duplicated_event = !duplicated(event_order))%>%
  # ungroup()%>%
  # group_by(ssid, tNo)%>%

```{r}

# compute amplitute of sacadesaccade_less3
accuracy_dataset_below6s$start

# create a new event order for this new detection

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  arrange(ssid, tNo,timerezero)%>%
  group_by(ssid, tNo,detection3, event_order_new)%>%
  mutate(sacc_posstart_x = if_else(detection3 == "s" & start == timerezero, gaze_x_cor_pix, NULL))%>%
 mutate(sacc_posend_x = if_else(detection3 == "s" & end == timerezero, gaze_x_cor_pix, NULL))%>%
    mutate(sacc_posstart_y = if_else(detection3 == "s" & start == timerezero, gaze_y_cor_pix, NULL))%>%
 mutate(sacc_posend_y = if_else(detection3 == "s" & end == timerezero, gaze_y_cor_pix, NULL))

accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
   arrange(ssid, tNo,timerezero)%>%
  group_by(ssid, tNo,detection3, event_order)%>%
 fill(sacc_posstart_x, .direction = "down")%>%
  fill(sacc_posend_x, .direction = "up")%>%
   fill(sacc_posstart_y, .direction = "down")%>%
  fill(sacc_posend_y, .direction = "up")

accuracy_dataset_below6s %>%
  subset(detection3 == "s")%>%
  arrange(ssid, tNo, timerezero)

  saveRDS(accuracy_dataset_below6s, "accuracy_dataset_below6sjan16witheventdur.rds")

# compute saccade amplitude displacement in pitagoean
accuracy_dataset_below6s<- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo, detection3, event_order_new)%>%
  mutate(sac_amp = sqrt((sacc_posstart_x - sacc_posend_x)^2 +(sacc_posstart_y - sacc_posend_y)^2))%>%
  mutate(sac_amp_visang = pixtoAng(hw = 52, d = 65, r = 1920, sac_amp))%>%
  mutate(sac_peak = mean(int_velocity))

#backup
  saveRDS(accuracy_dataset_below6s, "accuracy_dataset_below6sjan16witheventdur.rds")

accuracy_dataset_below6s %>%
  subset(detection3 == "s")%>%
   group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(sac_amp_visang, sac_peak))+
  geom_point()+
  geom_smooth(aes(group = ssid),method =  'lm', se = F)
  facet_grid(~detection3)

  accuracy_dataset_below6s %>%
  subset(detection3 == "s")%>%
   group_by(tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(sac_amp_visang, new_dur))+
  geom_point()+
  geom_smooth(,method =  'lm', se = F)
  facet_grid(~detection3)

accuracy_dataset_below6s$Event<- if_else(accuracy_dataset_below6s$detection2 == "f", "Fixation",
                                         if_else(accuracy_dataset_below6s$detection2 == "s", "Saccade", "Loss"))

acc_plots$parsing<- accuracy_dataset_below6s%>%  
group_by(ssid, tNo,detection2, Event)%>%
    summarise_at(c('new_dur'), max, na.rm = TRUE)%>%
  ggplot(aes(new_dur, fill = Event))+
  geom_histogram(aes(y=..density..),bins = 300, alpha=0.5, 
                position="identity")+
  geom_density(alpha = .2)+
  # geom_vline(xintercept = .1, linetype = "dashed")+
  p$graphstyle+
  xlab("Duration (s)")+
  ylab("Density")+
  scale_color_brewer(palette = "Dark2")+

  # summarise_at(c('Duration (s)')*1000, min, na.rm = TRUE)+
    theme(legend.position = c(0.6, 0.6))

 library(patchwork)

accuracy_dataset_below6s

acc_plots$loss<-  accuracy_dataset_below6s%>%  
group_by(ssid, tNo,detection2, Event)%>%
    summarise_at(c('gaze_loss_prop'), max, na.rm = TRUE)%>%
  ggplot(aes(gaze_loss_prop))+
  geom_histogram(aes(y=..density..),bins = 300, 
                position="identity")+
  # geom_density(alpha = .2)+
  # geom_vline(xintercept = .1, linetype = "dashed")+
  p$graphstyle+
  xlab("Loss")+
  ylab("Density")+
  scale_color_brewer(palette = "Dark2")

acc_plots$loss

acc_plots$loss<- acc_plots$loss+
  scale_x_continuous(limits=c(0,1), breaks = c(0,.5,1))

scale_y_continuous(limits=c(0,40), breaks = c(0,20,40))

acc_plots$loss

  # summarise_at(c('Duration (s)')*1000, min, na.rm = TRUE)+
    theme(legend.position = c(0.6, 0.6))

    library(patchwork)

acc_patch<- (acc_plots$accuracy | acc_plots$precision+theme(axis.title.y = element_blank())) /(acc_plots$loss|acc_plots$parsing+theme(axis.title.y = element_blank()))

acc_patch<- acc_patch+plot_annotation(tag_levels = 'A')
acc_patch

ggsave("acc_patch.tiff", acc_patch, device = "tiff", dpi = 800, height = 7, width = 10)  
#

```

saveRDS(accuracy_dataset_below6s, "accuracy_dataset_below6s.Rds")

```{r}

View(accuracy_dataset_below6s)

accuracy_dataset_below6s$timerezero

#store gazepoints for precision calculations

accuracy_dataset_below6s$gaze_x_cor_pix_for_prec<- if_else(accuracy_dataset_below6s$detection3 == "f", accuracy_dataset_below6s$gaze_x_cor_pix, NULL)

accuracy_dataset_below6s$gaze_y_cor_pix_for_prec<- if_else(accuracy_dataset_below6s$detection3 == "f", accuracy_dataset_below6s$gaze_y_cor_pix, NULL)

# check how parising is happening
 accuracy_dataset_below6s %>%
   subset(ssid == 304 & tNo == 16 )%>%
   ggplot(aes(timerezero, gaze_x_cor_pix_for_prec))+
   geom_line()+
   geom_line(aes(y = gaze_x_cor_pix), color = "green", linetype = "dotted")+
   geom_line(aes(y = int_velocity), color = 'red', alpha = .5)+
   geom_line(aes(y = gaze_y_cor_pix_for_prec), color = "blue")+
    geom_line(aes(y=gaze_y_cor_pix), color = "blue", linetype = "dotted")
   # geom_hline(aes(yintercept = threshold_velocity))+
   geom_vline(aes(xintercept = test))

```

Now compute precision
```{r}
 accuracy_dataset_below6s <- accuracy_dataset_below6s %>%
  arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo) %>%
  mutate(gaze_x_cor_pix_for_prec_lead = lead(gaze_x_cor_pix_for_prec, n = 1)) %>%
  mutate(gaze_x_cor_pix_for_prec_dif = gaze_x_cor_pix_for_prec_lead  - gaze_x_cor_pix_for_prec)%>%
  mutate(gaze_y_cor_pix_for_prec_lead= lead(gaze_y_cor_pix_for_prec, n = 1)) %>%
  mutate(gaze_y_cor_pix_for_prec_dif = gaze_y_cor_pix_for_prec_lead-gaze_y_cor_pix_for_prec) %>%
  mutate(v_precision_pix = sqrt(mean(gaze_y_cor_pix_for_prec_dif ^2, na.rm = TRUE))) %>%
  mutate(h_precision_pix = sqrt(mean(gaze_x_cor_pix_for_prec_dif ^2, na.rm = TRUE))) %>%

  mutate(gaze_cor_pix_euclidean_int_dif = sqrt((gaze_x_cor_pix_for_prec_dif)^2 +(gaze_y_cor_pix_for_prec_dif)^2))%>%
  mutate(precision_eucl_pix = sqrt(mean(gaze_cor_pix_euclidean_int_dif ^2, na.rm = TRUE)))%>%
  mutate(precision_avg_pix = (v_precision_pix + h_precision_pix)/2)%>%

  mutate(sd_h_precision_pix = sd(gaze_x_cor_pix_for_prec, na.rm = TRUE)) %>%
    mutate(sd_v_precision_pix = sd(gaze_y_cor_pix_for_prec, na.rm = TRUE)) %>%
    mutate(sd_precision_avg_pix = (sd_v_precision_pix + sd_h_precision_pix) / 2)

mean(accuracy_dataset_below6s$v_precision_pix, na.rm = TRUE) 
mean(accuracy_dataset_below6s$h_precision_pix, na.rm = TRUE)

mean(accuracy_dataset_below6s$precision_avg_pix, na.rm = TRUE) 
mean(accuracy_dataset_below6s$precision_eucl_pix, na.rm = TRUE)

mean(accuracy_dataset_below6s$sd_h_precision_pix, na.rm = TRUE) 
mean(accuracy_dataset_below6s$sd_v_precision_pix, na.rm = TRUE)

mean(accuracy_dataset_below6s$sd_h_precision_pix, na.rm = TRUE)

accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(precision_eucl_pix))+
  geom_histogram()+
  geom_vline(xintercept = 40)

accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(precision_avg_pix))+
  geom_histogram()+
  geom_vline(xintercept = 40)

accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(sd_precision_avg_pix))+
  geom_histogram()+
  geom_vline(xintercept = 40)

# conversions to degres

accuracy_dataset_below6s$precision_visang<- NULL
 accuracy_dataset_below6s <-  accuracy_dataset_below6s %>%
   arrange(ssid, tNo, timerezero)%>%
  group_by(ssid, tNo) %>%
   mutate(v_precision_visang = pixtoAng(hw = 32.5, d = 65 , r = 1200, v_precision_pix))%>%
   mutate(h_precision_visang = pixtoAng(hw = 52, d = 65 , r = 1920, h_precision_pix))%>% 
    mutate(avg_precision_visang = pixtoAng(hw = 52, d = 65 , r = 1920,precision_avg_pix ))%>%
    mutate(eucl_precision_visang = pixtoAng(hw = 52, d = 65 , r = 1920, precision_eucl_pix))%>%
    mutate(sd_v_precision_visang = pixtoAng(hw = 32.5, d = 65 , r = 1200, sd_v_precision_pix))%>%
     mutate(sd_h_precision_visang = pixtoAng(hw = 52, d = 65 , r = 1920, sd_h_precision_pix))%>%
     mutate(sd_avg_precision_visang = pixtoAng(hw = 32.5, d = 65 , r = 1200, sd_precision_avg_pix))

saveRDS(accuracy_dataset_below6s, file = "accuracy_dataset_below6sjan16eventdurprecis.rds")

#expeted ranges
accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(eucl_precision_visang))+
  geom_histogram()+
  geom_vline(xintercept = .5)

acc_plots$precision<- accuracy_dataset_below6s %>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(avg_precision_visang))+
  geom_histogram(aes(y=..density..), 
                position="identity")+
  # geom_histogram()+
  geom_vline(xintercept = .5, linetype = "dashed", color = "red", size = 1.5)+
  p$graphstyle+
  xlab("Precision (deg)")+
  ylab("Density")

acc_plots$precision

accuracy_dataset_below6s%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%
  ggplot(aes(sd_avg_precision_visang))+
  geom_histogram()+
  geom_vline(xintercept = .5)

accuracy_dataset_below6s %>%
    subset(detection3 == "f")%>%
  group_by(ssid, tNo)%>%
  mutate(losstest = n())
  summarise_if(is.numeric, mean, na.rm =TRUE)%>%

  ggplot(aes(gaze_loss_prop, new_dur))+
  geom_point()

# mutate(sd_eucl_precision_pix = sqrt())

# sd( (accuracy_dataset_below6s$new_dur  accuracy_dataset_below6s$duration)

  mutate(precision_pix = (h_precision_pix + v_precision_pix)/2) %>%
  mutate(v_precision_visang = pixtoAng(hw = 27.8, d = 62.45, r = 1080, stimsize = v_precision_pix)) %>%
  mutate(h_precision_visang = pixtoAng(hw = 51, d = 62.45, r = 1920, stimsize = h_precision_pix)) %>%
  mutate(precision_visang = (h_precision_visang + v_precision_visang)/2)

```

Precision analyses

```{r}

fixationscreen_only3<-

  View(accuracy_dataset_below6s)

test<- subset(accuracy_dataset_below6s, detection3 == "f")

fixationscreen_only3<- accuracy_dataset_below6s%>%
  subset(detection3 == "f")%>%
  group_by(ssid, tNo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

View(fixationscreen_only3)

unique(fixationscreen_only3$ssid)
fixationscreen_only3$ssid<- as.numeric(as.character(fixationscreen_only3$ssid))

fixationscreen_only3_lmer<- fixationscreen_only3
unique(fixationscreen_only3_lmer$ssid)

fixationscreen_only3_lmer$tNo2<- if_else(as.double(fixationscreen_only3_lmer$tNo)< 25, as.double(fixationscreen_only3_lmer$tNo), fixationscreen_only3_lmer$tNo - 24)

unique(fixationscreen_only3_lmer$tNo2)

#polynomial
fixationscreen_only3_lmer$ot1<- poly(fixationscreen_only3_lmer$tNo2, degree = 1, raw =FALSE)[,1]

# fixationscreen_only3_lmer$ot1[1,]
fixationscreen_only3_lmer$ot2<- poly(fixationscreen_only3_lmer$tNo2, degree = 2, raw =FALSE)[,2]

fixationscreen_only3_lmer$ot1_z<- scale(fixationscreen_only3_lmer$ot1, center = TRUE, scale = TRUE)[,1]
fixationscreen_only3_lmer$ot2_z<- scale(fixationscreen_only3_lmer$ot2, center = TRUE, scale = TRUE)[,1]

fixationscreen_only3_lmer%>%
  ggplot(aes(tNo, ot1_z))+
  geom_point()+
  geom_point(aes(y = ot2_z))

# fixationscreen_only3_lmer$spli_trial<- as.factor(fixationscreen_only3_lmer$spli_trial)
fixationscreen_only3_lmer$block<- if_else(fixationscreen_only3_lmer$tNo <25, -.5, +.5)

fixationscreen_only3_lmer$subject<- as.factor(as.character(fixationscreen_only3_lmer$ssid))
fixationscreen_only3

precision_models<- list()
library(lmerTest)

mean(fixationscreen_only3_lmer$avg_precision_visang, na.rm = TRUE) # 0.2719884
median(fixationscreen_only3_lmer$avg_precision_visang, na.rm = TRUE)# 0.2440743
sd(fixationscreen_only3_lmer$avg_precision_visang, na.rm = TRUE) #0.1080971
range(fixationscreen_only3_lmer$avg_precision_visang, na.rm = TRUE) #0.1065232 0.7233320

# vertical

mean(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE) # 0.3044482
median(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)# 0.2632572
sd(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE) # 0.1335215
range(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE) #0.1120570 0.9538436

# horizontal

mean(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE) # 0.2484294
median(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)#  0.2296711
sd(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE) # 0.09180471
range(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE) #0.09097805 0.60289107

fixationscreen_only3_lmer$ssid<- as.factor(as.character(fixationscreen_only3_lmer$ssid))

precision_models$prec_1 <- lmer(avg_precision_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | ssid), REML = FALSE,
                             data = fixationscreen_only3_lmer)

summary(precision_models$prec_1)

# fit data
(fixationscreen_only3_lmer) %>%
  ggplot(aes(tNo2, precision_visang))+
  # geom_smooth(aes(color = ssid), se = F)
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(precision_models$prec_1, fixationscreen_only3_lmer), colour = subject))+
  facet_grid(~block)

```

Linear mixed model fit by maximum likelihood . t-tests use Satterthwaite's method [lmerModLmerTest]
Formula: avg_precision_visang ~ (ot1_z + ot2_z) * block + (0 + ot1_z +      ot2_z) + (1 | ssid)
   Data: fixationscreen_only3_lmer

     AIC      BIC   logLik deviance df.resid 
 -6321.0  -6281.4   3167.5  -6335.0     2097

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-3.9049 -0.5446 -0.0864  0.4553  5.7883

Random effects:
 Groups   Name        Variance Std.Dev.
 ssid     (Intercept) 0.082648 0.28749 
 Residual             0.002479 0.04979 
Number of obs: 2104, groups:  ssid, 43

Fixed effects:
                Estimate   Std. Error           df t value Pr(>|t|)   
ot1_z          0.0007986    0.0011108 2061.0024024   0.719  0.47227   
ot2_z          0.0009260    0.0011245 2061.0040815   0.823  0.41033   
block          0.0065649    0.0022007 2061.0587841   2.983  0.00289 **
ot1_z:block    0.0014579    0.0022219 2061.0173472   0.656  0.51179   
ot2_z:block    0.0023842    0.0022486 2061.0224616   1.060  0.28914

# real data

(fixationscreen_only3_lmer) %>%
  ggplot(aes(tNo2, precision_visang))+
  # geom_line(aes(colour = subject))+
  geom_smooth(aes(color = subject), se = F)+
 # geom_vline(xintercept = 26)+
  #geom_line(aes(y = predict(precision_models$prec_1, fixationscreen_only3_lmer), colour = subject))+
  facet_grid(~spli_trial)

emmeans::emmeans(accurcay_models$prec_1, pairwise ~  block | ot2_z)
```{r}

#vertical precision

mean(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)
median(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)
sd(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)
min(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)
max(fixationscreen_only3_lmer$v_precision_visang, na.rm = TRUE)

#horizontal precision

mean(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)
median(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)
sd(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)
min(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)
max(fixationscreen_only3_lmer$h_precision_visang, na.rm = TRUE)

# vertical precision

precision_models$v_prec_1 <- lmer(v_precision_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | ssid), REML = FALSE,
                             data = fixationscreen_only3_lmer)

plot(precision_models$v_prec_1)
qqnorm(resid(precision_models$v_prec_1))

summary(precision_models$v_prec_1)

```
Linear mixed model fit by maximum likelihood . t-tests use Satterthwaite's method [lmerModLmerTest]
Formula: v_precision_visang ~ (ot1_z + ot2_z) * block + (0 + ot1_z + ot2_z) +      (1 | ssid)
   Data: fixationscreen_only3_lmer

     AIC      BIC   logLik deviance df.resid 
 -5217.6  -5178.1   2615.8  -5231.6     2097

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-3.8675 -0.4922 -0.0646  0.3834  6.5967

Random effects:
 Groups   Name        Variance Std.Dev.
 ssid     (Intercept) 0.105332 0.32455 
 Residual             0.004213 0.06491 
Number of obs: 2104, groups:  ssid, 43

Fixed effects:
                Estimate   Std. Error           df t value Pr(>|t|)   
ot1_z          0.0007272    0.0014480 2061.0035161   0.502  0.61558   
ot2_z          0.0009723    0.0014659 2061.0057547   0.663  0.50722   
block          0.0076593    0.0028688 2061.0786467   2.670  0.00765 **
ot1_z:block    0.0020506    0.0028965 2061.0234389   0.708  0.47906   
ot2_z:block    0.0031637    0.0029314 2061.0302560   1.079  0.28059   
---
```{r}
# horizontal precision

precision_models$h_prec_1 <- lmer(h_precision_visang ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | ssid), REML = FALSE,
                             data = fixationscreen_only3_lmer)

plot(precision_models$h_prec_1)
qqnorm(resid(precision_models$h_prec_1))

summary(precision_models$h_prec_1)

# fit data
(fixationscreen_only3_lmer) %>%
  ggplot(aes(tNo2, h_precision_visang))+
  # geom_smooth(aes(color = ssid), se = F)
 # geom_vline(xintercept = 26)+
  geom_line(aes(y = predict(precision_models$h_prec_1, fixationscreen_only3_lmer), colour = subject))+
  facet_grid(~spli_trial)

# real data

(fixationscreen_only3_lmer) %>%
  ggplot(aes(tNo2, h_precision_visang))+
  # geom_line(aes(colour = subject))+
  geom_smooth(aes(color = subject), se = F)+
 # geom_vline(xintercept = 26)+
  #geom_line(aes(y = predict(precision_models$prec_1, fixationscreen_only3_lmer), colour = subject))+
  facet_grid(~spli_trial)

saveRDS(fixationscreen_only3_lmer, "fixationscreen_only3_lmer_precision.rds")

saveRDS(accuracy_dataset_below6s, "accuracy_dataset_below6sjan16eventdurprecis.rds")
```

Heart rate variability

ALL WE NEED IS bounce

hr variability metrics to use
mean RMSSD

rr interval

```{r}

unique(accuracy_dataset_below6s$ssid)
library(readr)

read_plus_hrv <- function(flnm) {
  data.table::fread(flnm, select = c("ssid", "time", 'hr', "hrv", "hrp", "status", "tNo",  "Condition", "screencontent", "stim")) %>% 
    mutate(filename = flnm)
}

# now read them
library(purrr)
library(data.table)
library(dplyr)

# read all clean data for repixeling

Restperiod <-
  list.files(pattern = "*.csv", 
             full.names = T) %>% 
  map_df(~read_plus_hrv(.))

# keep only rest period

Restperiod1 <- subset(Restperiod, tNo == -7)

Restperiod1 %>%
  subset(ssid == 609)%>%
  ggplot(aes(time, hrp))+
  geom_line()

Restperiod1 <- select(Restperiod1, c(ssid, time, hrp))
Restperiod2<- select()
# df_dat <- split(Restperiod1, Restperiod1$ssid)

# Split dataframe by city
split_df <- split(Restperiod1[,3], list(Restperiod1$ssid))
Restperiod1$time

# Write out separate CSV for each city
for (ssid in names(split_df)) {
  # if (ssid == split_df[[ssid]]) {
  # }
    write_csv(split_df[[ssid]], paste0(ssid, "rest_hp.txt"), row.names = FALSE, col.names = FALSE, quote = FALSE)
  }

split_df[["344"]][,3]

lapply(names(SPLIT.DATA), function(nm) 
   write.csv(SPLIT.DATA[[nm]], paste0(nm, ".csv"), row.names = FALSE, quote = FALSE))

View(Restperiod1)

# unique(tmp.df4_full$ssid)
# tmp.df4_full<-  read_csv("_tmp.df4_full.csv")

hr<- subset(tmp.df4_full, is.na(hrp) == FALSE)

# rm(tmp.df4_full)
View(Restperiod)

# exclude 341 hr from biopac

```

Plotting Preferences

```{r}

p<- list()

sf = 1

p$graphstyle <-  theme(#base plot theme

  # axis
  axis.line.y = element_line(colour = "gray40", size = 0.5, lineend=0),
  axis.line.x = element_line(colour = 'gray40', size = 0.5, lineend=0),
  axis.ticks.x = element_line(colour = 'gray40', size = 0.5, lineend=0),
  strip.text.x = element_text(size = 11*(sf+1.5),  colour = "black"),
  strip.text.y = element_text(size = 11*(sf+1.5),  colour = "black"),

  #axis text

  axis.text.x = element_text(size = 16*(sf+.5), family = "sans", colour = "black", hjust =1.2), #angle = 45
  axis.text.y = element_text(size = 16*(sf+.5), family = "sans", colour = "black"),

  # titles
    plot.title = element_text(size = 16*(sf+1.5), hjust = .5),

  axis.title.y=element_text(size = 16*(sf+1), margin=margin(0,5,0,0)),
  axis.title.x=element_text(size = 16*(sf+1), margin=margin(0,5,0,0)),

  #ticks
    #axis.ticks = element_blank(),

  # panel
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill="transparent"),

  # legend
  legend.position = "none",
  # legend.direction =,
  text=element_text(size = 16, family = "sans"),

  # strip shades (reco rectagles)
  strip.background = element_blank())

```

```{r}

# 605 and 606 look the same

# gp heart rate
max(Restperiod1$hrp)
gp_heartpulse <-  Restperiod1 %>%
  subset(ssid == 339)%>%
  ggplot(aes(time - min(time), hrp/4740))+
  geom_line()+
    xlim(0,120)+
  theme_classic()+
  ylab("Pulse (AU)")+
  xlab("Time (s)")+
  p$graphstyle+
ggtitle("HP GPB")

```

# biopac heart rate

```{r}
# 120% .004s

rest_ss335_bio<- read.delim("335_IAPS_Physio_v3_rest.txt", skip = 40)

max(rest_ss335_bio$X2.73285)

nrow(rest_ss335_bio)
rest_ss335_bio$time<- seq(1,239725)

rest_ss335_bio$time_ms<- rest_ss335_bio$time/2000

# merge(rest_ss335_bio, as.data.frame(rest_ss335))

rest_ss335_bio%>%
  ggplot(aes(time_ms,X0.116638.1))+
  geom_line()+
    xlim(0,120)+
  theme_classic()

ss335_ecg_bioplot<- rest_ss335_bio%>%
  ggplot(aes(time_ms,X0.116638.1))+
  geom_line()+
    xlim(0,120)+
  theme_classic()+
  ylab("ECG (uv)")+
  xlab("Time (s)")+
  p$graphstyle+
  ggtitle("ECG MP160")

ss335_ecg_bioplot
gp_heartpulse

pulseechpatch<- gp_heartpulse+ss335_ecg_bioplot +plot_annotation(tag_levels = 'A')
pulseechpatch

ggsave("pulseechpatc.tiff", device = "tiff", dpi = 800, width = 12, height = 5)
```

Import parsed hrv from gazepoint

```{r}

library(readxl)
GP_HRVanalysys <- read_excel("GP_HRVanalysys.xlsx")
View(GP_HRVanalysys)

# Import the IBIS and artifact freefreeny

read_plus_ibi <- function(flnm) {
  data.table::fread(flnm) %>% 
    mutate(filename = flnm)
}

# now read
cleanIBI_gp <-
  list.files(pattern = "*.txt", 
             full.names = T) %>% 
  map_df(~read_plus_ibi(.))

cleanIBI_gp<- cleanIBI_gp%>%
  rename(IBI_clean = V1)

cleanIBI_gp$ssid<- substr(cleanIBI_gp$filename, 13,15)

# pointcare plots
cleanIBI_gp%>%
  # subset(ssid == 335)%>%
  group_by(ssid)%>%
  mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(IBI_clean_z, lead(IBI_clean_z)+IBI_clean_z))+
  geom_point()+
  geom_smooth(method = 'lm', se = F)
  facet_grid(~ssid)

  # distribution of IBI lenghts
  unique(cleanIBI_gp$ssid)
unique(Restperiod1$ssid)

time<- subset(Restperiod1, is.na(ssid) == FALSE & ssid != 341)

nrow(time)
nrow(cleanIBI_gp)

cleanIBI_gp$time<- Restperiod1$time

cleanIBI_gp %>%
  # subset(ssid == 335)%>%
  group_by(ssid)%>%

  # mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(IBI_clean))+
  geom_histogram()

cleanIBI_gp %>%
  # subset(ssid == 335)%>%
  group_by(ssid)%>%
    mutate(time == 1:n())%>%
  subset(ssid == 335)%>%
  # mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(time, IBI_clean))+
  geom_point()

```
# data from biopac

```{r}

GP_HRVanalysys$MEANHR
mean(GP_HRVanalysys$RMMSSD, na.rm = TRUE)

# import biopac
BIO_HRVanalysys <- read_excel("D:/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/Validation/BIO_HRVanalysys.xlsx")

View(BIO_HRVanalysys)
View(BIO_HRVanalysys)
BIO_HRVanalysys$ssid
GP_HRVanalysys$ssid

GP_HRVanalysys$ssid<- as.character(GP_HRVanalysys$ssid)
BIO_HRVanalysys$ssid<- as.character(BIO_HRVanalysys$ssid)

combined_bio_GP_HRV<- left_join(GP_HRVanalysys, BIO_HRVanalysys, by = "ssid")

View(combined_bio_GP_HRV)
# GP_HRVanalysys$

# r to R
combined_bio_GP_HRV%>%
  subset(ssid!= 608)%>%
ggplot(aes(MEANRR.x, MEANRR.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

combined_bio_GP_HRV_no601 <- subset(combined_bio_GP_HRV, ssid != 601)

cor.test(subset(combined_bio_GP_HRV, ssid!= 605)$MEANRR.x, subset(combined_bio_GP_HRV, ssid!= 605)$MEANRR.y, method = "kendal") # 0.9789474 T = 188, p-value < 0.00000000000000022

cor.test(subset(combined_bio_GP_HRV, ssid!= 605 & ssid!= 608)$MEANRR.x, subset(combined_bio_GP_HRV, ssid!= 605 & ssid!= 608)$MEANRR.y, method = "kendal")

# T = 169, p-value = 0.000000000000003775
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.9766082

cor.test(subset(combined_bio_GP_HRV)$MEANRR.x, subset(combined_bio_GP_HRV)$MEANRR.y, method = "kendal")

# T = 199, p-value = 0.000000000002767
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.8952381

```

# rMSSD
```{r}
combined_bio_GP_HRV %>%
  subset(ssid!= 608 & ssid!= 605)%>%
ggplot(aes(SDNN.x, SDNN.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid!= 605 & ssid!= 608)$SDNN.x, subset(combined_bio_GP_HRV, ssid!= 605 & ssid!= 608)$SDNN.y, method = "kendal")

# 
# 	Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV, ssid != 605 & ssid != 608)$SDNN.x and subset(combined_bio_GP_HRV, ssid != 605 & ssid != 608)$SDNN.y
# T = 148, p-value = 0.000001521
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.7309942

cor.test(subset(combined_bio_GP_HRV)$SDNN.x, subset(combined_bio_GP_HRV)$SDNN.y, method = "kendal")

# data:  subset(combined_bio_GP_HRV)$SDNN.x and subset(combined_bio_GP_HRV)$SDNN.y
# T = 169, p-value = 0.00004489
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6095238

```

PNN50

```{r}
combined_bio_GP_HRV %>%
  subset(ssid!= 608)%>%
ggplot(aes(pNN50.x, pNN50.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid!= 608)$pNN50.x, subset(combined_bio_GP_HRV, ssid!= 608)$pNN50.y, method = "kendal")

```
data:  subset(combined_bio_GP_HRV, ssid != 608)$pNN50.x and subset(combined_bio_GP_HRV, ssid != 608)$pNN50.y
z = 3.7951, p-value = 0.0001476
alternative hypothesis: true tau is not equal to 0
sample estimates:
      tau 
0.6357085

nn50

```{r}

combined_bio_GP_HRV %>%
  # subset(ssid!= 608)%>%
ggplot(aes(NN50.x, NN50.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV)$NN50.x, subset(combined_bio_GP_HRV)$NN50.y, method = "kendal")

```

RMSSD
```{r}
combined_bio_GP_HRV %>%
  subset( ssid!= 608)%>%
ggplot(aes(RMMSSD.x, RMMSSD.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid!= 608)$RMMSSD.x, subset(combined_bio_GP_HRV, ssid!= 608)$RMMSSD.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV, ssid != 608)$RMMSSD.x and subset(combined_bio_GP_HRV, ssid != 608)$RMMSSD.y
# T = 153, p-value = 0.00007322
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6105263

cor.test(subset(combined_bio_GP_HRV)$RMMSSD.x, subset(combined_bio_GP_HRV)$RMMSSD.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV)$RMMSSD.x and subset(combined_bio_GP_HRV)$RMMSSD.y
# T = 173, p-value = 0.00001127
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#      tau 
# 0.647619

```

```{r}
# r to R
combined_bio_GP_HRV%>%
ggplot(aes(MEANHR.x, MEANHR.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

```
FRequency measures
Of the frequency-domain methods, data related to the amount of low frequency (LF) heartbeats is often used (0.04 to 0.15 Hz) as a measure of sympathetic nervous system activity. Measurements of high-frequency (HF; 0.15 Hz to 0.4 Hz) and very-low-frequency (VLF) are also used

```{r}
# low frequency
combined_bio_GP_HRV%>%
  subset(ssid!= 601)%>%
ggplot(aes(lfperce.x, lfperce.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)
  # xlim(0,200)

cor.test(subset(combined_bio_GP_HRV, ssid != 601)$lfperce.x, subset(combined_bio_GP_HRV, ssid != 601)$lfperce.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV, ssid != 601)$lfperce.x and subset(combined_bio_GP_HRV, ssid != 601)$lfperce.y
# T = 163, p-value = 0.000001342
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.7157895

cor.test(subset(combined_bio_GP_HRV)$lfperce.x, subset(combined_bio_GP_HRV)$lfperce.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV)$lfperce.x and subset(combined_bio_GP_HRV)$lfperce.y
# T = 171, p-value = 0.00002292
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6285714

```

# vlf

```{r}
combined_bio_GP_HRV %>%
    subset(ssid!= 601)%>%
ggplot(aes(vlfperc.x, vlfperc.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid != 601)$vlfperc.x, subset(combined_bio_GP_HRV, ssid != 601)$vlfperc.y, method = "kendal")

cor.test(subset(combined_bio_GP_HRV, ssid != 601)$vlfpower.x, subset(combined_bio_GP_HRV, ssid != 601)$vlfpower.y, method = "kendal")
cor.test(subset(combined_bio_GP_HRV, ssid != 601)$lfnu.x, subset(combined_bio_GP_HRV, ssid != 601)$lfnu.y, method = "kendal")

cor.test(subset(combined_bio_GP_HRV, ssid != 601)$pNN50.x, subset(combined_bio_GP_HRV, ssid != 601)$pNN50.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV, ssid != 601)$vlfperc.x and subset(combined_bio_GP_HRV, ssid != 601)$vlfperc.y
# T = 163, p-value = 0.000001342
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.7157895

cor.test(subset(combined_bio_GP_HRV)$vlfperc.x, subset(combined_bio_GP_HRV)$vlfperc.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV)$vlfperc.x and subset(combined_bio_GP_HRV)$vlfperc.y
# T = 174, p-value = 0.000007786
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6571429

# hf
combined_bio_GP_HRV%>%
  # subset(ssid!= 601)%>%
ggplot(aes(hfperent.x, hfperent.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid != 601)$hfperent.x, subset(combined_bio_GP_HRV, ssid!= 601)$hfperent.y, method = "kendal")
# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV, ssid != 601)$hfperent.x and subset(combined_bio_GP_HRV, ssid != 601)$hfperent.y
# T = 159, p-value = 0.000007745
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6736842

cor.test(subset(combined_bio_GP_HRV)$hfperent.x, subset(combined_bio_GP_HRV)$hfperent.y, method = "kendal")

# Kendall's rank correlation tau
# 
# data:  subset(combined_bio_GP_HRV)$hfperent.x and subset(combined_bio_GP_HRV)$hfperent.y
# T = 171, p-value = 0.00002292
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.6285714

combined_bio_GP_HRV$group <- if_else(combined_bio_GP_HRV$ssid< 400, "NT", "CLIN")

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, RMMSSD.x))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, RMMSSD.x))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, RMMSSD.y))+
  geom_jitter(width = .3)+
  geom_boxplot()

####

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, MEANRR.x))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, MEANRR.y))+
  geom_jitter(width = .3)+
  geom_boxplot()

# frequency

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, lfperce.x))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, lfperce.y))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, hfperent.x))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
  group_by(group)%>%
  ggplot(aes(group, hfperent.y))+
  geom_jitter(width = .3)+
  geom_boxplot()

combined_bio_GP_HRV%>%
    subset(ssid!= 608)%>%
ggplot(aes(hfpower.x, hfpower.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

```

HF power

```{r}

combined_bio_GP_HRV%>%
    subset(ssid!= 608)%>%
ggplot(aes(hfpower.x, hfpower.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV, ssid != 608)$hfpower.x, subset(combined_bio_GP_HRV, ssid!= 608)$hfpower.y, method = "kendal")

```
Kendall's rank correlation tau

data:  subset(combined_bio_GP_HRV, ssid != 601)$hfpower.x and subset(combined_bio_GP_HRV, ssid != 601)$hfpower.y
T = 169, p-value = 0.00000005841
alternative hypothesis: true tau is not equal to 0
sample estimates:
      tau 
0.7789474

LF power

```{r}
combined_bio_GP_HRV%>%
    subset(ssid!= 608 & ssid!= 342)%>%
ggplot(aes(lfpower.x, lfpower.y, color = ssid))+
  geom_point()+
  geom_smooth(method = "lm", se = F)

cor.test(subset(combined_bio_GP_HRV)$hfpower.x, subset(combined_bio_GP_HRV)$hfpower.y, method = "kendal")

cor.test(subset(combined_bio_GP_HRV, ssid !=608& ssid!= 342)$hfpower.x, subset(combined_bio_GP_HRV, ssid !=608& ssid!= 342)$hfpower.y, method = "kendal")

```

# correlations

```{r}

cor.test(GP_HRVanalysys$MEANRR.x,BIO_HRVanalysys$MEANRR.x)

```

```{r}
# IBI data to compare from biopac

read_plus_ibi_bio <- function(flnm) {
  data.table::fread(flnm) %>% 
    mutate(filename = flnm)
}

# now read
cleanIBI_bio <-
  list.files(pattern = "*.txt", 
             full.names = T) %>% 
  map_df(~read_plus_ibi_bio(.))

cleanIBI_bio<- cleanIBI_bio%>%
  rename(IBI_clean = V1)

cleanIBI_bio$ssid<- substr(cleanIBI_bio$filename, 13,15)

# pointcare plots

cleanIBI_gp$IBI_clean_next<- lead(cleanIBI_gp$IBI_clean)

poitcare_gp<- cleanIBI_gp%>%
  # subset(ssid == 335)%>%
  group_by(ssid)%>%
  # mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(IBI_clean, IBI_clean_next,  color = ssid))+
  geom_point(alpha = .3)+
  geom_smooth(method = 'lm', se = F)+
  # facet_grid(~ssid)+
  xlab("RR[n] (ms)")+
  ylab("RR[n+1] (ms)")+
    p$graphstyle
  # ggtitle("GP3 Gazepoint")+
  p$graphstyle

poitcare_gp

poitcare_gp <- poitcare_gp +
  scale_y_continuous(limits = c(400, 1600), breaks = c(400, 1000, 1600))+
   scale_x_continuous(limits = c(300, 1600), breaks = c(400, 1000, 1600))

scale_x_continuous(limits = c(500, 1300), breaks = c(500, 900, 1300))

poitcare_gp

```

```{r}

  # cleanIBI_bio$ssid<- a

cleanIBI_bio$IBI_clean_next<- lead(cleanIBI_bio$IBI_clean)
pointcare_bio<-  cleanIBI_bio %>%
    # subset(ssid == 335)%>%
  subset(ssid != 341)%>%
  group_by(ssid)%>%
  # mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(IBI_clean, IBI_clean_next,  color = ssid))+
  geom_point(alpha = .3)+
  geom_smooth(method = 'lm', se = F)+
  # facet_grid(~ssid)+
  xlab("RR[n] (ms)")+
  ylab("RR[n+1] (ms)")+
    p$graphstyle
  ggtitle("MP150 BIOPAC")

pointcare_bio<- pointcare_bio+
  scale_y_continuous(limits = c(500, 1300), breaks = c(500, 900, 1300))+
 scale_x_continuous(limits = c(500, 1300), breaks = c(500, 900, 1300))

pointcare<- poitcare_gp+ pointcare_bio+ plot_annotation(tag_levels = 'A')

```

cleanIBI_gp$group <- if_else(cleanIBI_gp$ssid< 400, 1, 2)

  cleanIBI_gp %>%
  # subset(ssid != 341)%>%
  group_by(ssid, group)%>%
  # mutate(IBI_clean_z = scale(IBI_clean,center = TRUE)[,1])%>%
  ggplot(aes(IBI_clean, lead(IBI_clean)+IBI_clean,  color = as.factor(group)))+

  geom_smooth(method = 'lm', se = F)

  facet_grid(~ssid)

```

db_parsed_summarised_fix1.2$tNo2<- if_else(as.double(db_parsed_summarised_fix1.2$tNo)< 25, as.double(db_parsed_summarised_fix1.2$tNo), db_parsed_summarised_fix1.2$tNo - 24)

unique(fixationscreen_only3_lmer$tNo2)

#polynomial
unique(db_parsed_summarised_fix1.2$screencontent)

db_parsed_summarised_fix1.2$ot1<- poly(db_parsed_summarised_fix1.2$tNo2, degree = 1, raw =FALSE)

db_parsed_summarised_fix1.2$ot2<- poly(db_parsed_summarised_fix1.2$tNo2, degree = 2, raw =FALSE)[,2]

db_parsed_summarised_fix1.2$ot1_z<- scale(db_parsed_summarised_fix1.2$ot1, center = TRUE, scale = TRUE)
db_parsed_summarised_fix1.2$ot2_z<- scale(db_parsed_summarised_fix1.2$ot2, center = TRUE, scale = TRUE)

db_parsed_summarised_fix1.2$block<- if_else(db_parsed_summarised_fix1.2$tNo < 25, -.5, +.5)

db_parsed_summarised_fix1.2$subject<- as.factor(as.character(db_parsed_summarised_fix1.2$ssid))

precision_models$prec_stim <- lmer(RMS ~ (ot1_z + ot2_z) * block+
                              (0+ ot1_z+ot2_z)+ (1 | ssid), REML = FALSE,
                             data = subset(db_parsed_summarised_fix1.2,as.numeric(mean_fix_dur) < 3000))

summary(precision_models$prec_stim)

db_parsed_summarised_fix1.2 <- subset(db_parsed_summarised_fix, ssid <500 & tNo> 0)

db_parsed_summarised_fix1.2

 #precision correlation with calibratin accuracy

unique(tmp.df4_sum_below500_1.2$cali_order)
unique(fixationscreen_only2_lmer$tNo)
fixationscreen_only3_lmer$ssid<- as.character(fixationscreen_only3_lmer$subject)

tmp.df4_sum_below500_1.2$ssid<- as.character(tmp.df4_sum_below500_1.2$ssid)

fixationscreen_only3_lmer$cali_order<- as.character(if_else(fixationscreen_only3_lmer$tNo < 24, 2, 3))

tmp.df4_sum_below500_1.2_select<- select(tmp.df4_sum_below500_1.2, c(ssid, cali_order, avg_error1, cali_no_resc_agg))

fixationscreen_only3_lmer<- left_join(fixationscreen_only3_lmer, tmp.df4_sum_below500_1.2_select)

fixationscreen_only3_lmer %>%
  group_by(ssid, cali_order)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(avg_error1, precision_visang))+
  geom_point()+
geom_smooth(method = 'lm', se = F)+
  facet_grid(~cali_order)

fixationscreen_only2_lmer_1.2<- fixationscreen_only2_lmer%>%
  group_by(ssid, cali_order)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

ggplot(aes(avg_error1, precisi))+
  geom_point()+
geom_smooth(method = 'lm', se = F)+
  facet_grid(~cali_order)

cor.test(subset(fixationscreen_only2_lmer_1.2, cali_order == 2)$avg_error1, subset(fixationscreen_only2_lmer_1.2, cali_order == 2)$h_accuracy_visang)

cor.test(subset(fixationscreen_only2_lmer_1.2, cali_order == 3)$avg_error1, subset(fixationscreen_only2_lmer_1.2, cali_order == 3)$h_accuracy_visang)

# Save an object to a file
saveRDS(fixationscreen_only3_lmer, file = "fixationscreen_only3_lmer.rds")
# Restore the object
readRDS(file = "my_data.rds"

```{r}

View(fixationscreen_only1$)

summary(model_time_acc)

fixationscreen_only2_lmer

options(scipen = 999)

accurcay_models<- list()

 #accuracy correlation with calibratin accuracy

unique(tmp.df4_sum_below500_1.2$cali_order)
unique(fixationscreen_only2_lmer$tNo)
fixationscreen_only2_lmer$ssid<- as.character(fixationscreen_only2_lmer$ssid)

tmp.df4_sum_below500_1.2$ssid<- as.character(tmp.df4_sum_below500_1.2$ssid)

fixationscreen_only2_lmer$cali_order<- as.character(if_else(fixationscreen_only2_lmer$tNo < 24, 2, 3))

tmp.df4_sum_below500_1.2_select<- select(tmp.df4_sum_below500_1.2, c(ssid, cali_order, avg_error1, cali_no_resc_agg))

fixationscreen_only2_lmer<- left_join(fixationscreen_only2_lmer, tmp.df4_sum_below500_1.2_select)

fixationscreen_only2_lmer %>%
  group_by(ssid, cali_order)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(avg_error1, v_accuracy_visang))+
  geom_point()+
geom_smooth(method = 'lm', se = F)+
  facet_grid(~cali_order)

fixationscreen_only2_lmer_1.2<- fixationscreen_only2_lmer%>%
  group_by(ssid, cali_order)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

ggplot(aes(avg_error1, precisi))+
  geom_point()+
geom_smooth(method = 'lm', se = F)+
  facet_grid(~cali_order)

cor.test(subset(fixationscreen_only2_lmer_1.2, cali_order == 2)$avg_error1, subset(fixationscreen_only2_lmer_1.2, cali_order == 2)$h_accuracy_visang)

cor.test(subset(fixationscreen_only2_lmer_1.2, cali_order == 3)$avg_error1, subset(fixationscreen_only2_lmer_1.2, cali_order == 3)$h_accuracy_visang)

#neither vertical or horizontal or global reach significance

```

fixationscreen_only2$spli_trial %>%
  ggplot(aes(tNo,accuracy ))+
  geom_point()
  #stat_summary(geom = 'pointrange')+
  geom_vline(xintercept = 24)

#

cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 2)$cali_no_resc_agg, subset(tmp.df4_sum_below500_1.2, cali_order == 2)$gaze_loss_prop)

cor.test(subset(tmp.df4_sum_below500_1.2, cali_order == 3)$cali_no_resc_agg, subset(tmp.df4_sum_below500_1.2, cali_order == 3)$gaze_loss_prop)

Pupil analysis

```{r}
readRDS(file = "db_full6new_304_up_500less.rds")

db_full4new_stim_pupil$ssid
db_full4new_stim_pupil$subject
View(db_full6new_304_up)

db_full6new_304_up_500less<- subset(db_full6new_304_up, subject< 500 & tNo > 0)

db_full4new_stim_pupil$BIO_Global.Mean

db_full6new_304_up_500less$Bio_Mean_HR_z

unique(db_full6new_304_up_500less$screencontent)

db_full6new_304_up_500less$Bio_Mean_HR_z<- scale(db_full6new_304_up_500less$Bio_Mean_HR, center = TRUE, scale = TRUE)
db_full6new_304_up_500less$BIO_Global.Mean_z<- scale(db_full6new_304_up_500less$BIO_Global.Mean, center = TRUE, scale = TRUE)
db_full6new_304_up_500less$pup_basCor_z<- scale(db_full6new_304_up_500less$pup_basCor, center = TRUE, scale = TRUE)

db_full6new_304_up_500less$hr_interp<- scale(db_full6new_304_up_500less$hr_interp, center = TRUE, scale = TRUE)

# Pupil

```

```
```{r}

summary(lmer(arousal ~  pup_basCor_z + BIO_Global.Mean_z + Bio_Mean_HR_z +ArousalMean+ (1 | stimIAPS) +(1 | ssid), REML = FALSE,
                      data = subset(db_full6new_304_up_500less)))

summary(lmer(pup_basCor_z ~  ArousalMean + (1 | stimIAPS)  +(1 | ssid), REML = FALSE,
                      data = db_full6new_304_up_500less))

cognitivepupil$pup_arous_brigh <- lmer(pup_basCor ~  arousal_c +  BRIGHTNESSc +(1 | stimIAPS)  +(1 | ssid), REML = FALSE,
                      data = db_full6new_304_up_500less)
summary(cognitivepupil$pup_arous_brigh)

MuMIn::r.squaredGLMM(cognitivepupil$pup_arous_brigh)

library(lmerTest)

cognitivepupil$pup_arous_1 <- lmer(pup_basCor ~  arousal_c +( 1| stimIAPS)  +(1 | ssid), REML = FALSE,
                      data = db_full6new_304_up_500less)

library(readr)
library(tidyverse)
cognitivepupil$pup_arous_1bri <- lmer(pup_basCor ~  BRIGHTNESSc +( 1| stimIAPS)  +(1 | ssid), REML = FALSE,
                      data = db_full6new_304_up_500less)

MuMIn::r.squaredGLMM(cognitivepupil$pup_arous_1)
MuMIn::r.squaredGLMM(cognitivepupil$pup_arous_1bri) # 20

db_full6new_304_up_500less$fit_pup_arou_bright<- 
  predict(cognitivepupil$pup_arous_brigh, type = c("response"), terms = "BRIGHTNESSc")
?predict

pupil_plots<- list()

write_csv(db_full6new_304_up_500less, "Pupil_data_db_full6new_304_up_500less.csv")

pupil_plots$brightness<-  db_full6new_304_up_500less %>%
  #subset(!is.na(Alexithymia))%>%
  group_by(ssid,stimIAPS) %>%
  summarise_at(c("arousal_c", "TASc","BRIGHTNESSc", "pup_basCor", "fit_pup_arou_bright"), mean, na.rm = TRUE)%>%
  ggplot(aes(BRIGHTNESSc, pup_basCor, color = ssid))+
  geom_point(alpha = .3)+
  geom_smooth(aes(group = ssid), method = 'lm', se = F)+
  stat_smooth (geom="line", method = 'lm', alpha = .8, size = 1.5, color = "gray30")+
    ylab('Pupilary change (z)')+
  xlab('Brightness (z)')+
  p$graphstyle
    scale_color_brewer(palette = "Dark2")

pupil_plots$brightness<- pupil_plots$brightness+
  scale_y_continuous(limits = c(-12,4), breaks = c(-12,-4, 4))+
  scale_x_continuous(limits = c(-2.5,2), breaks = c(-2.5,-0, 2))

library(dplyr)

pupil_plots$arousalpup<-  db_full6new_304_up_500less %>%
  #subset(!is.na(Alexithymia))%>%
  group_by(ssid,stimIAPS) %>%
  summarise_at(c("arousal", "TASc", "pup_basCor", "fit_pup_arou_bright"), mean, na.rm = TRUE)%>%
  ggplot(aes(pup_basCor, arousal, color = ssid))+
  geom_point(alpha = .3)+
  geom_smooth(aes(group = ssid), method = 'lm', se = F)+
  stat_smooth (geom="line", method = 'lm', alpha = .8, size = 1.5, color = "gray30")+
    ylab('Pupilary change (z)')+
  xlab('Self reported arousal (z)')+
  p$graphstyle
  scale_color_brewer(palette = "Dark2")

pupil_plots$arousalpup

pupil_plots$arousalpup<- pupil_plots$arousalpup+
  scale_y_continuous(limits = c(-10,10), breaks = c(-10,0, 10))+
  scale_x_continuous(limits = c(-12,6), breaks = c(-12,6))

pupil_plots$arousalpup
# Save an object to a file
saveRDS(db_full6new_304_up_500less, file = "db_full6new_304_up_500less.rds")
# Restore the object

library(patchwork)
library(ggplot2)

patchpup<- pupil_plots$arousalpup+ pupil_plots$brightness+theme(axis.title.y = element_blank())
patchpup<- patchpup+plot_annotation(tag_levels = 'A')

ggsave("patchpup.tiff", patchpup, device = 'tiff', width = 10, height = 6, dpi = 800)
```

Quick check Hr pulse

```
```{r}

unique(one$ssid)

one1<- subset(one, ssid == 336)

# pp340 <- subset(tmp.df4_full, tNo = -7 & ssid == 340)

one1 %>%
  ggplot(aes(time - min(time), hrp))+
  geom_line()+
  xlim(0,30)

hrclass<-  kmeans(one1$hrp, 2, nstart = 25)

one1$hrclass<- hrclass$cluster

one1$time2hr<- if_else(one1$hrclass == 2, one1$time - min(one1$time), NULL)

stats::fil
one1 %>%
  ggplot(aes(time-min(time), stats::filter(hrp, 1)))+
  geom_line()+
  geom_vline(aes(xintercept = time2hr), alpha = .05, color = "red")+
  xlim(50, 80)+ theme_classic()

```