### 2022-02-02 Subj_Obj_Ar_Bayes_Prev.Rmd
Initial stab at bayes prevalence
Try and estimate bayes prevalence

---
title: "Subj_Arousal_bayes_prevalence"
author: "Helio"
date: "02/02/2022"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

Subj_Arousal_bayes_prevalence
pupil_arousal_findings$pupil_from_ar_slope_vale <- lmer(pup_basCor ~ (valence_c* arousal_c) +
                                                    Mean_gray_z  +
                                                    (1  | ssid)+
                                                    (0+ arousal_c | ssid),
                                                  REML = FALSE,
                      data = db_full4new_stim_screen_pupil_nopract%>%
                         subset(Group == "NT")%>%
                        subset(pupil_outlier == FALSE)%>%
                        subset(arousal_outler == FALSE))

summary(pupil_arousal_findings$pupil_from_ar_slope_vale)

```

```{r cars}
summary(cars)
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.