### 2022-02-02 Subj_Obj_Ar_Bayes_Prev.Rmd
adapted robins fucntions to do lm
adapted the functions to do lm instead of t.tests and do prevalence testing for multiple terms

---
title: "Subj_Arousal_bayes_prevalence"
author: "Helio"
date: "02/02/2022"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

Subj_Arousal_bayes_prevalence
pupil_arousal_findings$pupil_from_ar_slope_vale <- lmer(pup_basCor ~ (valence_c* arousal_c) +
                                                    Mean_gray_z  +
                                                    (1  | ssid)+
                                                    (0+ arousal_c | ssid),
                                                  REML = FALSE,
                      data = db_full4new_stim_screen_pupil_nopract%>%
                         subset(Group == "NT")%>%
                        subset(pupil_outlier == FALSE)%>%
                        subset(arousal_outler == FALSE))

# remeber this does contain pratice, juts a leftover of an old name not to break code
# db_full4new_stim_screen_pupil_nopract_no_outl$tNo

# 1st we need to conduct the test within each participant and store the p value

# let's do lm or lmer
# stoe just the data we need
db_full4new_stim_screen_pupil_nopract_no_outl<- db_full4new_stim_screen_pupil_nopract%>%
                         subset(Group == "NT")%>%
                        subset(pupil_outlier == FALSE)%>%
                        subset(arousal_outler == FALSE)

# 2.  LEVEL 1
#

# 2.1. Within-participant statistical test
# This loop performs within-participant statistical test. Here, a t-test for
# non-zero mean which is the simplest statistical test. In general, any
# statistical test can be used at Level 1.
Nsub = length(unique(db_full4new_stim_screen_pupil_nopract_no_outl$subject))

subject_id<- unique(db_full4new_stim_screen_pupil_nopract_no_outl$subject)

# p <- vector(mode="numeric",length=Nsub) #original
# instead for our case create a matrix
p <- matrix(data = NA, nrow = Nsub, ncol = 6) #

# t
# si = 1
# rm(si)

# p <- vector(mode="numeric",length=Nsub) # 
# for( si in 1:Nsub ){
#   t = t.test(rawdat[,si], mu=0)
#   p[si] = t[3]
# }

# test
# lmoutputtest<- summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
#    data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,subject == 320)))
# 
# lmoutputtest
# lmoutputtest[["coefficients"]][,4]

# this fucntion fits a multiple regression per individual and returns the summary
# then we stoe the p valus for each term in a matrix yo ue in level 2

db_full4new_stim_screen_pupil_nopract_no_outl$pup_resid

estimate <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  estimate[si,1:5] = t(matrix(t[["coefficients"]][,1])) 
  estimate[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(estimate)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(estimate)

# save estimates and se

# instead for our case create a matrix
se <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  se[si,1:5] = t(matrix(t[["coefficients"]][,2])) 
  se[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(se)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")

View(se)

t$coefficients[,4]

# now p values
p <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  p[si,1:5] = t(matrix(t[["coefficients"]][,4])) 
    p[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(p)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(p)

```

```{r}

# p holds p-values of test for each participant
alpha = 0.05  
indsig = p<alpha

# the binary variable indsig indicates whether the within-participant
# t-test is, or not, significant for each participant (1 entry for each
# participant)

# 2.2. Loading within-participant statistical test.
# You can also load your own within-participant statistical test results here.
# Load binary results into binary indsig vector with one entry per participant. 
# See also example_csv.R for an example.

#
# 3.  LEVEL 2
#

# Bayesian prevalence inference is performed with three numbers: 
# k, the number of significant participants (e.g. sum of binary indicator
# variable)
# n, the number of participants in the sample
# alpha, the false positive rate
n <- Nsub
# table(indsig)

# lets start with valence
k_2 <- sum(indsig[,2])
table(indsig[,2])

?binom.test
binom.test(k_2, 55, 0.05, alternative="two.sided",  )

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
bayesprev_posterior()
pdf_2 <- bayesprev_posterior(xvals, k_2, n) # function needed
pdf_2

plot(xvals, pdf_2, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_2, n)
pmap = bayesprev_posterior(xmap, k_2, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_2, n)
pbound = bayesprev_posterior(xbound, k_2, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_2, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_2, n)
h2 = bayesprev_posterior(i2, k_2, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# try arosual

```

```{r}
k_3 <- sum(indsig[,3])
k_3
# k_3/55 (20%)
12/55

binom.test(k_3, 55, 0.05, alternative="two.sided") 
# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
pdf_3 <- bayesprev_posterior(xvals, k_3, n) # function needed
pdf_3

plot(xvals, pdf_3, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_3, n)
pmap = bayesprev_posterior(xmap, k_3, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_3, n)
pbound = bayesprev_posterior(xbound, k_3, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_3, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_3, n)
h2 = bayesprev_posterior(i2, k_3, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

brighteness
```{r}
k_4 <- sum(indsig[,4])
k_4/55

# plot posterior distribution of population prevalence

xvals <- seq(0, 1, .01)
pdf_4 <- bayesprev_posterior(xvals, k_4, n) # function needed
pdf_4

plot(xvals, pdf_4, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_4, n)
pmap = bayesprev_posterior(xmap, k_4, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_4, n)
pbound = bayesprev_posterior(xbound, k_4, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_4, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_4, n)
h2 = bayesprev_posterior(i2, k_4, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# is the p value just a bonomial test on the counts of significants and non significant results?
```

<!-- interaction -->

<!-- arousal valence-->

```{r}
k_5 <- sum(indsig[,5])

# plot posterior distribution of population prevalence

xvals <- seq(0, 1, .01)
pdf_5 <- bayesprev_posterior(xvals, k_5, n) # function needed
pdf_5

plot(xvals, pdf_5, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_5, n)
pmap = bayesprev_posterior(xmap, k_5, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_5, n)
pbound = bayesprev_posterior(xbound, k_5, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_5, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_5, n)
h2 = bayesprev_posterior(i2, k_5, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

```{r}
#    Load the nonlinear equation solver
   library(nleqslv)
 #  Load the required functions from the current working directory
source("bayesprev.R")

  # load the data
  sigdat = p
    # read.csv("example_data.csv", header =FALSE);

  alpha = 0.05                       #   this specifies the alpha value used for the within-unit tests
  Ntests = nrow(sigdat);        #   number of tests (e.g. participants)
  Nsigtests_3 = sum(sigdat[,3]);    #   number of significant tests

  # Plot the posterior  pdf for the population prevalence proportion

  xvals <- seq(0, 1, .01)
  pdf_3.1 <- bayesprev_posterior(xvals, Nsigtests_3, Ntests)
  plot(xvals, pdf_3.1, type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

  # 0.95 lower bound for the population prevalence proportion

  b = bayesprev_bound(0.05, Nsigtests_3, Ntests)
  print(b)

  # MAP maximum a posteriori [MAP estimate of the population prevalence proportion

  m = bayesprev_map(Nsigtests_3, Ntests)
  print(m)

  # 96% HPDI for the population prevalence proportion highest posterior density intervals (HPDIs)

  int = bayesprev_hpdi(0.96, Nsigtests_3, Ntests)
  print(int)

  # Example of possible Bayesian prevalence analyses on a common plot.

  xvals <- seq(0, 1, .01)
  pdf_3.1 <- bayesprev_posterior(xvals, Nsigtests_3, Ntests)
  plot(xvals, pdf_3.1, type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

  # Add the MAP estimate as a point

  xmap = bayesprev_map(Nsigtests_3, Ntests)
  pmap = bayesprev_posterior(xmap, Nsigtests_3, Ntests)
  points(xmap, pmap, cex =2, col= "red", pch=16)

  # Add the .95 lower bound as a vertical line

  xbound = bayesprev_bound(0.05, Nsigtests_3, Ntests)
  pbound = bayesprev_posterior(xbound, Nsigtests_3, Ntests)
  lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

  # Add the 0.96 HPDI

  int = bayesprev_hpdi(0.96, Nsigtests_3, Ntests)
  i1 = int[1]
  i2 = int[2]
  h1 = bayesprev_posterior(i1, Nsigtests_3, Ntests)
  h2 = bayesprev_posterior(i2, Nsigtests_3, Ntests)
  lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

valence
```{r}

```

```{r}

range(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor)
?plot()
  plot(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor, 
       type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

testdensity<-density(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor, n = 55)

db_full4new_stim_screen_pupil_nopract_no_outl<- db_full4new_stim_screen_pupil_nopract_no_outl%>%
  group_by(ssid)%>%
  mutate(testdensity_x = density(pup_basCor, n = 1)$x,
         testdensity_y = density(pup_basCor, n = 1)$y)

db_full4new_stim_screen_pupil_nopract_no_outl%>%
  subset(ssid == 300)%>%
  # group_by(ssid)%>%
  mutate(testdensity_x = density(pup_basCor)$x,
         testdensity_y = density(pup_basCor)$y)
testdensity$x
testdensity$y

?density
testdensity
nrow(db_full4new_stim_screen_pupil_nopract_no_outl)

db_full4new_stim_screen_pupil_nopract_no_outl%>%
    subset(subject>303)%>%
    group_by(ssid)%>%
    mutate(density_1= density(pup_basCor))

install.packages('scales')

scales::rescale(0)
db_full4new_stim_screen_pupil_nopract_no_outl%>%
  group_by(ssid)%>%
  mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  mutate(rescaled_y = scales::rescale_max(testdensity_y, to = c(0,.4)))%>%
  ggplot(aes(x = pup_basCor))+
  stat_function(fun=dgamma, args=list(shape=1, scale=.5))+
  # geom_density()+
  geom_jitter(aes( x = pup_basCor_mean, y = rescaled_y),
              width = .1, height = 0)
  # stat_function(fun=dgamma, args=list(shape=1, scale=.5))
  geom_jitter()
  stat_summary(geom = "pointrange")

testdensity$data.name

 c(density(db_full4new_stim_screen_pupil_nopract_no_outl$ssid,n=n)$x, density(bar,n=n)$x)

```

```{r cars}
summary(cars)
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.