### 2022-02-03 Subj_Obj_Ar_Bayes_Prev1.Rmd
timeseries bayes prevalence on pupil
added a way to visualise the bayesian prevalence on the timeseries of pupil data

---
title: "Subj_Arousal_bayes_prevalence"
author: "Helio"
date: "02/02/2022"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

Subj_Arousal_bayes_prevalence
pupil_arousal_findings$pupil_from_ar_slope_vale <- lmer(pup_basCor ~ (valence_c* arousal_c) +
                                                    Mean_gray_z  +
                                                    (1  | ssid)+
                                                    (0+ arousal_c | ssid),
                                                  REML = FALSE,
                      data = db_full4new_stim_screen_pupil_nopract%>%
                         subset(Group == "NT")%>%
                        subset(pupil_outlier == FALSE)%>%
                        subset(arousal_outler == FALSE))

# remeber this does contain pratice, juts a leftover of an old name not to break code
# db_full4new_stim_screen_pupil_nopract_no_outl$tNo

# 1st we need to conduct the test within each participant and store the p value

# let's do lm or lmer
# stoe just the data we need
db_full4new_stim_screen_pupil_nopract_no_outl<- db_full4new_stim_screen_pupil_nopract%>%
                         subset(Group == "NT")%>%
                        subset(pupil_outlier == FALSE)%>%
                        subset(arousal_outler == FALSE)

# 2.  LEVEL 1
#

# 2.1. Within-participant statistical test
# This loop performs within-participant statistical test. Here, a t-test for
# non-zero mean which is the simplest statistical test. In general, any
# statistical test can be used at Level 1.
Nsub = length(unique(db_full4new_stim_screen_pupil_nopract_no_outl$subject))

subject_id<- unique(db_full4new_stim_screen_pupil_nopract_no_outl$subject)

# p <- vector(mode="numeric",length=Nsub) #original
# instead for our case create a matrix
p <- matrix(data = NA, nrow = Nsub, ncol = 6) #

# t
# si = 1
# rm(si)

# p <- vector(mode="numeric",length=Nsub) # 
# for( si in 1:Nsub ){
#   t = t.test(rawdat[,si], mu=0)
#   p[si] = t[3]
# }

# test
# lmoutputtest<- summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
#    data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,subject == 320)))
# 
# lmoutputtest
# lmoutputtest[["coefficients"]][,4]

# this fucntion fits a multiple regression per individual and returns the summary
# then we stoe the p valus for each term in a matrix yo ue in level 2

db_full4new_stim_screen_pupil_nopract_no_outl$pup_resid

estimate <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  estimate[si,1:5] = t(matrix(t[["coefficients"]][,1])) 
  estimate[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(estimate)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(estimate)

# save estimates and se

# instead for our case create a matrix
se <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  se[si,1:5] = t(matrix(t[["coefficients"]][,2])) 
  se[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(se)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")

View(se)

t$coefficients[,4]

# now p values
p <- matrix(data = NA, nrow = Nsub, ncol = 6) #

for(si in 1:Nsub ){
  t = summary(lm(pup_basCor ~ (valence_c* arousal_c) + Mean_gray_z,
  data = subset(db_full4new_stim_screen_pupil_nopract_no_outl,
                subject == subject_id[si])))
  p[si,1:5] = t(matrix(t[["coefficients"]][,4])) 
    p[si,6] = subject_id[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(p)<- c("Intercept", "valence_c", "arousal_c", "Mean_gray_z", "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(p)
```

```{r}

# p holds p-values of test for each participant
alpha = 0.05  
indsig = p<alpha

# the binary variable indsig indicates whether the within-participant
# t-test is, or not, significant for each participant (1 entry for each
# participant)

# 2.2. Loading within-participant statistical test.
# You can also load your own within-participant statistical test results here.
# Load binary results into binary indsig vector with one entry per participant. 
# See also example_csv.R for an example.

#
# 3.  LEVEL 2
#

# Bayesian prevalence inference is performed with three numbers: 
# k, the number of significant participants (e.g. sum of binary indicator
# variable)
# n, the number of participants in the sample
# alpha, the false positive rate
n <- Nsub
# table(indsig)

# lets start with valence
k_2 <- sum(indsig[,2])
table(indsig[,2])

?binom.test
binom.test(k_2, 55, 0.05, alternative="two.sided",  )

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
bayesprev_posterior()
pdf_2 <- bayesprev_posterior(xvals, k_2, n) # function needed
pdf_2

plot(xvals, pdf_2, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_2, n)
pmap = bayesprev_posterior(xmap, k_2, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_2, n)
pbound = bayesprev_posterior(xbound, k_2, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_2, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_2, n)
h2 = bayesprev_posterior(i2, k_2, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# try arosual

```

```{r}
k_3 <- sum(indsig[,3])
k_3
# k_3/55 (20%)
12/55

binom.test(k_3, 55, 0.05, alternative="two.sided") 
# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
pdf_3 <- bayesprev_posterior(xvals, k_3, n) # function needed
pdf_3

plot(xvals, pdf_3, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_3, n)
pmap = bayesprev_posterior(xmap, k_3, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_3, n)
pbound = bayesprev_posterior(xbound, k_3, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_3, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_3, n)
h2 = bayesprev_posterior(i2, k_3, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

brighteness
```{r}
k_4 <- sum(indsig[,4])
k_4/55

# plot posterior distribution of population prevalence

xvals <- seq(0, 1, .01)
pdf_4 <- bayesprev_posterior(xvals, k_4, n) # function needed
pdf_4

plot(xvals, pdf_4, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_4, n)
pmap = bayesprev_posterior(xmap, k_4, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_4, n)
pbound = bayesprev_posterior(xbound, k_4, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_4, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_4, n)
h2 = bayesprev_posterior(i2, k_4, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# is the p value just a bonomial test on the counts of significants and non significant results?
```

<!-- interaction -->

<!-- arousal valence-->

```{r}
k_5 <- sum(indsig[,5])

# plot posterior distribution of population prevalence

xvals <- seq(0, 1, .01)
pdf_5 <- bayesprev_posterior(xvals, k_5, n) # function needed
pdf_5

plot(xvals, pdf_5, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_5, n)
pmap = bayesprev_posterior(xmap, k_5, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_5, n)
pbound = bayesprev_posterior(xbound, k_5, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_5, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_5, n)
h2 = bayesprev_posterior(i2, k_5, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

Now Level 2 test

```{r}
#    Load the nonlinear equation solver
   library(nleqslv)
 #  Load the required functions from the current working directory
source("bayesprev.R")

  # load the data
  sigdat = p
    # read.csv("example_data.csv", header =FALSE);

  alpha = 0.05                       #   this specifies the alpha value used for the within-unit tests
  Ntests = nrow(sigdat);        #   number of tests (e.g. participants)
  Nsigtests_3 = sum(sigdat[,3]);    #   number of significant tests

  # Plot the posterior  pdf for the population prevalence proportion

  xvals <- seq(0, 1, .01)
  pdf_3.1 <- bayesprev_posterior(xvals, Nsigtests_3, Ntests)
  plot(xvals, pdf_3.1, type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

  # 0.95 lower bound for the population prevalence proportion

  b = bayesprev_bound(0.05, Nsigtests_3, Ntests)
  print(b)

  # MAP maximum a posteriori [MAP estimate of the population prevalence proportion

  m = bayesprev_map(Nsigtests_3, Ntests)
  print(m)

  # 96% HPDI for the population prevalence proportion highest posterior density intervals (HPDIs)

  int = bayesprev_hpdi(0.96, Nsigtests_3, Ntests)
  print(int)

  # Example of possible Bayesian prevalence analyses on a common plot.

  xvals <- seq(0, 1, .01)
  pdf_3.1 <- bayesprev_posterior(xvals, Nsigtests_3, Ntests)
  plot(xvals, pdf_3.1, type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

  # Add the MAP estimate as a point

  xmap = bayesprev_map(Nsigtests_3, Ntests)
  pmap = bayesprev_posterior(xmap, Nsigtests_3, Ntests)
  points(xmap, pmap, cex =2, col= "red", pch=16)

  # Add the .95 lower bound as a vertical line

  xbound = bayesprev_bound(0.05, Nsigtests_3, Ntests)
  pbound = bayesprev_posterior(xbound, Nsigtests_3, Ntests)
  lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

  # Add the 0.96 HPDI

  int = bayesprev_hpdi(0.96, Nsigtests_3, Ntests)
  i1 = int[1]
  i2 = int[2]
  h1 = bayesprev_posterior(i1, Nsigtests_3, Ntests)
  h2 = bayesprev_posterior(i2, Nsigtests_3, Ntests)
  lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

```{r}

range(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor)
?plot()
  plot(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor, 
       type ="l", xlab = expression(gamma), ylab ="Posterior density", lwd=3)

testdensity<-density(db_full4new_stim_screen_pupil_nopract_no_outl$pup_basCor, n = 55)

db_full4new_stim_screen_pupil_nopract_no_outl<- db_full4new_stim_screen_pupil_nopract_no_outl%>%
  group_by(ssid)%>%
  mutate(testdensity_x = density(pup_basCor, n = 1)$x,
         testdensity_y = density(pup_basCor, n = 1)$y)

db_full4new_stim_screen_pupil_nopract_no_outl%>%
  subset(ssid == 300)%>%
  # group_by(ssid)%>%
  mutate(testdensity_x = density(pup_basCor)$x,
         testdensity_y = density(pup_basCor)$y)
testdensity$x
testdensity$y

?density
testdensity
nrow(db_full4new_stim_screen_pupil_nopract_no_outl)

db_full4new_stim_screen_pupil_nopract_no_outl%>%
    subset(subject>303)%>%
    group_by(ssid)%>%
    mutate(density_1= density(pup_basCor))

install.packages('scales')

scales::rescale(0)

bind_bayes1 <- left_join(as.data.frame(p), as.data.frame(estimate), by = "sid")
bind_bayes2 <- left_join(as.data.frame(bind_bayes1), as.data.frame(se), by = "sid")
bind_bayes2

bind_test<- bind_cols(p, estimate, se)

db_full4new_stim_screen_pupil_nopract_no_outl%>%
  group_by(ssid)%>%
  mutate(pup_basCor_mean = mean(pup_basCor, na.rm = TRUE))%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  mutate(rescaled_y = scales::rescale_max(testdensity_y, to = c(0,.4)))%>%
  ggplot(aes(x = pup_basCor))+
  # stat_function(fun=dgamma, args=list(shape=1, scale=.5))+
  geom_jitter(aes( x = pup_basCor_mean, y = rescaled_y),
              width = .1, height = 0)

bind_bayes2$xdensity<- testdensity[["x"]]
bind_bayes2$ydensity<- testdensity[["y"]]

testgensity

bind_bayes2 %>%

  group_by(sid)%>%
  # mutate(testdensity_x = density(arousal_c.x, n = 2)$x,
  #        testdensity_y = density(arousal_c.x, n = 2)$y)%>%

  ggplot(aes(x = arousal_c.x))+
   # stat_function(fun=dgamma, args=list(shape=1))
  geom_density() +
  geom_point(aes(x = xdensity, y =ydensity))

bind_bayes2%>%
  ggplot(aes(as.numeric(as.factor(as.integer(sid))), arousal_c.y, color = arousal_c.x<.05))+
  geom_point()+
  geom_pointrange(aes(ymin = arousal_c.y- arousal_c, ymax = arousal_c.y+ arousal_c ))

bind_bayes2%>%
  ggplot(aes(as.numeric(as.factor(as.integer(sid))), valence_c.y, color = valence_c.x<.05))+
  geom_point()+
  geom_pointrange(aes(ymin = valence_c.y- valence_c, ymax = valence_c.y+ valence_c ))

bind_bayes2%>%
  ggplot(aes(as.numeric(as.factor(as.integer(sid))), `valence:arousal.y` , 
             color = `valence:arousal.x` <.05))+
  geom_point()+
  geom_pointrange(aes(ymin = `valence:arousal.y` - `valence:arousal`, 
                      ymax = `valence:arousal.y` + `valence:arousal` ))

```

Try with the timecourse now.
start with only one predictor - arousal

then plot the effect size in terms of a heatmap

```{r}

tmp.df4_full_stim_downs_jun2021_60bins_NT<- subset(tmp.df4_full_stim_downs_jun2021_60bins,
                                    Group == "NT")

b = 1

subject_id2<- unique(tmp.df4_full_stim_downs_jun2021_60bins_NT$ssid)
s = 1

b = 1

testtimecourseforbayprev<-   subset(tmp.df4_full_stim_downs_jun2021_60bins_NT,
                                    ssid =  subject_id[s],
                                    timebin_no == nbins[b])

unique(testtimecourseforbayprev$timebin_no)

tmp.df4_full_stim_downs_jun2021_60bins_NT%>%
  group_by(ssid, timebin_no)%>%
  summarise_at(c("pup_basCor"), mean, na.rm = TRUE)

# try and figure out how many timebins per participant we need to cerate to hold data
tmp.df4_full_stim_downs_jun2021_60bins_NT%>%
  subset(timebin_no<56)%>%
  group_by(ssid, timebin_no)%>%
  summarise_at(c("pup_basCor"), mean, na.rm = TRUE) %>%
    group_by(ssid)%>%
  mutate(n_timebin = n())
  # total of 3025 rows
  #  summarise_at(c("pup_basCor", "n_timebin"), mean, na.rm = TRUE)

tmp.df4_full_stim_downs_jun2021_60bins_NT_55<- tmp.df4_full_stim_downs_jun2021_60bins_NT%>%
  subset(timebin_no<56)

# create the dataset to hold participant, timebin code, statistics for each timebin
# seems like the majority have 55

nsubj_2 = 55
n_bins_2 = 55

timecourse_60bin_result_df_for_bp <-timecourse_60bin_result_df <- data.frame(
  ssid = rep(NA, nsubj_2*n_bins_2),
  timebins= rep(NA, nsubj_2*n_bins_2), 
                                       Estimate_ar= rep(NA, nsubj_2*n_bins_2), 
                                       se_ar= rep(NA, nsubj_2*n_bins_2), 
                                       t_ar=rep(NA, nsubj_2*n_bins_2), 
                                       p_ar=rep(NA, nsubj_2*n_bins_2)
                                       # Estimate_vl= rep(NA, nsubj_2*n_bins_2), 
                                       # t_vl=rep(NA, nsubj_2*n_bins_2), 
                                       # p_vl=rep(NA, nsubj_2*n_bins_2),
                                       # Estimate_br= rep(NA, nsubj_2*n_bins_2), 
                                       # t_br= rep(NA, nsubj_2*n_bins_2), 
                                       # p_br= rep(NA, nsubj_2*n_bins_2),
                                       # Estimate_ar_vl= rep(NA, nsubj_2*n_bins_2), 
                                       # t_ar_vl= rep(NA, nsubj_2*n_bins_2), 
                                       # p_ar_vl= rep(NA, nsubj_2*n_bins_2),
                                       # cov= as.character(rep(NA, nsubj_2*n_bins_2)
                                                         )
timecourse_60bin_result_df_for_bp

# now fit level 1 within individual tests
0 = 1 # remember to always rezero it
s=2
b = 3
# nbins

n_bins_3<- unique(tmp.df4_full_stim_downs_jun2021_60bins_NT_55$timebin_no)
subject_id_3<- unique(tmp.df4_full_stim_downs_jun2021_60bins_NT_55$ssid)

s = 1
b = 2

for (s in 1:length(subject_id2)) {
  for (b in 1:length(n_bins_3)) {

    message(sprintf("$$$$$RUNING ssid %s timebin %i", subject_id2[s], n_bins_3[b]))
        lm_bin_pup  <- lm(pup_basCor ~ arousal_c,
                      data = subset(tmp.df4_full_stim_downs_jun2021_60bins_NT_55,
                                    ssid ==  subject_id2[s] &
                                    timebin_no == n_bins_3[b]))
        # save the summary    
        lm_bin_pup_summary<- summary(lm_bin_pup)

        # store statistics for particiapnt and timebin
        timecourse_60bin_result_df_for_bp[i,1]<- subject_id[s] #ssid 
        timecourse_60bin_result_df_for_bp[i,2]<- n_bins_3[b] #timebin
          # timecourse_60bin_result_df[b,1]<- tmp.df4_full_stim_downs_jun2021_60bins$timerezero3[b] 
            timecourse_60bin_result_df_for_bp[i,3] <- lm_bin_pup_summary$coefficients[2] #estimate, note to avoid intercept start on 2 for estimate, then 4 for se, the 6 for t and so on
            timecourse_60bin_result_df_for_bp[i,4] <- lm_bin_pup_summary$coefficients[4] #se
            timecourse_60bin_result_df_for_bp[i,5] <- lm_bin_pup_summary$coefficients[6] #t
             timecourse_60bin_result_df_for_bp[i,6] <- lm_bin_pup_summary$coefficients[8] #p

       i = i+1
     # b = b+1
  }

  }

timecourse_60bin_result_df_for_bp<- timecourse_60bin_result_df_for_bp[442:nrow(timecourse_60bin_result_df_for_bp),]

# heatmap

library(viridis)
install.packages('hrbrthemes')

library(hrbrthemes)
timecourse_60bin_result_df_for_bp%>%
  ggplot(aes(y = as.factor(as.numeric(ssid)), x = timebins))+
  geom_tile(aes(fill = Estimate_ar))+
  scale_fill_viridis(discrete=FALSE) 
  theme_ipsum()
  theme_classic()

  ?p.adjust
  timecourse_60bin_result_df_for_bp$p_ar_adjust<- p.adjust(timecourse_60bin_result_df_for_bp$p_ar,
                                                           method = "holm", n = 55)

timecourse_60bin_result_df_for_bp<-  timecourse_60bin_result_df_for_bp%>%
    group_by(ssid)%>%
    mutate(p_adjusted = p.adjust(p_ar, method = "holm", n = n()))

timecourse_60bin_result_df_for_bp %>%
  ggplot(aes(y = as.factor(as.numeric(ssid)), x = timebins/10))+
  geom_tile(aes(fill = p_adjusted))+
  scale_fill_viridis(discrete=FALSE) +
  # theme_ipsum()
  theme_classic()

timecourse_60bin_result_df_for_bp %>%
  ggplot(aes(x = timebins/10, y = t_ar))+
  stat_summary(geom = "line", size = 2)+
  stat_summary( geom = "errorbar", mult = 1) 
  stat_summary()

  geom_tile(aes(fill = p_adjusted))+
  scale_fill_viridis(discrete=FALSE) +
  # theme_ipsum()
  theme_classic()

  timecourse_60bin_result_df_for_bp %>%
  ggplot(aes(x = timebins/10, y = t_ar))+
  stat_summary(geom = "line", size = 2)+
  stat_summary( geom = "errorbar", mult = 1) 
  stat_summary()

  geom_tile(aes(fill = p_adjusted))+
  scale_fill_viridis(discrete=FALSE) +
  # theme_ipsum()
  theme_classic()

  #########
  p

  # flag max
  timecourse_60bin_result_df_for_bp<- timecourse_60bin_result_df_for_bp%>%
    group_by(ssid)%>%
   mutate(max_T =  max(t_ar))

  # new t = 3.5

timecourse_60bin_result_df_for_bp%>%
  group_by(ssid)%>%
  summarize_if(is.numeric, min, na.rm = TRUE)%>%
  ggplot(aes(x = as.factor(as.numeric(ssid)), y = p_adjusted))+
  geom_point(size = 2, color = "blue")+
  geom_segment( aes(x=as.factor(as.numeric(ssid)), xend=as.factor(as.numeric(ssid)), 
                    y=0, yend=p_adjusted),color = "blue")+
  geom_hline(yintercept = .05, linetype = "dashed")+
  geom_hline(yintercept = .1, linetype = "dashed", color = "red")+
  coord_flip()+
  theme_bw()

timecourse_60bin_result_df_for_bp_minp<- timecourse_60bin_result_df_for_bp%>%
  group_by(ssid)%>%
  summarize_at(c("p_adjusted"), min, na.rm = TRUE)%>%
  mutate(indsign = p_adjusted<.05)

View(timecourse_60bin_result_df_for_bp_minp)

##############################

# test prevalence

# step 2 - count significance

# p holds p-values of test for each participant
alpha = 0.05

# the binary variable indsig indicates whether the within-participant
# t-test is, or not, significant for each participant (1 entry for each
# participant)

# 2.2. Loading within-participant statistical test.
# You can also load your own within-participant statistical test results here.
# Load binary results into binary indsig vector with one entry per participant. 
# See also example_csv.R for an example.

#
# 3.  LEVEL 2
#

# Bayesian prevalence inference is performed with three numbers: 
# k, the number of significant participants (e.g. sum of binary indicator
# variable)
# n, the number of participants in the sample
# alpha, the false positive rate

unique(timecourse_60bin_result_df_for_bp_minp$ssid)
n <- nrow(timecourse_60bin_result_df_for_bp_minp)
# table(indsig)

# lets start with valence
k_3_pupilts <- sum(timecourse_60bin_result_df_for_bp_minp$indsign)

# ?binom.test - added by helio - is the number0f significant particpants significant against chance
binom.test(k_3_pupilts, n, 0.05, alternative="two.sided")  #pp-value = 0.01932

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)

# bayesprev_posterior()
pdf_3_pupts <- bayesprev_posterior(xvals, k_3_pupilts, n) # function needed
pdf_3_pupts

plot(xvals, pdf_3_pupts, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_3_pupilts, n)
pmap = bayesprev_posterior(xmap, k_3_pupilts, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_3_pupilts, n)
pbound = bayesprev_posterior(xbound, k_3_pupilts, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_3_pupilts, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_3_pupilts, n)
h2 = bayesprev_posterior(i2, k_3_pupilts, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# try arosual

# get prevalenc for the timeseries
# for ech bin do the count
# we need xmap - center pf the MAP on x axis
# and k - # k, the number of significant participants (e.g. sum of binary indicator
# variable)

# create a dataframe fro timebins

timecourse_60bin_forprevalence <-timecourse_60bin_result_df <- data.frame(
  # ssid = rep(NA, n_bins_2), # we just need bins not sids
  timebins= rep(NA, n_bins_2), 
  k_tb= rep(NA, n_bins_2), 
  xmap = rep(NA, n_bins_2),
  pmap = rep(NA, n_bins_2),
  # Add the .95 lower bound as a vertical line
xbound = rep(NA, n_bins_2),
  # bayesprev_bound(0.95, k_3_pupilts, n)
pbound = rep(NA, n_bins_2),
  # bayesprev_posterior(xbound, k_3_pupilts, n)
    # bayesprev_posterior(xmap, k_3_pupilts, n)
t = rep(NA, n_bins_2))

# create the k in advance

timecourse_60bin_result_df_for_bp<- timecourse_60bin_result_df_for_bp %>%
  group_by(timebins)%>%
  mutate(k_3_pupilts = sum(p_adjusted<alpha),
        xmap = bayesprev_map(k_3_pupilts, n()),
        pmap = bayesprev_posterior(xmap, k_3_pupilts, n()),
        i1 = bayesprev_hpdi(0.96, k_3_pupilts, n)[1],
         i2 = bayesprev_hpdi(0.96, k_3_pupilts, n)[2],
        )

  for (b in 1:length(n_bins_3)) {
    # for (s in 1:length(subject_id2)) {

    message(sprintf("$$$$$RUNING timebin %i", n_bins_3[b]))

    timecourse_60bin_forprevalence[i,1]<- n_bins_3[b] #timebin
    timecourse_60bin_forprevalence[i,2]<- sum(timecourse_60bin_result_df_for_bp$indsign)

        lm_bin_pup  <- lm(pup_basCor ~ arousal_c,
                      data = subset(tmp.df4_full_stim_downs_jun2021_60bins_NT_55,
                                    ssid ==  subject_id2[s] &
                                    timebin_no == n_bins_3[b]))
        # save the summary    
        lm_bin_pup_summary<- summary(lm_bin_pup)

        # store statistics for particiapnt and timebin
        timecourse_60bin_result_df_for_bp[i,1]<- subject_id[s] #ssid 
        timecourse_60bin_result_df_for_bp[i,2]<- n_bins_3[b] #timebin
          # timecourse_60bin_result_df[b,1]<- tmp.df4_full_stim_downs_jun2021_60bins$timerezero3[b] 
            timecourse_60bin_result_df_for_bp[i,3] <- lm_bin_pup_summary$coefficients[2] #estimate, note to avoid intercept start on 2 for estimate, then 4 for se, the 6 for t and so on
            timecourse_60bin_result_df_for_bp[i,4] <- lm_bin_pup_summary$coefficients[4] #se
            timecourse_60bin_result_df_for_bp[i,5] <- lm_bin_pup_summary$coefficients[6] #t
             timecourse_60bin_result_df_for_bp[i,6] <- lm_bin_pup_summary$coefficients[8] #p

       i = i+1
     # b = b+1
  }

  }

timecourse_60bin_result_df_for_bp%>%
  group_by(timebins)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(timebins, t_ar))+
  geom_line(color = "blue")+
  geom_line(aes(y = xmap, size = 2))+
   # scale_y_continuous(
   #  "mpg (US)",
   #  sec.axis = sec_axis(~ . * 1.20, name = "mpg (UK)")
  )

timecourse_60bin_result_df_for_bp%>%
  group_by(timebins)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  ggplot(aes(timebins, xmap))+
  geom_line(color = "blue", size = 2)+
   # geom_line(aes(y = pmap),color = "blue", linetype = "dashed")+
   # geom_line(aes(y = xbound),color = "red", linetype = "dashed")+
    geom_line(aes(y = i1),color = "blue", linetype = "dashed")+
          geom_line(aes(y = i2),color = "blue", linetype = "dashed")+
  stat_summary(aes(fun.ymin=i1, fun.ymax=i2), geom="ribbon", alpha=0.8) +
  theme_bw()

  geom_line(aes(y = xmap, size = 2))+

```

p.adjust(.05, method = "bonferroni", 55)

timecourse_60bin_result_df_for_bp$p_adjusted_count<- timecourse_60bin_result_df_for_bp$p_adjusted< alpha

timecourse_60bin_result_df_for_bp

  # p holds p-values of test for each participant
alpha = 0.05  
indsig = p<alpha

# the binary variable indsig indicates whether the within-participant
# t-test is, or not, significant for each participant (1 entry for each
# participant)

# 2.2. Loading within-participant statistical test.
# You can also load your own within-participant statistical test results here.
# Load binary results into binary indsig vector with one entry per participant. 
# See also example_csv.R for an example.

#
# 3.  LEVEL 2
#

# Bayesian prevalence inference is performed with three numbers: 
# k, the number of significant participants (e.g. sum of binary indicator
# variable)
# n, the number of participants in the sample
# alpha, the false positive rate
n <- Nsub
# table(indsig)

# lets start with valence
k_2 <- sum(indsig[,2])
table(indsig[,2])

?binom.test
binom.test(k_2, 55, 0.05, alternative="two.sided",  )

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
bayesprev_posterior()
pdf_2 <- bayesprev_posterior(xvals, k_2, n) # function needed
pdf_2

plot(xvals, pdf_2, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_2, n)
pmap = bayesprev_posterior(xmap, k_2, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_2, n)
pbound = bayesprev_posterior(xbound, k_2, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_2, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_2, n)
h2 = bayesprev_posterior(i2, k_2, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

do just scr prevalence

```{r}

 # remeber this does contain pratice, juts a leftover of an old name not to break code

# 1st we need to conduct the test within each participant and store the p value

# get the data w need

# unique(db_full4new_stim_screen_pupil_nopract_nt_304$tNo)
db_full4new_stim_screen_pupil_nopract_nt_304

# 2.  LEVEL 1
#

# 2.1. Within-participant statistical test
# This loop performs within-participant statistical test. Here, a t-test for
# non-zero mean which is the simplest statistical test. In general, any
# statistical test can be used at Level 1.
Nsub_scr = length(unique(db_full4new_stim_screen_pupil_nopract_nt_304$subject))

subject_id_scr<- unique(db_full4new_stim_screen_pupil_nopract_nt_304$subject)

# p <- vector(mode="numeric",length=Nsub) #original
# instead for our case create a matrix
p_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #
e_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for estimate
se_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for se
tval_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for t value

si = 1
# estimate = effect size when centered
for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  e_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,1])) 
  e_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(e_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(e_scr)

# s se
  for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  se_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,2])) 
  se_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(se_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  View(se_scr)

t$coefficients[,4]

# now p values
  for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  p_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,4])) 
  p_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(p_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  View(p_scr)

  # t values

    for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  tval_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,3])) 
  tval_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(tval_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  View(tval_scr)

```

step 2 - count significance
```{r}

# p holds p-values of test for each participant
alpha = 0.05  
indsig_scr = p_scr <alpha
View(indsig_scr)

# the binary variable indsig indicates whether the within-participant
# t-test is, or not, significant for each participant (1 entry for each
# participant)

# 2.2. Loading within-participant statistical test.
# You can also load your own within-participant statistical test results here.
# Load binary results into binary indsig vector with one entry per participant. 
# See also example_csv.R for an example.

#
# 3.  LEVEL 2
#

# Bayesian prevalence inference is performed with three numbers: 
# k, the number of significant participants (e.g. sum of binary indicator
# variable)
# n, the number of participants in the sample
# alpha, the false positive rate
n <- Nsub_scr
# table(indsig)

# lets start with valence
k_2_scr <- sum(indsig_scr[,2])

# ?binom.test - added by helio - is the number0f significant particpants significant against chance
binom.test(k_2_scr, n, 0.05, alternative="two.sided")  #p-value = 0.04107

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)

# bayesprev_posterior()
pdf_2_scr <- bayesprev_posterior(xvals, k_2_scr, n) # function needed
pdf_2_scr

plot(xvals, pdf_2_scr, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_2_scr, n)
pmap = bayesprev_posterior(xmap, k_2_scr, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_2_scr, n)
pbound = bayesprev_posterior(xbound, k_2_scr, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_2_scr, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_2_scr, n)
h2 = bayesprev_posterior(i2, k_2_scr, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

# try arosual

```

arousal scr

```{r}

k_3_scr <- sum(indsig_scr[,3])

# ?binom.test - added by helio - is the number0f significant particpants significant against chance
binom.test(k_3_scr, n, 0.05, alternative="two.sided")  #p-value = 0.04107

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)

# bayesprev_posterior()
pdf_3_scr <- bayesprev_posterior(xvals, k_3_scr, n) # function needed
pdf_3_scr

plot(xvals, pdf_3_scr, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_3_scr, n)
pmap = bayesprev_posterior(xmap, k_3_scr, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_3_scr, n)
pbound = bayesprev_posterior(xbound, k_3_scr, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_3_scr, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_3_scr, n)
h2 = bayesprev_posterior(i2, k_3_scr, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

do the sigend test the hard oway

```{r}
?lm
p_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #
e_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for estimate
se_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for se
tval_scr <- matrix(data = NA, nrow = Nsub_scr, ncol = 5) #for t value

si = 1
# estimate = effect size when centered
for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  e_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,1])) 
  e_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(e_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  # names(matrix(t(t[["coefficients"]])))
  View(e_scr)

# s se
  for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  se_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,2])) 
  se_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(se_scr)<- c("Intercept", "valence_c", "arousal_c",  "valence:arousal", "sid")
  View(se_scr)

t$coefficients[,4]

# now p values
  for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  p_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,4])) 
  p_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(p_scr)<- c("Intercept_p", "valence_c_p", "arousal_c_p",  "valence:arousal_p", "sid")
  View(p_scr)

  # t values

    for(si in 1:Nsub_scr ){
  t_scr = summary(lm(log1p(BIO_CDA.PhasicMax+.1)  ~ (valence_c* arousal_c),
  data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                subject == subject_id_scr[si])))
  tval_scr[si,1:4] = t(matrix(t_scr[["coefficients"]][,3])) 
  tval_scr[si,5] = subject_id_scr[si]  
  # the first t transposes the data, not we need to change si to row
  }

  colnames(tval_scr)<- c("Intercept_p", "valence_c_p", "arousal_c_p",  "valence:arousal_p", "sid")
  View(tval_scr)

e_scr
tval_scr
# colnames(tval_scr)<- paste(colnames(tval_scr),"pval")
tval_scr<- as.data.frame(tval_scr)

# tval_scr$`sid pval`<- "sid"

e_p_scr<- left_join(as.data.frame(e_scr), as.data.frame(p_scr))
View(e_p_scr)

e_p_scr$valence_c_p_onetail<- e_p_scr$valence_c_p/2
e_p_scr$valence_c_p_onetail_indsig <- e_p_scr$valence_c_p_onetail<alpha

sum(e_p_scr$valence_c_p_onetail_indsig)

# do the same for pupil
```

```{r}

k_4_scr <- sum(indsig_scr[,4])
k_4_scr
# ?binom.test - added by helio - is the number0f significant particpants significant against chance
binom.test(k_4_scr, n, 0.05, alternative="two.sided")  #p-value = 0.04107

# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)

# bayesprev_posterior()
pdf_4_scr <- bayesprev_posterior(xvals, k_4_scr, n) # function needed
pdf_4_scr

plot(xvals, pdf_4_scr, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_4_scr, n)
pmap = bayesprev_posterior(xmap, k_4_scr, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_4_scr, n)
pbound = bayesprev_posterior(xbound, k_4_scr, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_4_scr, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_4_scr, n)
h2 = bayesprev_posterior(i2, k_4_scr, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```

```{r}
k_3 <- sum(indsig[,3])
k_3
# k_3/55 (20%)
12/55

binom.test(k_3, 55, 0.05, alternative="two.sided") 
# plot posterior distribution of population prevalence
# install.packages("HDInterval")
library(HDInterval) # this package is needed
xvals <- seq(0, 1, .01)
pdf_3 <- bayesprev_posterior(xvals, k_3, n) # function needed
pdf_3

plot(xvals, pdf_3, type ="l", xlab = expression(gamma), 
     ylab ="Posterior density", lwd=3)

# Add the MAP estimate as a point
# MAP = Maximum

xmap = bayesprev_map(k_3, n)
pmap = bayesprev_posterior(xmap, k_3, n)
points(xmap, pmap, cex =2, col= "red", pch=16)

# Add the .95 lower bound as a vertical line
xbound = bayesprev_bound(0.95, k_3, n)
pbound = bayesprev_posterior(xbound, k_3, n)
lines(c(xbound, xbound), c(0, pbound+0.5), col="blue", lwd=3)

# Add the 0.96 HPDI
int = bayesprev_hpdi(0.96, k_3, n)
i1 = int[1]
i2 = int[2]
h1 = bayesprev_posterior(i1, k_3, n)
h2 = bayesprev_posterior(i2, k_3, n)
lines(c(i1, i2), c(h1, h2), col ="green", lwd=3)

```