### 2022-03-03 FactoShyny.Rmd
tracking latests changes
forgot to track these for a bit

---
title: "Tryfactoshyny"
author: "Helio"
date: "15/02/2022"
output: html_document
---

```{r}
install.packages("Factoshiny")

library(Fact)

View(clustering_new_pre_PCA3.1)

Factoshiny::MFAshiny(clustering_new_pre_PCA3.1)

clustering_new_pre_PCA3.1.scale

clustering_new_pre_PCA3.1.scale$ssid<- as.factor(clustering_new_pre_PCA3.1.scale$ssid)
clustering_new_pre_PCA3.1$ssid<- as.factor(clustering_new_pre_PCA3.1.scale$ssid)
Factoshiny::MFAshiny(clustering_new_pre_PCA3.1.scale)

Factoshiny::MFAshiny(clustering_new_pre_PCA3.1.scale_agg_bystim)

Factoshiny::MFAshiny(clustering_new_pre_PCA3.1.scale_agg_bystim_scaleafteragg)

# just agg by ssid
clustering_new_pre_PCA3.1.scale_agg_bysid
Factoshiny::MFAshiny(clustering_new_pre_PCA3.1.scale_agg_bysid)
clustering_new_pre_PCA3.1.scale_aft_agg_bysid
clustering_new_pre_PCA3.1_copy

# group_by(ssid, val_cut_iaps
         # clustering_new_pre_PCA3.1

clustering_new_pre_PCA_2_agg_by_val_sel$ssid<- as.factor(clustering_new_pre_PCA_2_agg_by_val_sel$ssid)
# clustering_new_pre_PCA_2_agg_by_val_sel
Factoshiny::MFAshiny(clustering_new_pre_PCA_2_agg_by_val_sel1)

clustering_new_pre_PCA_2_agg_by_val_sel1<- clustering_new_pre_PCA_2_agg_by_val_sel%>%
  subset(!is.na(valence_split))
```