### 2022-05-28 PCA figures1.Rmd
MFA figures

---
title: "Figures"
author: "Helio"
date: "15/02/2022"
output: html_document
---

```{r}
library(FactoMineR)
library(factoextra)

newDF <- clustering_new_pre_PCA3.1[,c("ssid","stimDescription","val_cut_iaps","pup_basCor","Bio_Mean_HR_dif","log_BIO_CDA.PhasicMax","arousal","valence")]
newDF2 <- clustering_new_pre_PCA3.1[,c("stimDescription","val_cut_iaps","pup_basCor","Bio_Mean_HR_dif","log_BIO_CDA.PhasicMax","arousal","valence")]
res.MFA<-MFA(newDF,group=c(3,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=FALSE)

?plot.MFA

colnames(newDF)
colnames(newDF) <- c("subject", "stimuli", "norm", "pupil", "HR", "SCR", "arousal", "valence")

table(if_else(newDF$norm == "neg", "negative",
                    if_else(newDF$norm == "mid", "mid",
                     if_else(newDF$norm == "pos", "positive", NULL))))

newDF$norm<- if_else(newDF$norm == "neg", "negative",
                    if_else(newDF$norm == "mid", "mid",
                     if_else(newDF$norm == "pos", "positive", NULL)))

table(is.na(newDF$norm))

colnames(newDF)
res.MFA_val<-MFA(newDF,group=c(3,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=FALSE)

plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0.2',habillage='val_cut_iaps',title="Individual factor map")

# drop sid

newDF2_stim_val <- newDF[,c("stimuli","norm","pupil", "HR", "SCR", "arousal", "valence")]
res.MFA_stim_val<-MFA(newDF2_stim_val,group=c(2,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)

plot.MFA(res.MFA_stim_val)
?plot.MFA

Factoshiny::MFAshiny(newDF2_stim_val)
factoextra::fviz_mfa(res.MFA_stim_val, geom = "point",
                      choix="ind",
                      invisible= c('quali'))

library(FactoMineR)
library(tidyverse)
plot.MFA(res.MFA_stim_val, 
         choix="ind",

         label = "none",

         lab.par=TRUE,
         invisible= c('ind', 'quali'),

         # habillage='norm',
         col.ind = "cos2",
         title="Individual factor map") +
  theme_classic()
  p$graphstyle_int

newDF <- newDF2_stim_val[,c("stimuli","norm","pupil","HR","SCR","arousal","valence")]

sf = 1 # scaling factor to make it easier to change sizes of everything, in which case change here
p<- list()
p$graphstyle_int <-  theme(#base plot theme

  # axis lines
  axis.line.y = element_line(color="black", size = 1.5),
  axis.line.x = element_line(color="black", size = 1,5),

  axis.title.y=element_text(size = 14*(sf+.1), margin=margin(0,5,0,0)),
  axis.title.x=element_text(size = 14*(sf+.1), margin=margin(0,5,0,0)),

  # text
  strip.text.x = element_text(size = 15*(sf+.1),  colour = "black"),
  strip.text.y = element_text(size = 15*(sf+.1),  colour = "black"),

  text=element_text(size = 14),#family = "arial", 
  axis.text.x = element_text(size = 14*(sf+.1),colour = "black"), # family = "arial", 
  axis.text.y = element_text(size = 14*(sf+.1), colour = "black"),#√

  # panel
  # pane
  # panel.grid.major = element_line(colour="black"),
  # panel.background = element_blank(),
  # panel.background = element_rect(fill="black"),
  # panel.border = element_rect(color="black"),
  # strip shades (reco rectagles)
   # strip.background = element_rect(fill="black"),

  # legend
  legend.position = "top",
  legend.key = element_rect(colour = "transparent", fill="transparent"),
  legend.direction = "horizontal",

  legend.text = element_text(size = 10*(sf+.3)),
  # legend.title = element_text(size = 10*(sf+.3)),
  # legend line tends to be really small in print so adjust here
  # legend.key.height = unit(5, "cm"),
  legend.key.width = unit(1, "cm"),
  # lege
  # legen

  # title
  plot.title = element_text(hjust = .5)

  #legend.text = element_text(size = 10*sf),
  # legend.title=element_blank(),
  #legend.text = element_blank(),
  #axis.ticks = element_blank(),
)

newDF2_stim_val %>%
  group

# centering ensures that we don't prioritise specific participants due to their extreme values
# we still capture the covarince within participants
clustering_new_pre_PCA3.1%>%
  group_by(ssid)%>%
  mutate_if(is.numeric, scale)

newDFscale_sid_norm<- newDF%>%
  group_by(subject)%>%
  mutate_if(is.numeric, scale)%>%
  group_by(subject, norm)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

newDFscale_stim<- newDF$subject%>%
  group_by(subject, subject_orig)%>%
  mutate_if(is.numeric, scale)%>%
  group_by(stimuli, norm)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

newDFscale<- newDF%>%
  group_by(subject)%>%
  mutate_if(is.numeric, scale)

colnames(newDFscale)
res.MFA_stim_val<-MFA(newDF2_stim_val,group=c(2,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)
res.MFA_stim_val_scale<-MFA(newDFscale[,c(2,4:8)],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)
res.MFA_stim_val_scal2e<-MFA(newDFscale[,3:8],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)

FactoMineR::MFA(newDFscale[,3:8],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)

res.MFA_stim_val_scale$group$cos2
Factoshiny::MFAshiny(newDFscale)

colnames(newDFscale_sid_norm)
res.MFA_stim_val_scal2e_sidnorm <-MFA(newDFscale_sid_norm[,1:7],group=c(2,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),
                                      num.group.sup=c(1),graph=TRUE)

# stim plot

res.MFA_stim_val_scal2e_stim <-MFA(newDFscale_stim[,1:7],group=c(2,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),
                                      num.group.sup=c(1),graph=TRUE)

# BADIA

res.MFA_stim_val_scal2e_stim

clustering_new_pre_PCA2

design_notm<- factor(newDFscale_stim$norm)
library(TExPosition)

TInPosition::tepBADA.inference.battery(newDFscale_stim[,3:7], 
                                                       scale = FALSE, 
                                                       center = TRUE, 
                                                       # DESIGN = factor( clustering_new_pre_PCA2$arousal_3cat ),
                                       DESIGN = newDFscale_stim$norm,
                                                       make_design_nominal = TRUE,
                                                       group.masses = NULL, 
                                                       weights = NULL, 
                                                       graphs = TRUE, 
                                                       k = 2, 
                                                       test.iters = 1000, 
                                                       critical.value = 2)

bada_all<- TInPosition::tepBADA.inference.battery(newDFscale[,4:8], 
                                                       scale = FALSE, 
                                                       center = FALSE, 
                                                       # DESIGN = factor( clustering_new_pre_PCA2$arousal_3cat ),
                                       DESIGN = newDFscale$norm,
                                                       make_design_nominal = TRUE,
                                                       group.masses = NULL, 
                                                       weights = NULL, 
                                                       graphs = TRUE, 
                                                       k = 2, 
                                                       test.iters = 1000, 
                                                       critical.value = 2)

# sid

newDFscale_sid<- newDF%>%
  group_by(subject)%>%
  mutate_if(is.numeric, scale, scale = FALSE)%>%
  group_by(subject)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

  mean(newDFscale_sid$HR)

res.MFA_stim_val_scal2e_sid <-MFA(newDFscale_sid[,1:6],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),
                                      num.group.sup=c(1),graph=TRUE)

# get the first two components and do clustering on it

res.MFA_stim_val_scal2e_stim$global.pca

MFAstimtoKmeans <- bind_cols(as.data.frame(res.MFA_stim_val_scal2e_stim[["ind"]][["coord"]]), newDFscale_stim[,c(1:2)])
MFAsidtoKmeans <- bind_cols(as.data.frame(res.MFA_stim_val_scal2e_sidnorm[["ind"]][["coord"]]), newDFscale_sid_norm[,c(1:2)])

MFAsid2toKmeans <- bind_cols(as.data.frame(res.MFA_stim_val_scal2e_sid[["ind"]][["coord"]]), newDFscale_sid[,c(1:2)])

db_full_sum_2$subject<-as.character(db_full_sum_2$ssid)
left_join(MFAsid2toKmeans,db_full_sum_2 )%>%
  ggplot(aes(Dim.1, TASc))+
  geom_point()+
  ggpubr::stat_cor()

left_join(MFAsid2toKmeans,db_full_sum_2 )%>%
  ggplot(aes(Dim.2, TASc))+
  geom_point()+
  ggpubr::stat_cor()

left_join(MFAsid2toKmeans,db_full_sum_2 )%>%
  ggplot(aes(Dim.2, AQc))+
  geom_point()+
  ggpubr::stat_cor()

multilMFAtokmeans<-bind_cols(as.data.frame(res.MFA_stim_val_scale[["ind"]][["coord"]]), newDFscale[,c(1:3,9)])

multilMFAtokmeans$subject
cor.test(MFAstimtoKmeans$Dim.1, MFAstimtoKmeans$Dim.2)
write_csv( MFAsidtoKmeans, "MFAsidtoKmeans.csv")
write_csv( MFAstimtoKmeans, "MFAstimtoKmeans.csv")
write_csv( multilMFAtokmeans, "multilMFAtokmeans.csv")

Factoshiny::MFAshiny(newDFscale_sid_norm)

Factoshiny::MFAshiny(newDFscale_stim)

library(readr)
HCPCshinyvariableand_clusters <- read_csv("~/OneDrive - Nexus365/InteroStudy2020/analysis/DataAnalysisJanuary2020/DataAnalysisJan2020/HCPCshinyvariableand_clusters1.csv")
View(HCPCshinyvariableand_clusters)

HCPCshinyvariableand_clusters<- HCPCshinyvariableand_clusters%>%
 gather(cluster, value, -X1)

write_csv(HCPCshinyvariableand_clusters, "HCPCshinyvariableand_clusters.csv")

HCPCshinyvariableand_clusters$X1<- if_else(HCPCshinyvariableand_clusters$X1 == "norm=positive",
                                           "positive",
                                            if_else(HCPCshinyvariableand_clusters$X1 == "norm=negative", "negative",
                                                     if_else(HCPCshinyvariableand_clusters$X1 == "norm=mid", "mid",HCPCshinyvariableand_clusters$X1)))

HCPCshinyvariableand_clusters$X1<- factor(HCPCshinyvariableand_clusters$X1, levels = c("positive", "mid", "negative", "pupil", "SCR", "HR", "valence", "arousal"))

HCPCshinyvariableand_clusters%>%
  mutate(variable = as.factor(if_else(X1 == "mid", "norm",
                                      if_else(X1 == "negative", "norm",
                                              if_else(X1 == "positive", "norm", "quantitative")))))%>%
  # subset(variable == "norm")%>%

         # HCPCshinyvariableand_clusters %>%
ggplot(aes(cluster,X1,  fill = value))+
    geom_tile()+
  # facet_grid(~variable)+
  theme_classic()+
  # geom_text(aes(label = sig))+
  p$graphstyle+
  scale_fill_viridis_c("inferno")

```

```{r}
plot.MFA(res.MFA_stim_val, 
         choix="ind",
         partial='all',
         lab.par=TRUE,
         invisible= c('ind','quali'),
         select='cos2  0',
         habillage='group',
         title="Individual factor map")+
  theme_classic()
  p$graphstyle_int

  # plot the scaled within a person
mfa_plots$ind1

res.MFA_stim_val_scale
 mfa_plots$ind1 <- plot.MFA(res.MFA_stim_val_scale, 
         choix="ind",
         partial='all',
         label = "norm",
         lab.par=TRUE,
         invisible= c('ind','quali'),
         select='cos2  0',
         habillage='group',
         title="Individual factor map",
         alpha = .2,
         shadowtext = TRUE,
         cex = 1)+
  xlab(paste0("Dimension 1 (",round(mean(res.mfa_bio_vs_self2_scale$eig[1,2], na.rm = TRUE),2),"%[95%CI:",
              paste0(round(mean(eigen_first1$perc_5[eigen_first1$component == "comp 1"], na.rm = TRUE),2), ",",
                     round(mean(eigen_first1$perc_95[eigen_first1$component == "comp 2"], na.rm = TRUE),2), "])")))+
  ylab(paste0("Dimension 2 (",round(mean(res.mfa_bio_vs_self2_scale$eig[2,2], na.rm = TRUE),2),"%[95%CI:",
              paste0(round(mean(eigen_first1$perc_5[eigen_first1$component == "comp 1"], na.rm = TRUE),2), ",",
                     round(mean(eigen_first1$perc_95[eigen_first1$component == "comp 2"], na.rm = TRUE),2), "])")))+
  annotate("text", label = paste0("total var: ", round(res.mfa_bio_vs_self2_scale$eig[2,3], 2), "%"), x = 2, y = -1.3, size = 5)+
       scale_color_brewer(palette = "Dark2", labels = c("autonomic", "subjective"))+
  theme_classic()+
  p$graphstyle_int+
   guides(colour = guide_legend(override.aes = list(size=2, stroke=1.5)))

 guides(fill = guide_legend(override.aes = list(linetype = "solid", size = 1) ))

 mfa_plots$ind1
   coord_fixed(ratio = 1)

   # custom plot
 mfa_plots$ind1

 ggsave("indi_factomap.tiff",  mfa_plots$ind1, device = "tiff", width = 10, height = 6, dpi = 800)
```

```{r}
  ?geom_point
ind_inertial<- as.data.frame(res.MFA_stim_val_scale[["quali.var.sup"]][["within.inertia"]])

ind_coord<- as.data.frame(res.MFA_stim_val_scale[["quali.var.sup"]][["coord"]])

ind_inertial%>%
  mutate(stimuli = row.names(ind_inertial))%>%
  ggplot(aes(x = Dim.1, Dim.2))+
  geom_point()

ind_parial_group<-as.data.frame(res.MFA_stim_val_scale[["quali.var.sup"]][["coord.partiel"]])

# strsplit("test.fuck", "[.]")[[1]][1]
ind_parial_group<- ind_parial_group%>%
  mutate(variable = row.names(ind_parial_group))%>%
  mutate(group = if_else(grepl("autonomic",variable), "autonomic", "subjective"))%>%
  group_by(variable)%>%
  mutate(stimuli = strsplit(variable, "[.]")[[1]][1])

ind_coord<- ind_coord%>%
  mutate(stimuli = row.names(ind_coord))

ind_coord%>%
  mutate(stimuli = row.names(ind_coord))%>%
  ggplot(aes(x = Dim.1, Dim.2))+
  # coord_fixed(ratio = 1)+
  geom_point(shape = 0, size = 2)+
  ylim(-1.5,1)+
  geom_line(aes)

ind_coord2_aut<- left_join(ind_coord, subset(ind_parial_group, group == "autonomic"), by = "stimuli")
ind_coord2_aut_subj<- left_join(ind_coord2_aut, subset(ind_parial_group, group != "autonomic"), by = "stimuli")

left_join(ind_coord, ind_parial_group, by = "stimuli")

ind_coord2_aut_subj$Dim.1.x_center<- ind_coord2_aut_subj$Dim.1.x
ind_coord2_aut_subj%>%
    ggplot(aes(x = Dim.1.x, Dim.2.x))+
  # coord_fixed(ratio = 1)+
  geom_point(shape = 0, size = 2)+
   geom_point(aes(Dim.1.y, Dim.2.y), shape = 0, size = 2, color = "red")+
  geom_line(aes(Dim.1.x_center, Dim.2.y))
  scale_color_brewer(palette = "Dark2", direction = -1)

  # ylim(-1.5,1)+
  geom_l(aes(x = Dim.1.x, y = Dim.1.y))

left_join(ind_coord, ind_parial_group, by = "stimuli") %>%

    ggplot(aes(x = Dim.1.x, Dim.2.x))+
  # coord_fixed(ratio = 1)+
  geom_point(shape = 0, size = 2)+
   geom_point(aes(Dim.1.x, Dim.2.y, group = group), shape = 0, size = 2, color = "red")+
  geom_line(aes(Dim.1.x, Dim.2.y, color = group, group = group))
  scale_color_brewer(palette = "Dark2", direction = -1)

  # ylim(-1.5,1)+
  geom_l(aes(x = Dim.1.x, y = Dim.1.y))

  left_join(ind_coord, ind_parial_group, by = "stimuli") %>%
    # subset(group == "autonomic")%>%
      mutate(Dim.2centerx = Dim.2.x)%>%
     mutate(Dim.1centery = Dim.1.y)%>%
 ggplot(aes(x = Dim.1.x, Dim.2.x))+
 # coord_fixed(ratio = 1)+
# geom_point(shape = 7, size = 2, alpha = .5)+
geom_point(aes(Dim.1.y, Dim.2.y, color = group), shape = 0, size = 2)+

geom_line(aes(Dim.1.y, Dim.2.y, group = stimuli), linetype = "dashed", alpha = .3)+
    geom_text(aes(label = stimuli, x = Dim.1.x, y = Dim.2.x))+
    theme_classic()+
    p$graphstyle_int
```

# just norm

```{r}
?plot.MFA

res.MFA_stim_val_scale_norm<-MFA(newDFscale[,3:8],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)
plot.MFA(res.MFA_stim_val_scale_norm, 
         geom = "point",
         choix="ind",
         partial='all',
         # label = "norm",
         lab.par=TRUE,
         invisible= c('ind', "qua"),
         select='cos2  0',
         habillage='norm',
         title="Individual factor map",
         alpha = .2,
         shadowtext = TRUE)+
  theme_classic()+
  p$graphstyle_int
quartz()

mfa_plots$ind2_cat_val<- plot.MFA(res.MFA_stim_val_scale_norm, 
         choix="ind",
         partial='all',
         label = "norm",
         lab.par=TRUE,
         invisible= c('ind','quali'),
         select='cos2  0',
         habillage='group',
         title="Individual factor map",
         alpha = .2,
         # ellipse= TRUE,
         graph.type = "ggplot",
         cex=2) + #text size
  xlab("Dimension 1")+
   xlab("Dimension 2")+
  ggtitle("factor map")+
    xlab("Dimension 1")+
   ylab("Dimension 2")+
  scale_color_brewer(palette = "Dark2")+
    # scale_fill_brewer(palette = "Dark2")+
  theme_classic()+
  p$graphstyle_int+
  theme(legend.position = "none",
        plot.title = element_blank())

mfa_plots$ind2_cat_val+
  coord_fixed(ratio = 1)

   ggsave("indi_factomap_catval.tiff", mfa_plots$ind2_cat_val, device = "tiff", width = 10, height = 6, dpi = 800)

 mfa_plots$ind2_cat_val<-  mfa_plots$ind2_cat_val +scale_y_continuous(
  labels = scales::number_format(accuracy = 0.1))+ #force two digits in y 
     ylim(-.2,.2)

    ggsave("indi_factomap_catval.tiff", mfa_plots$ind2_cat_val, device = "tiff", width = 7, height = 5, dpi = 800)
```

mfa_plots$ind12<- mfa_plots$ind1+
    annotate("text", label = normtest$norm, x = normtest$Dim.1, y =normtest$Dim.2, size = 5, color = "blue")+
geom_point(aes(x =  normtest$Dim.1, y  = normtest$Dim.2), color = "blue", size = 5, shape = 2)

mfa_plots$ind12

# individuals
library(tidyverse)
newDFscale1<- newDFscale%>%
  subset(subject!= 900)

res.MFA_stim_val_scale_ssid<-MFA(newDFscale[,c(1,4:8)],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)

```{r}
dev.off()

mfa_plots$ind13_sid<- plot.MFA(res.MFA_stim_val_scale_ssid, 
         choix="ind",
         # geom = c("point"),
         partial='all',
         lab.par=TRUE,
         # label = "subject",
         invisible= c("ind", "quali"),
         select='cos2  0',
         habillage='group',
         title="Individual factor map"
         # alpha = .1
         )+
         # ylim = c(-1,1))

    xlab("Dimension 1")+
   ylab("Dimension 2")+
  scale_color_brewer(palette = "Dark2")+
    # scale_fill_brewer(palette = "Dark2")+
  theme_classic()+
  p$graphstyle_int+
  theme(legend.position = "none",
        plot.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())

  mfa_plots$ind13_sid
    coord_fixed(ratio = 1)
   ggsave("indi_factomap_csid.tiff",   mfa_plots$ind13_sid, device = "tiff", width = 10, height = 6, dpi = 800)

      ggsave("indi_factomap_csid.pdf",   mfa_plots$ind13_sid, device = "pdf", width = 10, height = 6, dpi = 800)

   theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
mfa_plots$ind13_sid

knitr::plot_crop(indi_factomap_csid.pdf)

```
```{r fig.height = 2, fig.width = 10}
theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
mfa_plots$ind13_sid
#your plot code
```

patchwork

```{r}
library(patchwork)

mfa_plots$ind1+mfa_plots$ind2_cat_val+ mfa_plots$ind13_sid +
  plot_layout(ncol = 3, widths = c(2,1,1), heights = c(1,1,1))

fviz_mfa_var(res.mfa, "group")

theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))

```

Groups representatin

```{r}
mfa_plots

seq(0,1,.1)
res.MFA_stim_val_scale$group$coord.sup
?  geom_label

View(res.MFA_stim_val_scale)
# considering setting the seize of the size of qualitatie s,aller than the others

test<-bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))%>%
  mutate(group = rownames(bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))))

group_df<- bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))%>%
  mutate(group = rownames(bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))))%>%
  mutate(group = factor(group, levels = c("subjective","autonomic",  "qualitative"),
                        labels = c("subjective","autonomic", "categorical")))

group_df$tyextsize<- if_else(group_df$group == "autonomic", 2,
                             if_else(group_df$group == "subjective", 2, 1))
                             ))

group_df$tyextsize

mfa_plots$groups <- bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))%>%
  mutate(group = rownames(bind_rows(as.data.frame(res.MFA_stim_val_scale$group$coord),
          as.data.frame(res.MFA_stim_val_scale$group$coord.sup))))%>%
  mutate(group = factor(group, levels = c("autonomic", "subjective","qualitative"  ),
                        labels = c("autonomic","subjective","categorical")))%>%
  ggplot(aes(x = seq(0,1,.1), y =seq(0,1,.1)))+
  geom_vline(xintercept = res.mfa_bio_vs_self2_scale$group$coord[1,1], linetype = "dashed")+
  geom_vline(xintercept = res.mfa_bio_vs_self2_scale$group$coord[2,1], linetype = "dashed")+
    geom_hline(yintercept = res.mfa_bio_vs_self2_scale$group$coord[1,2], linetype = "dashed")+
  geom_hline(yintercept = res.mfa_bio_vs_self2_scale$group$coord[2,2], linetype = "dashed")+
  geom_point(aes(x = Dim.1, y = Dim.2, shape = group, fill = group, color = group), size = 8, alpha = .5)+

   # size = tyextsize*8
  geom_text(aes(x = Dim.1, y = Dim.2+.05, label = group, color = group 
               ), size = 8)+
    # annotate("text",label = group x = Dim.1, y = Dim.2))+
  xlim(0,1)+
      theme_classic()+
      xlab("Dimension 1")+
   ylab("Dimension 2")+
  ggtitle("Groups representation")+
  scale_color_brewer(palette = "Dark2", direction = -1)+
    scale_fill_brewer(palette = "Dark2",direction = -1)+
  theme(legend.position = "none",

        # axis.text.x = element_blank(),
        # axis.text.y = element_blank(),
        plot.title = element_text(hjust = .5, size = 16*(sf+.5)))

  # theme(g) = element_blank())

  mfa_plots$groups<- mfa_plots$groups+
    scale_colour_brewer(palette = "Dark2",
                      labels=c("categorical","autonomic","subjective"))+
    scale_fill_brewer(palette = "Dark2",
                      labels=c("categorical","autonomic","subjective"))

  mfa_plots$groups  +p$graphstyle_int
```
```{r}
  mfa_plots$groups<-   mfa_plots$groups+
    # coord_fixed(ratio = 1)+
    p$graphstyle_int+
    theme(legend.title = element_blank())
  # 
  # 
  mfa_plots$groups<- mfa_plots$groups+    
    # p$graphstyle_int+
   guides(
  fill = guide_legend(

    override.aes = aes(label = "")
  )
)

mfa_plots$groups<- mfa_plots$groups+
scale_y_continuous(
  labels = scales::number_format(accuracy = 0.2))
mfa_plots$groups

 ggsave("groups.tiff", mfa_plots$groups, device = "tiff", width = 10, height = 6, dpi = 800)

  ggsave("groups2.tiff", mfa_plots$groups, device = "tiff", width = 10, height = 10, dpi = 800)
    ggsave("groups3.tiff", mfa_plots$groups, device = "tiff", width = 10, height = 8, dpi = 800)

    ggsave("groups5.tiff", mfa_plots$groups, device = "tiff", width = 7, height = 7, dpi = 800)
     ggsave("groups3.tiff", mfa_plots$groups, device = "tiff", width = 7, height = 5, dpi = 800)

     ggsave("groups3.tiff", mfa_plots$groups, device = "tiff", width = 7, height = 7, dpi = 800)
```

```{r}
mfa_plots$cor_circle<- plot.MFA(res.MFA_stim_val_scale, choix="var",habillage='group',title="Correlation circle",
         cex = 2.7) +
    xlim(-1,1)+
  scale_color_brewer(palette = "Dark2", direction = -1)+
  theme_classic()+

      xlab("Dimension 1")+
   ylab("Dimension 2")+
  # ggtitle("Groups representation")+
  scale_color_brewer(palette = "Dark2")+
    scale_fill_brewer(palette = "Dark2")+
    p$graphstyle_int+
  theme(legend.position = "none",

        # axis.text.x = element_blank(),
        # axis.text.y = element_blank(),
        plot.title = element_text(hjust = .5, size = 16*(sf+.5)))

mfa_plots$cor_circle+
  theme(plot.margin=unit(c(0,-.4,0,-.5), "null"))
mfa_plots$cor_circle

 ggsave("cor_circle.tiff",   mfa_plots$cor_circle, device = "tiff", width = 10, height = 6, dpi = 800)

```

```{r}
mfa_plots$partial<- plot.MFA(res.MFA_stim_val_scale, choix="axes",habillage='group',cex=1.35,cex.main=1.35,cex.axis=1.35)+
  xlim(-1,1)+
   theme_classic()+
     xlab("Dimension 1")+
   ylab("Dimension 2")+
  # ggtitle("Groups representation")+
  scale_color_brewer(palette = "Dark2")+
    scale_fill_brewer(palette = "Dark2")+
    p$graphstyle_int+
  theme(legend.position = "none",

        # axis.text.x = element_blank(),
        # axis.text.y = element_blank(),
        plot.title = element_text(hjust = .5, size = 16*(sf+.5)))

mfa_plots$partial

 ggsave("partial.tiff",   mfa_plots$partial, device = "tiff", width = 10, height = 6, dpi = 800)

circle_partialpatch <- (mfa_plots$cor_circle+theme(axis.text.x = element_blank(),
                                                   axis.title.x = element_blank()))/mfa_plots$partial

 ggsave("circle_partialpatch.tiff",   circle_partialpatch, device = "tiff", width = 5, height = 10, dpi = 800)

```

```{r}

mfa_plots_eigen_value_boot<-
eigen_first%>%
  mutate(component =as.factor(substr(rownames(eigen_first), 1,6)))%>% #retrieve names for components
   # gather(dimension, value_dimenion, -group, -sim_no) %>%
  group_by(component)%>%
  # subset(!is.na(eigenvalue) & component == "comp 1" )%>% #select only component 1 and remove nas I introduced in the dataframe
  mutate(perc_5 = quantile(eigenvalue,.025, na.rm = TRUE), 
         perc_95 = quantile(eigenvalue,.975, na.rm = TRUE))%>%
  ungroup()%>%
  mutate(eig_obs_dim = if_else(component == "comp 1", 
                               res.mfa_bio_vs_self2_scale$eig[1],
                               res.mfa_bio_vs_self2_scale$eig[2])) %>% #create 95% to descrive interval around the observed estimate
  mutate(Dimension =  if_else(component == "comp 1", "Dimension 1", "Dimension 2"))%>%
  subset(Dimension == "Dimension 1")%>%
  # also store observed eige for each dimensions
  # mutate(probabcheck= sum(eigenvalue>=res.mfa_bio_vs_self2_k2$eig[1])/n())
  ggplot(aes(eigenvalue))+
  geom_histogram(aes(y=..density..), alpha = .3) +  # scale histogram y

  geom_errorbar(aes(xmin = perc_5, xmax=perc_95, y = 1), color = "gray30", size = 1.5)+
  geom_point(aes(x = eig_obs_dim, y = 1),size = 3, color = "red", alpha = .3)+

  theme_classic()+
    # geom_vline(aes(xintercept = perc_95),size = 2, linetype = "dashed", alpha = .6)+
    geom_vline(xintercept = 1,size = 2, linetype = "dashed", alpha = .6)+
  # geom_vline(aes(xintercept = perc_5), linetype = "dashed")+
    # geom_vline(aes(xintercept = perc_95), linetype = "dashed")+

  geom_hline(yintercept = 0, linetype = "dashed")+
  # facet_grid(~component)+
  # coord_fixed(ratio = 1)
  ggtitle("distribution of boostraped eigenvalues for\n largest dimension (10000 samples)")+
  annotate("text", label = "95%CI", x = 1.15, y = 2, size = 5*(sf+.8))+
    p$graphstyle_int

mfa_plots_eigen_value_boot

reshufrds<- read_rds("eigen_first_reshuf.Rds")

reshufrds<- as.data.frame(reshufrds)
# install.packages(expres)
mfa_plots$resample<-
reshufrds%>%
  mutate(component =as.factor(substr(rownames(reshufrds), 1,6)))%>% #retrieve names for components
   mutate(Dimension =  if_else(component == "comp 1", "Dimension 1", "Dimension 2"))%>%
  subset(!is.na(eigenvalue))%>% #select only component 1 and remove nas I introduced in the dataframe
  group_by(component)%>%

   mutate(perc_5 = quantile(eigenvalue,.025, na.rm = TRUE), 
         perc_95 = quantile(eigenvalue,.975, na.rm = TRUE))%>%
  ungroup()%>%
    group_by(component)%>%
  mutate(eig_obs_dim = if_else(component == "comp 1", 
                               res.mfa_bio_vs_self2_scale$eig[1],
                               res.mfa_bio_vs_self2_scale$eig[2])) %>% #store observed to visualise ion the null
  mutate(prob = paste(((sum(eigenvalue>=eig_obs_dim))/n())*100, paste0("%"))) %>%
  subset(Dimension == "Dimension 1")%>%
  ggplot(aes(eigenvalue))+
  geom_histogram(aes(y=..density..),bins = 500, alpha = .3) +  # scale histogram y
  # geom_density(col = "red", alpha = .001)+
  # facet_grid(~component)+
    geom_vline(aes(xintercept = perc_95),size = 2, linetype = "dashed", alpha = .6)+
  geom_point(aes(x = eig_obs_dim, y = 1),size = 3,  color = "red", alpha = .5)+

  theme_classic()+
  ggtitle("null distribution of eigenvalues for\n largest dimension (10000 samples)")+
  annotate("text", label = "p < .001", x = 1.16, y = 3, hjust = .5, color = "red", size = 5*(sf+.8))+
   annotate("text", label = paste0(expression(alpha), " < .05"), x = 1.05, y = 20, hjust = .5, size = 5*(sf+.8), parse = TRUE)+
  p$graphstyle_int

mfa_plots_eigen_value_boot+mfa_plots$resample

mfa_plots$resample

```

```{r}
left_join(cor_first1, cor.test_obsmfa, by = c("variable", "dimension")) %>%
  subset(dimension == "Dim.1")%>%
  ggplot(aes(value_dimenion.x, color = variable)) +
  geom_histogram(aes(y=..density..), alpha=0.5,
                 position="identity")+
  geom_density(alpha=.2) +
   geom_errorbarh(aes(xmax = perc_95, xmin = perc_5, y =-2 ))+ #color = "gray30"
  # facet_grid(~dimension)+
  geom_point(aes(x = value_dimenion.y, y = -2))+
    geom_vline(xintercept = 0, color = "red", linetype = "dashed", alpha = .3, size = 1)
  # facet_grid(~dimension)+
  # geom_vline(xintercept = .5, color = "red", linetype = "dashed", alpha = .1, size = 1)+

  theme_classic()

  mfa_plots$cor_var_dimension<-
  mfa_boot_corfirst%>%
  subset(dimension == "Dim.1")%>%
    mutate(measure = factor(measure, levels = c("pupil", "SCR", "HR", "valence", "arousal")))%>%
  ggplot(aes(value_dimenion.x, fill =measure, color = measure)) +
  geom_histogram(aes(y=..density..,fill =measure,color = measure, linetype = measure),alpha=0.2,
                 position="identity")+
  geom_density(aes(fill = measure, linetype = measure), alpha = .1) +
   geom_errorbarh(aes(xmax = perc_95, xmin = perc_5, y =-2 , linetype = measure))+ #color = "gray30"
  # facet_grid(~dimension)+
  geom_point(aes(x = value_dimenion.y, y = -2))+
    geom_vline(xintercept = 0,linetype = "dashed", alpha = .5, size = 1)+
    xlim(-1,1)+
    xlab("correlation with largest dimension")+
  # facet_grid(~dimension)+
  # geom_vline(xintercept = .5, color = "red", linetype = "dashed", alpha = .1, size = 1)+
  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+

  theme_classic()+
  p$graphstyle_int+
    theme(legend.title = element_blank())

    mfa_plots$cor_var_dimension<-mfa_plots$cor_var_dimension+
      guides(colour=guide_legend(nrow=2,byrow=TRUE))

```

groups

```{r}

mfa_plots$group_boot<-

left_join(groups_first_1boot_CIcorrected2, groups_first_obs,by = c("group", "dimension"))%>%
  subset(!is.na(value_dimenion))%>%
  group_by(dimension,group)%>%
   mutate(perc_5 = bootBCa(mean(value_dimenion_obs), value_dimenion, n = 2600)[1],
         perc_95 = bootBCa(mean(value_dimenion_obs), value_dimenion, n = 2600)[2])%>%

  subset(dimension == "Dim.1")%>%
    mutate(group = if_else(group == "auto", "autonomic", "subjective"))%>%

   ggplot(aes(value_dimenion, color = group)) +

   geom_histogram(aes(y=..density..,fill =group,color = group, linetype = group),alpha=0.2,
                 position="identity")+
  geom_density(aes(fill = group, linetype = group), alpha = .1) +
  #  geom_histogram(aes(y=..density..), alpha=0.5,
  #                position="identity")+
  # geom_density(alpha=.2) +

   geom_errorbarh(aes(xmax = perc_95, xmin = perc_5, y =-1, fill =group,color = group, linetype = group))+
    geom_point(aes(x = value_dimenion_obs, y = -1))+
   # geom_point(aes(x = value_dimenion_obs, y = -3))+
       geom_vline(xintercept = 0,linetype = "dashed", alpha = .5, size = 1)+

  # facet_grid(~dimension)+
  # geom_vline(xintercept = .5, color = "red", linetype = "dashed", alpha = .1, size = 1)+
   # geom_vline(xintercept = -.5, color = "red", linetype = "dashed", alpha = .1, size = 1)+
  theme_classic()+
   # xlim(-1,1)+
    xlab("correlation with largest dimension")+
  # facet_grid(~dimension)+
  # geom_vline(xintercept = .5, color = "red", linetype = "dashed", alpha = .1, size = 1)+
  scale_color_brewer(palette = "Dark2")+
   scale_fill_brewer(palette = "Dark2")+

  theme_classic()+
  p$graphstyle_int+
    theme(legend.title = element_blank())

mfa_plots$patchbootresample <-  (mfa_plots$resample+ mfa_plots_eigen_value_boot +
  mfa_plots$group_boot + mfa_plots$cor_var_dimension)+
  plot_layout(nrow = 2)

ggsave("patch_res_boot.tiff", mfa_plots$patchbootresample,
       device = "tiff",
       width = 10, height = 10, dpi = 800)

```

patchwork

```{r}
?plot_layout
((mfa_plots$ind1 +mfa_plots$ind2_cat_val/mfa_plots$ind13_sid)+ 
    (mfa_plots$groups + mfa_plots$cor_circle / mfa_plots$partial))+
  plot_layout(nrow = 2)+
  plot_annotation(tag_levels = "A")

 # widths = c(2,1,1), heights = c(1,1)

```

```{r}

plot.MFA(res.MFA_stim_val_scale, choix="group",
         cex=2.5,
         cex.main=2.5,
         cex.axis=1.7)+
    theme_classic()+
      xlab("Dimension 1")+
   ylab("Dimension 2")+

  scale_color_brewer(palette = "Dark2", direction = -1)+
    scale_fill_brewer(palette = "Dark2", direction = -1)+
  geom_vline(xintercept = res.mfa_bio_vs_self2_scale$group$coord[1,1], linetype = "dashed")+
  geom_vline(xintercept = res.mfa_bio_vs_self2_scale$group$coord[2,1], linetype = "dashed")+
    geom_hline(yintercept = res.mfa_bio_vs_self2_scale$group$coord[1,2], linetype = "dashed")+
  geom_hline(yintercept = res.mfa_bio_vs_self2_scale$group$coord[2,2], linetype = "dashed")+

  p$graphstyle_int+

```

ssid
```{r}

unique(as.integer(as.factor(newDFscale$subject)))
newDFscale$subject_orig<- newDFscale$subject

newDFscale$subject<- as.factor(as.integer(as.factor(newDFscale$subject)))

newDFscale1<- newDFscale%>% subset(subject!= 49)
res.MFA_stim_val_scale_ssid<-MFA(newDFscale1[,c(1,4:8)],group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=TRUE)

plot.MFA(res.MFA_stim_val_scale_ssid, 
choix="ind",
         partial='all',
         label = FALSE,
         lab.par=FALSE,
         invisible= c('ind','quali'),
         select='cos2  0',
         habillage='group',
         title="Individual factor map",
         alpha = .2,
         shadowtext = FALSE,
REPEL = TRUE)+
  theme_classic()

  p$graphstyle_int+
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()
        )

?factoextra::fviz_mfa_ind

factoextra::fviz_mfa_ind(res.MFA_stim_val_scale_ssid, 
                         choice="quali",
                         axes = c(1,2),
                         geom = "point",
                          habillage= 'none',
                         col.partial = "group",
                         repel = TRUE,
         partial='all',
select.ind = list(name = "subject"),
 alpha.ind = .02
)

```

  p$graphstyle_int

# add the neg pos etc based on their coordiantes

plot.MFA(res.MFA_stim_val, choix="ind",partial='all',lab.par=TRUE,invisible= c('ind','quali'),select='cos2  0',habillage='group',title="Individual factor map")

plot.MFA(res.MFA, choix="var",habillage='group',title="Correlation circle")
plot.MFA(res.MFA, choix="group")
plot.MFA(res.MFA, choix="axes",habillage='group')

```

?factoextra::fviz_mfa_ind
factoextra::fviz_mfa_ind(res.MFA_stim_val, 
                         axes = c(1, 2),
                         geom = c("text"),

                         select.ind = list(name = c("stimuli")),
                         # col.ind = "cos2",
                         habillage = "none",
                         addEllipses = TRUE,
             # gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             # col.partial = "group"
             )

```

facto

label = "none", # hide individual labels
             habillage = iris$Species, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE # Concentration ellipses

fviz_mfa_ind(res.MFA_stim_val,
             geom = "none",
              choix=c("ind"),
               habillage='norm',
         lab.par=FALSE,
         invisible= c('ind'),

         # partial = "all"
         )

fviz_mfa_ind(res.MFA_stim_val, partial = "norm")

```

plot.MFA(res.MFA, 
         choix="ind",
         lab.par=TRUE,
         invisible= c('ind', 'quali'),
         # ellipse = 'val_cut_iaps',
         select='cos2  0.0',
         habillage='val_cut_iaps',
         title="Individual factor map",
         lab.grpe	 = TRUE)+
  theme_classic()
# boolean, if TRUE, the labels of the groups are drawn)

plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0',habillage='val_cut_iaps',title="Individual factor map")

testplot<- plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0',habillage='val_cut_iaps',title="Individual factor map")+
  theme_classic()

testplot$data

mfa_plots<- list()

factoextra::fviz_mfa_ind(res.MFA, 
                         choix="ind",

                         title="Individual factor map")
  theme_classic()

  # factor map of indivduals

quartz()
loadfonts(dev="win")
install.packages("extrafont")
library(extrafont)
extrafont::font_import()
plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0',habillage='val_cut_iaps',title="Individual factor map")+
  theme_classic()+
  p$graphstyle_int

# big lcuster by group
plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= 'quali',select='cos2  0',habillage='val_cut_iaps',title="Individual factor map")

# show that individuals are spread out
plot.MFA(res.MFA, choix="ind",
         lab.par=FALSE,
         invisible= 'quali',
         select='cos2  0',habillage='ssid',title="Individual factor map")

# we can select individuals significanlty contributing to the dimensions

newDF_sid <- clustering_new_pre_PCA3.1[,c("ssid","pup_basCor","Bio_Mean_HR_dif","log_BIO_CDA.PhasicMax","arousal","valence")]
res.MFA_sid<-MFA(newDF_sid,group=c(1,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=FALSE)
plot.MFA(res.MFA_sid, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0.2',
         habillage= 'ssid',
         title="Individual factor map")

?factoextra::fviz_mfa_ind

fviz_mfa_ind(res.MFA_sid, col.ind = "cos2", 
             choix="ind",
             invisible= c('ind','quali'),
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)

plot.MFA(res.MFA, choix="var",habillage='group',title="Correlation circle")
plot.MFA(res.MFA, choix="group")
plot.MFA(res.MFA, choix="axes",habillage='group')

newDF_sid_va <- clustering_new_pre_PCA3.1[,c("ssid","val_cut_iaps","pup_basCor","Bio_Mean_HR_dif","log_BIO_CDA.PhasicMax","arousal","valence")]
res.MFA_sid_val<-MFA(newDF_sid_va,group=c(2,3,2), type=c("n","c","c"),name.group=c("qualitative","autonomic","subjective"),num.group.sup=c(1),graph=FALSE)
plot.MFA(res.MFA_sid_val, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0.2',habillage='val_cut_iaps',title="Individual factor map")

plot.MFA(res.MFA, choix="ind",lab.par=FALSE,invisible= c('ind','quali'),select='cos2  0.2',habillage='val_cut_iaps',title="Individual factor map")
plot.MFA(res.MFA, choix="var",habillage='group',title="Correlation circle")
plot.MFA(res.MFA, choix="group")
plot.MFA(res.MFA, choix="axes",habillage='group')

```