### 2022-12-06 Power.Rmd
PowerRmd

---
title: "Power simulations"
author: "Helio"
date: '2022-12-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

summary(pupil_arousal_findings$pupil_from_arousal)

install.packages("simr")

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(simr)
library(lmerTest)
# Power for predictor 'x', (95% confidence interval):
power_models<- list()

table(is.na(db_full4new_stim_screen_pupil_nopract$arousal_c))
table(is.na(db_full4new_stim_screen_pupil_nopract$pup_basCor))

power_models$pupil <- lmer(pup_basCor ~ arousal_c +(1  | ssid), REML = FALSE,
                      data = subset(db_full4new_stim_screen_pupil_nopract, 
                                    # pupil_outlier == FALSE & 
                                     Group == "NT" & !is.na(pup_basCor) &
                                      !is.na(arousal_c)
                                    # arousal_outler == FALSE
                                    ))

db_full4new_stim_screen_pupil_nopract
power_models$pupil1<- power_models$pupil
fixef(power_models$pupil1)["arousal_c"]<- .4
sim_.4 <- powerSim(power_models$pupil1, nsim = 100)
sim_.4
# Power for predictor 'arousal_c', (95% confidence interval):
#       100.0% (96.38, 100.0)

?powerCurve

curve<- powerCurve(power_models$pupil1, nsim = 100)

plot(curve)

model3 <- extend(power_models$pupil1, along="ssid", n=1)

pc3 <- powerCurve(power_models$pupil1, along="ssid", nsim = 10)

plot(pc3)

extendeddata_along_ssid<- extend(power_models$pupil1, along="ssid", n=100)

pc_extended100 <- powerCurve(extendeddata_along_ssid, along="ssid", nsim = 100)
pc_extended100

plot(pc_extended100)

```
SCR

```{r}
power_models$BIO_CDA.PhasicMax <- lmer(BIO_CDA.PhasicMax ~ valence_c + (1|ssid) + (1|stimIAPS),
                        REML = FALSE,
                        data = subset(db_full4new_stim_screen_pupil_nopract_nt_304,
                                      !is.na(arousal_c)&
                                        !is.na(BIO_CDA.PhasicMax)))

table(is.na(db_full4new_stim_screen_pupil_nopract_nt_304$arousal_c))
table(is.na(db_full4new_stim_screen_pupil_nopract_nt_304$BIO_CDA.PhasicMax))

summary(power_models$BIO_CDA.PhasicMax)
power_models$BIO_CDA.PhasicMax1<- power_models$BIO_CDA.PhasicMax
# 0.01571715 
fixef(power_models$BIO_CDA.PhasicMax)["valence_c"]

# reverse log to get proper unites wot meaninful effect size
0.01571715

ex

fixef(power_models$BIO_CDA.PhasicMax1)["valence_c"] <- -.1

summary(power_models$BIO_CDA.PhasicMax)

summary(power_models$BIO_CDA.PhasicMax1)
sim_SCR <- powerSim(power_models$BIO_CDA.PhasicMax1,test = fixed("arousal_c:valence_c", "z"),
                    nsim = 100)

# extendedscr100<- extend(power_models$BIO_CDA.PhasicMax1, along="ssid", n=100)
# extendedscr100_int<- extend(power_models$BIO_CDA.PhasicMax1, along="ssid", n=100)

 pc_scr100<- powerCurve(extendedscr100, along="ssid", nsim = 100)
 plot(pc_scr100)

extendedscr1002<- extend(power_models$BIO_CDA.PhasicMax1, along="ssid", n=100)

# pc_scr100_2<- powerCurve(extendedscr1002, along="ssid", nsim = 100)

# extendedscr100_int@call
pc_scr100_int<- powerCurve(extendedscr1002,
                           test = fixed("arousal_c:valence_c", "z"),
                           along="ssid", nsim = 50)

plot(pc_scr100_int)
```