### 2023-01-31 anonymise.Rmd
push

---
title: "Anonymise"
author: "Helio"
date: "2023-01-21"
output: html_document
---

```{r}
colnames(clustering_new_pre_PCA3.1)
colnames(clustering_new_pre_PCA3)

unique(clustering_new_pre_PCA3.1$ssid)
clustering_new_pre_PCA3.1

table(clustering_new_pre_PCA3.1$ssid)

data_osf<- clustering_new_pre_PCA3[,c(1,4:13,16:17)]

# create a factor variable with original levels
factor_var <- as.factor(as.character(data_osf$ssid))
# create a vector of letters and numbers
letters_and_numbers <- c(letters, 0:9)
# generate unique random strings for each level
random_strings <- lapply(levels(factor_var), function(x) paste(sample(letters_and_numbers, 10, replace = TRUE), collapse = ""))
names(random_strings) <- levels(factor_var)

idsnew<- data.frame(matrix(nrow=50))
idsnew$ssid<- as.vector(unique(data_osf$ssid))
# factor_var

idsnew$ssid_rand<- as.vector(unlist(unlist(random_strings)))
library(tidyverse)
data_osf_v1<- left_join(data_osf,idsnew)%>%
  select(c(15,2:13))

data_osf_v1

saveRDS(data_osf_v1,"data_RDS_v1.RDS", compress = F)

```